### 2021-08-11 Discussion.tex
add files

\section{\label{sec:discussion}Discussion}
\label{discussion}

The class of separable covariance matrix estimators \eqref{separable} that we proposed in this paper appears to be very promising. Many existing procedures already explicitly or implicitly target this class, and our proposed estimate \eqref{proposed} of the optimal separable estimator outperforms a number of existing covariance matrix estimators. This is surprising because our approach vectorizes the matrix and therefore cannot take matrix structure, such as positive-definiteness, into account. This suggests that a vectorized approach combined with a positive-definiteness constraint may have improved performance. The resulting estimator would necessarily not be separable, because the estimate of the $jk$th entry would depend on more than just the $j$th and $k$th observed features, so the $g$-modeling estimation strategy is insufficient. More work is needed.

Though our estimator performs well in simulations and in real data, providing theoretical guarantees is difficult. In the standard mean vector estimation problem with $Y_i \sim N(\theta_i, 1)$, \citet{jiang2009general} showed that an empirical Bayes estimator based on a nonparametric maximum likelihood estimate of the prior on the $\theta_i$ can indeed asymptotically achieve the same risk as the oracle optimal separable estimator. However, this was in a simple model with a univariate prior distribution. \citet{saha2020nonparametric} extended these results to multivariate $\bs{Y}_i \sim N(\bs{\theta}_i, \bs{\Sigma}_i)$ with a multivariate prior on the $\bs{\theta}_i$, but assumed that the $\bs{Y}_i$ were independent. In contrast, our covariance matrix estimator is built from arbitrarily dependent $\bs{A}_{jk}$. These imposes significant theoretical difficulties that will require substantial work to address; we leave this for future research.

Finally, we have so far assumed that our data multivariate normal. To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gp hat} and in our proposed estimator \eqref{proposed}. If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.

### 2021-08-11 Introduction.tex
add files

\section{Introduction}
\label{introduction}

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications, such as portfolio management \citep{ledoit2003improved}, genomics \citep{schafer2005shrinkage}, and array signal processing \citep{abramovich2001locally}. However, in modern problems the number of features can be of the same order as or exceed the sample size, and the standard sample covariance matrix estimator behaves poorly in this regime. To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrix. These can roughly be divided into two groups, according to whether they impose assumptions about the structure of population covariance matrix.

Structured methods make structural assumptions about the population covariance matrix. One class models the population covariance matrix as sparse. The most common method to address this problem is thresholding \citep{rothman2009generalized, cai2011adaptive}. Penalized likelihood methods \citep{xue2012positive} can also estimate large-scale sparse covariance matrix by penalizing a log-likelihood function. Another class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension. Other common structured methods assume that the covariance matrix is banding \citep{li2017estimation} or Toeplitz matrix \citep{liu2017covariance}.

In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet can still outperform the sample covariance matrix. A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix. More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}. These shrink the eigenvalues of the sample covariance matrix toward clusters. Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global mean.

Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance. However, they modify only the sample eigenvalues and not the sample eigenvectors. It is known that sample eigenvectors are not consistent estimators of population eigenvectors when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}. This suggests that there may exist a class of unstructured estimators that can outperform nonlinear shrinkage.

%In this paper, we propose a new covariance matrix estimator which is approximately optimal among all separable estimators. Our method is motivated by the similarity between covariance estimation and vector mean estimation. Covariance matrix estimation under Frobenius loss can be viewed as a vector estimation problem by converting the matrix to a vector. \citep{cai_liu} used this motivation to generalize thresholding method in mean vector estimation to large sparse covariance matrix estimation. Nevertheless, they are not totally the same because covariance matrix has repeated entries due to its symmetry. In mean vector estimation problem, optimality has been achieved among all separable decision rules. \citep{jiang_zhang2009} and \citep{brown_greenshtein} proposed empirical Bayes methods to approximate the optimal separable estimator. We proposed to apply vector estimation methods to covariance estimation problem and expect that these methods will also have good properties here. Result in simulation and real data set shows that our estimator has good performance in different scenarios.

Here we propose a new unstructured estimator for high-dimensional covariance matrices. Our approach centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem. We do this because it allows us to use a nonparametric empirical Bayes shrinkage procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex}. We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator. Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis.

The article is organized as follows. In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach. In Section \ref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset. Finally, Section \ref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub.

### 2021-08-11 Method.tex
add files

\section{\label{sec:method}Method}
\subsection{\label{sec:compound}Compound decision problem formulation}
Suppose we have $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$. The purpose of this paper is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$ that minimizes the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.

Our proposed approach is motivated by two observations. First, \eqref{frobenius risk} shows that estimating $\bs{\Sigma}$ under Frobenius risk is equivalent to simultaneously estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under a loss function that aggregates errors across components. Second, this type of vector estimation problem has been well-studied in the compound decision literature. Thus, recent advances in vector estimation may be profitably applied to covariance matrices.

We first briefly review compound decision problems. Introduced by \citet{robbins1951asymptotically}, these problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk
\begin{equation}
\label{compound risk}
R(\bs{\theta},\bs{\delta}) = \frac{1}{n}\sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))
\end{equation}
where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is squared error loss \citep{johnstone2017gaussian}.

A key property of compound decision problems is that while a given $Y_i$ seems to offer no information about any specific $\theta_j$ when $j \ne i$, borrowing information across all components of $\bs{Y}$ to estimate $\bs{\theta}$ is superior to estimating each $\theta_i$ using the corresponding $Y_i$ alone. A classical example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. It is well-known that when $n \geq 3$, the James-Stein estimator 
%which is recognized as a breakthrough discovery. In mean vector estimation problem, given $X_i\sim N(\theta_i,\sigma^2)$, $i=1,\ldots, n$, we want to estimate the multiple mean parameters $(\theta_1,\ldots,\theta_n)$. James-Stein estimator
%\begin{equation}
%\label{js}
%\hat{\theta}_{\text{JS}} = (1-frac{N\sigma^2}{\|\bs{X}\|^2}) \bs{X}
%\end{equation}
dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$. A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.

%It has been shown that for the above risk, the combination of minimax decision rule for each parameter might be inadmissible. In fact, suppose for each $i$, there exists a minimax decision rule $\delta_i^*$ minimizing $\sup_{\theta_i\in \Omega}EL(\theta_i, \delta^*(X_i))$. However, for the decision rule $\bs{\delta}^*(X) = \{\delta_1^*(X_1),\ldots,\delta_{n}^*(X_n)\}$ there might be exists another decision rule $\bs{\delta}(\bs{X})=\{\delta_1(\bs(X)),\ldots,\delta_n(\bs(X))\}$ such that 
%\begin{equation}
%R(\bs{\theta},\bs{\delta}) \leq R(\bs{\theta},\bs{\delta^*})+o(1)
%\end{equation}

We propose to apply some of these recent vector estimation ideas to covariance matrix estimation. Clearly, covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can be viewed as a compound decision problem. Furthermore, some existing covariance matrix estimation procedures can already be interpreted as taking a vector approach. The sample covariance matrix $\bs{S}$, for instance, can be thought of as estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$ using maximum likelihood. Less trivially, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective. Their adaptive thresholding method is a version of the soft thresholding method of \citet{donoho1995adapting}, which was originally developed to estimate a sparse mean vector in the Gaussian sequence problem.

Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be interpreted as a vector estimator. The estimator is defined as
\begin{equation}
\label{linear model}
\hat{\bs{\Sigma}}_{\text{LW}} = \hat{\rho}_{\text{LW}} \bs{S} + (1-\hat{\rho}_{\text{LW}})\hat{\mu} \bs{I}
\end{equation} 
where $\hat{\mu} = \text{tr}(\bs{S})/p$, and $\hat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\hat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$. Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \eqref{frobenius risk}. We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
\begin{equation}
  \label{linear class}
  \beta_S s_{jk} + \beta_I u_{jk},
\end{equation}
where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. It is straightforward to show that
$$\hat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
is an unbiased estimate of the risk \eqref{frobenius risk}, where $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n$ is the sample variance of $s_{jk}$. The optimal estimator in this class can now be chosen by minimizing $\hat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}. Proposition \ref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \eqref{linear model}.
\begin{prop}
  \label{prop:linear}
  Define the estimator $\hat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\hat{\sigma}_{jk} = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, where $(\hat{\beta}_S, \hat{\beta}_I) = \argmin_{\beta_S, \beta_I} \hat{R}(\beta_S, \beta_I)$. Then $\hat{\bs{\Sigma}}_V$ = $\hat{\bs{\Sigma}}_{LW}$.
\end{prop}

\subsection{\label{sec:proposed}Proposed estimator}

The previous section argues that treating covariance matrix estimation as a vector estimation problem can be a fruitful strategy, but discusses only estimators linear in $s_{jk}$. We propose to consider a larger class, the class of so-called separable rules. In the standard compound decision problem of estimating $\bs{\theta}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ \citep{robbins1951asymptotically}. Here we generalize this to the problem of estimating a vectorized matrix. For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, define the class of separable rules
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} : \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})\},
\end{equation}
where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature. In other words, rules in $\mathcal{S}$ estimate the $jk$th entry of the covariance matrix using a fixed function $t$ of only observations from features $j$ and $k$.

We propose to search for the optimal estimator within $\mathcal{S}$. This is sensible because $\mathcal{S}$ includes the sample covariance $(s_{11}, \ldots, s_{pp})^\top$, the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
\[
\delta_{jk}(\bs{X}) = \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n + \beta_I I(\bs{X}_{\cdot j} = \bs{X}_{\cdot k}),
\]
and the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}. Therefore the optimal separable estimator $\bs{\delta}^\star$ that minimizes the scaled squared Frobenius risk \eqref{frobenius risk} will perform at least as well as these three estimators, and may perform better. Targeting the optimal separable rule is standard in the compound decision literature \citep{zhang2003compound}.

The optimal $\bs{\delta}^\star$ is an oracle estimator and cannot be calculated in practice, as the true risk is unknown. In the classical compound decision framework, empirical Bayes methods are used to estimate the oracle optimal separable rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}. We take a similar approach here. To simplify notation, denote $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})^\top$ as $\bs{A}_{jk}$. The density of $f(\cdot \mid \bs{\eta}_{jk})$ of $\bs{A}_{jk}$ depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation. When $r_{jk} \ne 1$, $\bs{A}_{jk}$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
\[
\bs{C}_{jk}=\begin{bmatrix}
&\sigma_j^2 & \sigma_j\sigma_k r_{jk} \\
& \sigma_j\sigma_k r_{jk}  & \sigma_k^2
\end{bmatrix}.
\]
When $r_{jk} = 1$, $\bs{A}_{jk}$ consists of mean-zero univariate normals with variances $\sigma_j^2$.

Now consider the Bayesian model
\begin{equation}
  \label{bayesian}
  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
By the fundamental theorem of compound decisions \citep{robbins1951asymptotically, jiang2009general}, this is closely related to the vectorized covariance matrix estimation problem under Frobenius risk \eqref{frobenius risk}: for any separable $\bs{\delta} \in \mathcal{S}$ \eqref{separable}, the Frobenius risk can be written as
\begin{align*}
  R(\bs{\Sigma}, \bs{\delta})
  =\,&
  \frac{1}{p^2} \sum_{j,k = 1}^p
  \int \{t(\bs{A}) - \sigma_{jk} \}^2 f(\bs{A} \mid \bs{\eta}_{jk}) d\bs{A}\\
  =\,&
  \int \int \{t(\bs{A}) - g(\bs{\eta})\}^2 f(\bs{A} \mid \bs{\eta}) dG_p(\bs{\eta}) d\bs{A}
  =
  \bb{E}[\{t(\bs{A}) -  g(\bs{\eta})\}^2],
\end{align*}
where $g(a, b, \gamma) = a b \gamma$ and the final expectation is the Bayes risk of estimating $\sigma_{jk}$. The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star(\bs{A}_{jk})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid \bs{A}\}$ minimizes the Bayes risk.

Based on this result, we propose the following empirical Bayes procedure. We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the prior $G_p$. Under the Bayesian model \eqref{bayesian}, and the working assumption that the $\bs{A}_{jk}$ are independent across $jk$, we estimate $G_p$ using
\begin{equation}
  \label{Gp hat}
  \hat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
\end{equation}
where $\mathcal{G}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$. Of course, the $\bs{A}_{jk}$ are not independent, so $\hat{G}_p$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\hat{G}_p$, we estimate the vectorized $\bs{\Sigma}$ using
\begin{equation}
  \label{proposed}
  \hat{\bs{\delta}}(\bs{X})
  =
  (\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
  \quad
  \hat{t}(\bs{A}) = \frac{\int g(\bs{\eta}) f(\bs{A} \mid \bs{\eta})d\hat{G}_p(\bs{\eta})}{\int f(\bs{A} \mid \bs{\eta}) d\hat{G}_p(\bs{\eta})}.
\end{equation}
The $\hat{t}$ estimates the Bayes rule $t^\star$ and $\hat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.

Our proposed procedure is an example of what \citet{efron2014two} calls $g$-modeling, an approach to empirical Bayes problems that proceeds by modeling the prior. A major advantage of nonparametric estiation of the prior is that it allows the data itself to determine how best to shrink the estimator. In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}. Theoretical justification of our proposed $\hat{\bs{\delta}}$ is difficult and is discussed in Section \ref{discussion}. Nevertheless, our numerical results in Section \ref{numerical results} show that in practice, our $\hat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.

\subsection{\label{implementation}Implementation}

Calculating the estimated prior $\hat{G}_p$ \eqref{Gp hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$. \citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points. The EM algorithm has traditionally been used to estimate the locations of the support points and the masses at those points \citep{laird1978nonparametric}, but this is a difficult nonconvex optimization problem.

Instead, we maximize the pairwise composite likelihood over a fixed grid of support points, similar to recent $g$-modeling procedures for standard compound decision problems; this restores convexity \citep{jiang2009general, koenker2014convex, feng2018approximate}. Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$. We can then use the EM algorithm to estimate the masses $\hat{\bs{w}}=\{\hat{w}_{1},\ldots, \hat{w}_{D}\}$ at those points via the iteration
\[
\hat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_l)}
\]
over $k$. Early stopping of the EM algorithm can be useful \citep{koenker2019comment}, and more sophisticated convex optimization procedures can be used as well \citep{koenker2014convex}. Our proposed estimator \eqref{proposed} then becomes
\[
\hat{\bs{\delta}}(\bs{X})
=
(\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
\quad
\hat{t}(\bs{A}) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}.
\]

Ideally, the grid points should be chosen to densely cover the parameter space. However, the fact that $G_p$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$. Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\hat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$. This reduces the size of the support set, but even in this case the computation complexity grows like $O(p^2)$.

Here we propose a clustering-based exemplar algorithm to further improve computational efficiency. Let $s_j$, $s_k$, and $\gamma_{jk}$ be the sample variances and correlation between the $j$th and $k$th covariates. We first apply $K$-means clustering to identify $K$ clusters among the $p (p-1) / 2$ off-diagonal sample points $(s_j,s_k,\gamma_{jk})$ and $\lceil{K^{1/2}}\rceil$ clusters among the $p$ diagonal sample points $(s_j,s_j,1)$. We then use the $K+\lceil{K^{1/2}}\rceil$ cluster centroids as our support points. We cluster the off- and on-diagonal observations separately to ensure that the support points $\bs{\xi}_\tau$ are such that $f(\bs{A}_{jk} \mid \bs{\xi}_\tau) \ne 0$ when both $j = k$ and $j \ne k$. Figure \ref{fig:sim1_frobenius} shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table \ref{tab:sim1_time} shows that they can be significantly faster.

%We constructed our support points by taking the Cartesian product of grids along each of the three dimensions of the points $\bs{\xi}_\tau = (a, b, \gamma)$. The first two components $a$ and $b$ correspond to standard deviations $\sigma_j$ and $\sigma_k$, so we assume that these are supported on $d$ equally spaced points $\min_j s_{jj} = \tilde{\sigma}_1 \leq \tilde{\sigma}_2 \ldots \leq \tilde{\sigma}_d = \max_j s_{jj}$, where $s_{jj}$ is the same variance of the $j$th feature. The last component $\gamma$ corresponds to the correlation $r_{jk}$, so we assume it is supported on $d$ equally spaced points $-1 = \tilde{\gamma}_1\leq \tilde{\gamma}_2 \ldots \leq \tilde{\gamma}_d = 1$. Therefore $d^3$ grid points would be sufficient for the support of $\bs{\xi}_\tau$. However, we can reduce the number of required points thanks to the symmetry of $f(\cdot \mid \bs{\xi}_\tau)$: $f(\cdot \mid a, b, \gamma) = f(\cdot \mid b, a, \gamma)$, and when $\gamma = 1$, $f(\cdot \mid a, b, \gamma) = f(\cdot \mid a, a, \gamma)$. Furthermore, we assume that $\gamma \ne -1$. Therefore we only consider $\bs{\xi}_\tau$ supported on
%\begin{equation}
%  \label{grid}
%  \{(a, b, \gamma) \in (\tilde{\sigma}_1, \ldots, \tilde{\sigma}_d)^2 \times (\tilde{\gamma}_2, \ldots, \tilde{\gamma}_{d - 1}) : a < b\}
%  \cup
%  \{(a, a, 1) : a  \in (\tilde{\sigma}_1, \ldots, \tilde{\sigma}_d)\}
%\end{equation}
%This requires $d + d (d-1) (d-2) / 2$ grid points.

%Ideally, the number of grid points should be chosen to be as large as possible. However, the fact that $G_p$ is multivariate poses difficulties, as for example our strategy of using a grid of $d$ points in each dimension requires a total of $D = O(d^3)$ grid points. Better computational solutions are necessary. One promising alternative is the approach of \citet{tao2014convex}, who studied the dual problem of \eqref{Gp hat} and used unequally spaced grid points.  Another alternative might be parametric or semiparametric modeling of $G_p$, along the lines of \citet{efron2019bayes}. We found that our implementation strategy is computationally feasible for $d \leq 30$, and numerical results in Section \ref{numerical results} suggest that these work well enough.

\subsection{\label{posdef}Positive definiteness correction}
Our proposed estimator \eqref{proposed} is not guaranteed to be positive-definite. To correct this, we reshape our vector estimator back into a matrix and then identify the closest positive-definite matrix. \citet{higham1988computing} and \citet{huang2017calibration} showed that the projection of a $p \times p$ symmetric matrix $\bs{B}$ onto the space of positive semi-definite matrices is
\[
P_0(\bs{B})
=
\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert
=
\bs{Q}\text{diag}\{\max(\lambda_1,0),\max(\lambda_2,0),\ldots,\max(\lambda_p,0)\}\bs{Q}^\top,
\]
where $\Vert \cdot \Vert$ denotes the Frobenius norm, $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues. Projections in terms of other matrix norms are also possible.

To guarantee positive-definiteness, we follow \citet{huang2017calibration} and replace non-positive eigenvalues with a chosen positive value $c$ smaller than the least positive eigenvalue $\lambda_{\min}^+$, so that the corrected estimate is
\begin{equation}
  \label{near posdef}
  P_0(\bs{B})
  =
  \bs{Q}\text{diag}\{\max(\lambda_1,c),\max(\lambda_2,c),\ldots,\max(\lambda_p,c)\}\bs{Q}^\top.
\end{equation}
\citet{huang2017calibration} suggest $c_{\alpha}=10^{-\alpha}\lambda_{\min}^+$, where the parameter $\alpha$ is chosen to minimize $\Vert B - P_{c_{\alpha}}(B) \Vert + \alpha$ over a uniform partition of $\{\alpha_1,\ldots,\alpha_K\}$ of $[0,\alpha_K]$. In this paper we chose $K=20$ and $\alpha_K=10$.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "msg"
%%% End:

### 2021-08-11 Numerical_Results.tex
add files

\section{\label{numerical results}Numerical Results}
\subsection{\label{models}Models}

We considered five models for the population covariance matrix. For the first four settings, $\bs{\Sigma} = \text{diag}(\bs{s}) \bs{C} \text{diag}(\bs{s})$, where $\bs{C}$ is correlation matrix and $\bs{s}$ is a vector of standard deviations.
\begin{itemize}
\item Model 1. The standard deviations were independently generated from $\mathcal{U}(1,1.5)$ and the correlation matrix followed Model 2 of \citet{cai2011adaptive}:
  \[
  \bs{C}=
  \begin{pmatrix}&\bs{A}_1 & \bs{0} \\ &\bs{0} &\bs{I}_{p/2\times p/2}\end{pmatrix},
  \]
  where the $jk$th entry of $\bs{A}_1$ is $a_{jk} = \max(1- \vert j - k \vert / 10, 0)$. This setting modeled a sparse covariance matrix.

\item Model 2. The first $p / 2$ standard deviations equaled 1, the last $p / 2$ equaled 2, and the correlation matrix was
  \[
  \bs{C} = \begin{pmatrix} & \bs{C}_{11} & \bs{C}_{12} \\  & \bs{C}_{21} & \bs{C}_{22}\end{pmatrix},
  \]
  where $\bs{C}_{11}$ and $\bs{C}_{22}$ were $p/2 \times p/2$ compound symmetric matrices with correlation parameters 0.8 and 0.2, respectively, and $\bs{C}_{12}$ and $\bs{C}_{21}$ were $p/2 \times p/2$ matrices with entries equal to 0.4. This model was designed such that larger $\sigma_j$ and $\sigma_k$ tended to correspond to larger $r_{jk}$.

\item Model 3. The standard deviations were generated independently from $\mathcal{U}(1, 1.5)$ and $\bs{C}$ was a compound symmetric matrix with correlation parameter 0.7. This modeled a dense covariance matrix.

\item Model 4. This setting was the same as Model 3 except with correlation parameter 0.9. This high level of dependence tested the robustness of the pairwise composite likelihood estimator \eqref{Gp hat}.

\item Model 5. With $\bs{U}$ a randomly generated orthogonal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4. This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
\end{itemize}

In each scenario, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$. We generated $200$ replicates and reported average errors under the following three norms, where $\hat{\bs{\Sigma}}$ is the estimated matrix with entries $\hat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
\begin{itemize}
\item Frobenius: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\hat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$, a version of \eqref{frobenius risk},

\item Spectral: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\hat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\hat{\bs{\Sigma}} - \bs{\Sigma}$, and

\item Matrix $\ell_1$: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \hat{\sigma}_{jk} - \sigma_{jk} \vert$.
\end{itemize}

\subsection{\label{optimalK}Clustering-based exemplar algorithm}
We first studied the behavior of our $K$-means clustering-based exemplar algorithm for different $K$, described in Section \ref{implementation}. For a given $p$, we let $K = rp$ for different ratios ratios $r=2,1,0.5,0.25$. We compared these choices for $K$ to the full exemplar method. For all these estimators, we show the result after applying positive-definiteness correction.

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{img/sim1_frobenius.pdf}
  \caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.}
  \label{fig:sim1_frobenius}
\end{figure}

\begin{table}
\centering
\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{rccc}
\hline
            & p=30 & p=100 & p=200 \\
\hline
Exemplar method   & 0.0748          & 4.2575         & 47.8616       \\
$K=2p$	      & 0.0445 	     & 1.3166	   & 12.2612         \\
$K=p$            & 0.0291         & 0.8744         & 8.0862         \\
$K=p/2$         & 0.0207         & 0.5861         & 5.3757         \\
$K=p/4$      & 0.0153         &0.4132          & 3.7415         \\
\hline
\end{tabular}
\caption{\label{tab:sim1_time} Average running time for different ratios.}
\end{table}

Figure \ref{fig:sim1_frobenius} presents the Frobenius norm error estimates from Model 1 to Model 5. Table \ref{tab:sim1_time} shows the running time only for Model 1, because the running time does not vary much across different models. The results show that different $K$ exhibit similar performance and are comparable to the full exemplar method. Letting $K = p$ seemed to provide a good balance between accuracy and speed, so we implement our proposed method with $K = p$ in the rest of this paper.

\subsection{\label{compared}Methods compared}

In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness. We compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:

\begin{itemize}
\item Sample: the sample covariance matrix.

\item Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \eqref{linear model}.

\item QIS: the Quadratic-Inverse Shrinkage estimator of \citet{ledoit2019quadratic}, a recently developed nonlinear shrinkage method. QIS performs linear shrinkage on the sample eigenvalues of the covariance matrix in inverse eigenvalue space. A bandwidth parameter is required, which we choose following the paper's recommendation.

\item NERCOME: the Nonparametric Eigenvalue-Regularized COvariance Matrix Estimator of \citet{lam2016nonparametric}. This nonlinear shrinkage method randomly splits the samples into two groups, one for estimating eigenvectors and the other for estimating eigenvalues. Combining the estimates gives a matrix. Following the article, we repeated this procedure 50 times and took the final covariance matrix estimator to be the average of the individual matrices.

\item Adap: the adaptive thresholding method of \citep{cai2011adaptive} for sparse covariance matrices, which applies soft thresholding to entries of the sample covariance matrix. The threshold method is adaptive to the entry's variance and involves a tuning parameter. We fixed the parameter at 2, as recommended.
\end{itemize}

In addition to the above estimators, we also implemented the two following oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$.
\begin{itemize}
\item OracNonlin: the optimal rotation-invariant covariance estimator, defined in \citet{ledoit2019quadratic}, with $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$. The sample covariance, the linear shrinkage estimator of \citet{ledoit2004well}, and the nonlinear shrinkage estimators QIS and NERCOME are all rotation-invariant.

\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \eqref{separable}. It equals our proposed estimator \eqref{proposed} except with the true $G_p$ \eqref{bayesian} instead of $\hat{G}_p$ \eqref{Gp hat}. The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{img/sim2_frobenius.pdf}
  \caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.}
  \label{fig:sim2_frobenius}
\end{figure}

Figures \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators. Our MSG methods had the lowest or near-lowest errors across all settings except for Model 4 which has high correlations 0.9. This is not surprising because our method assumed independence of $\bs{A}_{jk}$.  In some cases, for example in Models 1 and 2, the improvement was substantial. Model 2 was especially interesting because the standard deviation and correlations were related. Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_p$ \eqref{bayesian} and leverage it to provide much more accurate estimates. The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5. In every setting, correcting MSG for positive-definiteness never increased the risk and decreased the risk in some cases. We also did experiments for Spectral norm and Matrix $\ell_1$. The results for this two norms are very similar to Frobenius norm. One exception is Adap has the lowest error in terms of Matrix $\ell_1$ norm in Model 1 because of its sparsity. Though our estimator was motivated in terms of the Frobenius norm error, it performed extremely well in terms of the other two norms as well.

%We were surprised to find that our proposed MSG methods with $d = 30$ did not perform as well as with $d = 20$, and in some cases, especially Model 4, were much worse. Intuitively, more grid points should lead to more accurate results. Indeed, we ran additional simulations using MSG with $d = 5$ and $d = 10$ and found that those results were consistent with this intuition. We believe the counterintuitive performance of $d = 30$ stems from the specific values of the specific points comprising the grid \eqref{grid} in these simulations. In a set of unreported experiments, we let $r_{jk}$ be supported on equally spaced points between the smallest and largest observed sample correlations, instead of between $[-1, 1]$ as in \eqref{grid}. In these results, $d = 20$ and $d = 30$ had similar performance, though neither performed as well as the positive-definite-corrected MSG with $d = 20$ reported here. It was computationally infeasible to implement $d$ much higher than 30. This highlights the importance of developing better computational strategies for multivariate $g$-modeling problems \citep{tao2014convex}.

Finally, the simulations show that the class of separable estimators \eqref{separable} proposed in this paper is fundamentally different from the class of rotation-invariant estimators, as the oracle optimal estimators in these two classes behave very differently. For example, the oracle separable estimator had vanishing risk in Model 2, while the oracle rotiation-invariant estimator does not. Separable estimators seemed better for Models 1 and 2 while rotation-invariant estimators were superior in Models 3 and 4. They seem comparable in Model 5.

\subsection{Data analysis}
\label{gene analysis}
Covariance matrix estimation is often used to reconstruct gene networks \citep{markowetz2007inferring}. We applied our MSG and the other covariance matrix estimators described in Section \ref{compared} to gene network estimation using data from a small round blue-cell tumor microarray experiment, which was also studied by \citet{cai2011adaptive}. \citet{osareh2009classification} report the expression of 2308 genes from 63 samples from four groups: 12 neuroblastoma, 20 rhabdomyosarcoma, 8 Burkitt lymphoma, and 23 Ewing's sarcoma patients. We followed the same data preprocessing as \citet{cai2011adaptive} and sorted the genes in decreasing order according to their $F$-statistic
\begin{equation}
F = \frac{1}{k-1}\sum_{m=1}^kn_m(\bar{x}_m - \bar{x})^2 /   \frac{1}{n-k}\sum_{m=1}^k (n_m-1)\hat{\sigma}_m^2 
\end{equation}
where $k = 4$ is the number of patient categories, $n_m$, $\bar{x}_m$, and $\hat{\sigma}_k$ represent the sample size, sample mean, and sample variance of the gene's expression in the $m$th category, respectively, and $\bar{x}$ is the global mean. We proceeded with the top 40 genes and bottom 160 genes.

We applied various methods to estimate the covariance matrix of these 200 genes. To measure the accuracy of the estimators, we split the 63 samples into two subsets $\bs{X}_1$ and $\bs{X}_2$, ensuring that each subset consisted of the same number of subjects from each of the four disease groups. After centering the variables to have zero mean, we used $\bs{X}_1$ to calculate covariance matrix estimates and compared these to the sample covariance matrix $\bs{S}_2$ of $\bs{X}_2$, which served as a proxy for the unknown true covariance matrix. We measured the errors using the Frobenius, spectral, and matrix $\ell_1$ norms. We repeated this process 200 times.

Table \ref{tab:tab1} reports the average errors across the replications. Our MSG methods had the lowest average error. The positive-definiteness correction slightly reduced the risk as well. The linear shrinkage estimate was almost as accurate, but the other methods were much less accurate. These results suggest that our estimator can perform well in realistic settings, where the mean-zero multivariate normal distributional assumption on the data may not be met.

\begin{table}
\centering
\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{rccc}
\hline
            & Frobenius & Spectral & Matrix $\ell_1$ \\
\hline
MSG   & 24.07          & 12.89         & 47.12         \\
MSGCor & \textbf{23.90} & \textbf{12.87}& \textbf{46.63}         \\
Adap        & 28.81          & 17.20         & 50.71         \\
Linear      & 24.59         & 13.57         & 51.43         \\
QIS         & 28.66          & 15.94         & 59.30         \\
NERCOME     & 29.76          & 20.13         & 57.30         \\
Sample      & 28.39          & 16.10         & 57.90         \\
\hline
\end{tabular}
\caption{\label{tab:tab1} Average gene expression covariance matrix estimation errors. Bold entries highlight the smallest errors in each column.}
\end{table}

In addition to comparing the numerical accuracies, we also investigated whether our estimator gave qualitatively different gene networks compared to the other approaches. First, Figure \ref{network} illustrates the covariance matrices in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects. To avoid completely connected graphs, we sparsified the matrix estimates by thresholding the smaller entries of each matrix to zero. Since the adaptive thresholding method of \citet{cai2011adaptive} naturally produced a sparse estimated matrix, we thresholded the other matrix estimates to match the sparsity level of the \citet{cai2011adaptive} estimate.

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{img/network.pdf}
  \caption{Gene networks recovered by the different covariance matrix estimation methods.}
  \label{network}
\end{figure}

The results show several interesting features. First, there appear to be two major clusters, which are disconnected in every estimated network except for the one produced by the adaptive thresholding approach. Second, the larger cluster appears to contain two sub-clusters, and this finer structure was only recovered by MSG and QIS, and to a lesser extent the linear shrinkage estimator and NERCOME. Finally, the nodes in the networks estimated by QIS and NERCOME appear to be clustered more tightly together compared to in the other networks. These observations suggest that MSG produces qualitatively different networks, in addition to lower estimation errors.

Finally, we also compared the estimated degrees of the genes in the different networks. For each estimated network, we ordered the 200 genes by degree and then selected the top 20\%, denoting this set $J_k$ for the $k$th network. For each pair of networks $k$ and $k'$, we calculated the similarity between their most connected genes using Jaccard index $\vert J_k \cap J_{k'} \vert / \vert J_k \cup J_{k'} \vert$. Figure \ref{top20} visualizes these similarities. Interestingly, however, among all estimators, they were also the most similar to the unbiased sample covariance matrix. Together with the above results, this indicates that MSG may simultaneously give the lowest error and, at least in terms of degree estimation, the most unbiased results.

\begin{figure}
  \centering
  \includegraphics[width=0.85\textwidth]{img/top20.pdf}
  \caption{Similarities of gene degrees between the estimated networks. Each number reports the Jaccard index between the top 20\% most connected genes of each pair of networks.}
  \label{top20}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "msg"
%%% End:

### 2021-08-11 msg.tex
add files

% \documentclass[12pt]{article}
\documentclass[useAMS,referee,usenatbib]{biom}
%\usepackage[utf8]{inputenc}
%\usepackage{amsmath, authblk}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape} 
%\usepackage{graphics}
%\usepackage{afterpage}
%\usepackage{xcolor}
%\usepackage{citeppp}
%\usepackage[usenatbib]{natbib}
%\usepackage[round]{natbib}
%\usepackage{url} 
\usepackage{graphicx}
%\usepackage{subcaption}

%\usepackage[margin = 1in]{geometry}

\usepackage[figuresright]{rotating}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
%\def\bs{\boldsymbol}
\def\bs{\bmath}
\def\bb{\mathbb}

\newtheorem{prop}{Proposition}
%\newtheorem{lemma}{Lemma}

\title[Empirical Bayes Covariance Matrix Estimation]{A Nonparametric Empirical Bayes Approach to Covariance Matrix Estimation}

\author{Huiqin Xin$^{*}$\email{huiqinx2@illinois.edu } and
Sihai Dave Zhao$^{**}$\email{sdzhao@illinois.edu} \\
Department of Statistics, University of Illinois at Urbana-Champaign, Champaign, Illinois}
%
%\author{John Author$^{*}$\email{author@address.edu} and
%Kathy Authoress$^{**}$\email{email2@address.edu} \\
%Department of Statistics, University of Warwick, Coventry CV4 7AL, U.K.}

\begin{document}

%\date{{\it Received October} 2007. {\it Revised February} 2008.  {\it
%Accepted March} 2008.}

\label{firstpage}
\begin{abstract}
  We propose an empirical Bayes method to estimate high-dimensional covariance matrices. Our procedure centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem. Drawing from the compound decision theory literature, we introduce a new class of decision rules that generalizes several existing procedures. We then use a nonparametric empirical Bayes $g$-modeling approach to estimate the oracle optimal rule in that class. This allows us to let the data itself determine how best to shrink the estimator, rather than shrinking in a pre-determined direction such as toward a diagonal matrix. Simulation results and a gene expression network analysis shows that our approach can outperform a number of state-of-the-art proposals in a wide range of settings, sometimes substantially.
\end{abstract}

\begin{keywords}
Compound decision theory; $g$-modeling; nonparametric maximum likelihood; separable decision rule.
\end{keywords}

\maketitle

\section{Introduction}
\label{introduction}

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications, such as portfolio management \citep{ledoit2003improved}, genomics \citep{schafer2005shrinkage}, and array signal processing \citep{abramovich2001locally}. However, in modern problems the number of features can be of the same order as or exceed the sample size, and the standard sample covariance matrix estimator behaves poorly in this regime. To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrix. These can roughly be divided into two groups, according to whether they impose assumptions about the structure of population covariance matrix.

Structured methods make structural assumptions about the population covariance matrix. One class models the population covariance matrix as sparse. The most common method to address this problem is thresholding \citep{rothman2009generalized, cai2011adaptive}. Penalized likelihood methods \citep{xue2012positive} can also estimate large-scale sparse covariance matrix by penalizing a log-likelihood function. Another class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension. Other common structured methods assume that the covariance matrix is banding \citep{li2017estimation} or Toeplitz matrix \citep{liu2017covariance}.

In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet can still outperform the sample covariance matrix. A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix. More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}. These shrink the eigenvalues of the sample covariance matrix toward clusters. Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global mean.

Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance. However, they modify only the sample eigenvalues and not the sample eigenvectors. It is known that sample eigenvectors are not consistent estimators of population eigenvectors when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}. This suggests that there may exist a class of unstructured estimators that can outperform nonlinear shrinkage.

%In this paper, we propose a new covariance matrix estimator which is approximately optimal among all separable estimators. Our method is motivated by the similarity between covariance estimation and vector mean estimation. Covariance matrix estimation under Frobenius loss can be viewed as a vector estimation problem by converting the matrix to a vector. \citep{cai_liu} used this motivation to generalize thresholding method in mean vector estimation to large sparse covariance matrix estimation. Nevertheless, they are not totally the same because covariance matrix has repeated entries due to its symmetry. In mean vector estimation problem, optimality has been achieved among all separable decision rules. \citep{jiang_zhang2009} and \citep{brown_greenshtein} proposed empirical Bayes methods to approximate the optimal separable estimator. We proposed to apply vector estimation methods to covariance estimation problem and expect that these methods will also have good properties here. Result in simulation and real data set shows that our estimator has good performance in different scenarios.

Here we propose a new unstructured estimator for high-dimensional covariance matrices. Our approach centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem. We do this because it allows us to use a nonparametric empirical Bayes shrinkage procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex}. We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator. Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis.

The article is organized as follows. In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach. In Section \ref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset. Finally, Section \ref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub.

\section{\label{sec:method}Method}
\subsection{\label{sec:compound}Compound decision problem formulation}
Suppose we have $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$. The purpose of this paper is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$ that minimizes the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.

Our proposed approach is motivated by two observations. First, \ref{frobenius risk} shows that estimating $\bs{\Sigma}$ under Frobenius risk is equivalent to simultaneously estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under a loss function that aggregates errors across components. Second, this type of vector estimation problem has been well-studied in the compound decision literature. Thus, recent advances in vector estimation may be profitably applied to covariance matrices.

We first briefly review compound decision problems. Introduced by \citet{robbins1951asymptotically}, these problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk
\begin{equation}
\label{compound risk}
R(\bs{\theta},\bs{\delta}) = \frac{1}{n}\sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))
\end{equation}
where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is squared error loss \citep{johnstone2017gaussian}.

A key property of compound decision problems is that while a given $Y_i$ seems to offer no information about any specific $\theta_j$ when $j \ne i$, borrowing information across all components of $\bs{Y}$ to estimate $\bs{\theta}$ is superior to estimating each $\theta_i$ using the corresponding $Y_i$ alone. A classical example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. It is well-known that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$. A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.

We propose to apply some of these recent vector estimation ideas to covariance matrix estimation. Clearly, covariance matrix estimation under the Frobenius risk \ref{frobenius risk} can be viewed as a compound decision problem. Furthermore, some existing covariance matrix estimation procedures can already be interpreted as taking a vector approach. The sample covariance matrix $\bs{S}$, for instance, can be thought of as estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$ using maximum likelihood. Less trivially, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective. Their adaptive thresholding method is a version of the soft thresholding method of \citet{donoho1995adapting}, which was originally developed to estimate a sparse mean vector in the Gaussian sequence problem.

Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be interpreted as a vector estimator. The estimator is defined as
\begin{equation}
\label{linear model}
\hat{\bs{\Sigma}}_{\text{LW}} = \hat{\rho}_{\text{LW}} \bs{S} + (1-\hat{\rho}_{\text{LW}})\hat{\mu} \bs{I}
\end{equation} 
where $\hat{\mu} = \text{tr}(\bs{S})/p$, and $\hat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\hat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$. Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \ref{frobenius risk}. We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
\begin{equation}
  \label{linear class}
  \beta_S s_{jk} + \beta_I u_{jk},
\end{equation}
where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. It is straightforward to show that
$$\hat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
is an unbiased estimate of the risk \ref{frobenius risk}, where $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n$ is the sample variance of $s_{jk}$. The optimal estimator in this class can now be chosen by minimizing $\hat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}. Proposition \ref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \ref{linear model}.
\begin{prop}
  \label{prop:linear}
  Define the estimator $\hat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\hat{\sigma}_{jk} = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, where $(\hat{\beta}_S, \hat{\beta}_I) = \argmin_{\beta_S, \beta_I} \hat{R}(\beta_S, \beta_I)$. Then $\hat{\bs{\Sigma}}_V$ = $\hat{\bs{\Sigma}}_{LW}$.
\end{prop}

\subsection{\label{sec:proposed}Proposed estimator}

The previous section argues that treating covariance matrix estimation as a vector estimation problem can be a fruitful strategy, but discusses only estimators linear in $s_{jk}$. We propose to consider a larger class, the class of so-called separable rules. In the standard compound decision problem of estimating $\bs{\theta}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ \citep{robbins1951asymptotically}. Here we generalize this to the problem of estimating a vectorized matrix. For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, define the class of separable rules
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} : \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})\},
\end{equation}
where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature. In other words, rules in $\mathcal{S}$ estimate the $jk$th entry of the covariance matrix using a fixed function $t$ of only observations from features $j$ and $k$.

We propose to search for the optimal estimator within $\mathcal{S}$. This is sensible because $\mathcal{S}$ includes the sample covariance $(s_{11}, \ldots, s_{pp})^\top$, the class of linear estimators \ref{linear class} used by \citet{ledoit2004well}, which can be expressed as
\[
\delta_{jk}(\bs{X}) = \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n + \beta_I I(\bs{X}_{\cdot j} = \bs{X}_{\cdot k}),
\]
and the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}. Therefore the optimal separable estimator $\bs{\delta}^\star$ that minimizes the scaled squared Frobenius risk \ref{frobenius risk} will perform at least as well as these three estimators, and may perform better. Targeting the optimal separable rule is standard in the compound decision literature \citep{zhang2003compound}.

The optimal $\bs{\delta}^\star$ is an oracle estimator and cannot be calculated in practice, as the true risk is unknown. In the classical compound decision framework, empirical Bayes methods are used to estimate the oracle optimal separable rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}. We take a similar approach here. To simplify notation, denote $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})^\top$ as $\bs{A}_{jk}$. The density of $f(\cdot \mid \bs{\eta}_{jk})$ of $\bs{A}_{jk}$ depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation. When $r_{jk} \ne 1$, $\bs{A}_{jk}$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
\[
\bs{C}_{jk}=\begin{bmatrix}
&\sigma_j^2 & \sigma_j\sigma_k r_{jk} \\
& \sigma_j\sigma_k r_{jk}  & \sigma_k^2
\end{bmatrix}.
\]
When $r_{jk} = 1$, $\bs{A}_{jk}$ consists of mean-zero univariate normals with variances $\sigma_j^2$.

Now consider the Bayesian model
\begin{equation}
  \label{bayesian}
  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
By the fundamental theorem of compound decisions \citep{robbins1951asymptotically, jiang2009general}, this is closely related to the vectorized covariance matrix estimation problem under Frobenius risk \ref{frobenius risk}: for any separable $\bs{\delta} \in \mathcal{S}$ \ref{separable}, the Frobenius risk can be written as
\begin{align*}
  R(\bs{\Sigma}, \bs{\delta})
  =\,&
  \frac{1}{p^2} \sum_{j,k = 1}^p
  \int \{t(\bs{A}) - \sigma_{jk} \}^2 f(\bs{A} \mid \bs{\eta}_{jk}) d\bs{A}\\
  =\,&
  \int \int \{t(\bs{A}) - g(\bs{\eta})\}^2 f(\bs{A} \mid \bs{\eta}) dG_p(\bs{\eta}) d\bs{A}
  =
  \bb{E}[\{t(\bs{A}) -  g(\bs{\eta})\}^2],
\end{align*}
where $g(a, b, \gamma) = a b \gamma$ and the final expectation is the Bayes risk of estimating $\sigma_{jk}$. The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star(\bs{A}_{jk})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid \bs{A}\}$ minimizes the Bayes risk.

Based on this result, we propose the following empirical Bayes procedure. We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the prior $G_p$. Under the Bayesian model \ref{bayesian}, and the working assumption that the $\bs{A}_{jk}$ are independent across $jk$, we estimate $G_p$ using
\begin{equation}
  \label{Gp hat}
  \hat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
\end{equation}
where $\mathcal{G}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$. Of course, the $\bs{A}_{jk}$ are not independent, so $\hat{G}_p$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}. Using $\hat{G}_p$, we estimate the vectorized $\bs{\Sigma}$ using
\begin{equation}
  \label{proposed}
  \hat{\bs{\delta}}(\bs{X})
  =
  (\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
  \quad
  \hat{t}(\bs{A}) = \frac{\int g(\bs{\eta}) f(\bs{A} \mid \bs{\eta})d\hat{G}_p(\bs{\eta})}{\int f(\bs{A} \mid \bs{\eta}) d\hat{G}_p(\bs{\eta})}.
\end{equation}
The $\hat{t}$ estimates the Bayes rule $t^\star$ and $\hat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.

Our proposed procedure is an example of what \citet{efron2014two} calls $g$-modeling, an approach to empirical Bayes problems that proceeds by modeling the prior. A major advantage of nonparametric estiation of the prior is that it allows the data itself to determine how best to shrink the estimator. In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}. Theoretical justification of our proposed $\hat{\bs{\delta}}$ is difficult and is discussed in Section \ref{discussion}. Nevertheless, our numerical results in Section \ref{numerical results} show that in practice, our $\hat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.

\subsection{\label{implementation}Implementation}

Calculating the estimated prior $\hat{G}_p$ \ref{Gp hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$. \citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points. The EM algorithm has traditionally been used to estimate the locations of the support points and the masses at those points \citep{laird1978nonparametric}, but this is a difficult nonconvex optimization problem.

Instead, we maximize the pairwise composite likelihood over a fixed grid of support points, similar to recent $g$-modeling procedures for standard compound decision problems; this restores convexity \citep{jiang2009general, koenker2014convex, feng2018approximate}. Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$. We can then use the EM algorithm to estimate the masses $\hat{\bs{w}}=\{\hat{w}_{1},\ldots, \hat{w}_{D}\}$ at those points via the iteration
\[
\hat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_l)}
\]
over $k$. Early stopping of the EM algorithm can be useful \citep{koenker2019comment}, and more sophisticated convex optimization procedures can be used as well \citep{koenker2014convex}. Our proposed estimator \ref{proposed} then becomes
\[
\hat{\bs{\delta}}(\bs{X})
=
(\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
\quad
\hat{t}(\bs{A}) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}.
\]

Ideally, the grid points should be chosen to densely cover the parameter space. However, the fact that $G_p$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$. Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\hat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$. This reduces the size of the support set, but even in this case the computation complexity grows like $O(p^2)$.

Here we propose a clustering-based exemplar algorithm to further improve computational efficiency. Let $s_j$, $s_k$, and $\gamma_{jk}$ be the sample variances and correlation between the $j$th and $k$th covariates. We first apply $K$-means clustering to identify $K$ clusters among the $p (p-1) / 2$ off-diagonal sample points $(s_j,s_k,\gamma_{jk})$ and $\lceil{K^{1/2}}\rceil$ clusters among the $p$ diagonal sample points $(s_j,s_j,1)$. We then use the $K+\lceil{K^{1/2}}\rceil$ cluster centroids as our support points. We cluster the off- and on-diagonal observations separately to ensure that the support points $\bs{\xi}_\tau$ are such that $f(\bs{A}_{jk} \mid \bs{\xi}_\tau) \ne 0$ when both $j = k$ and $j \ne k$. Figure \ref{fig:sim1_frobenius} shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table \ref{tab:sim1_time} shows that they can be significantly faster.

\subsection{\label{posdef}Positive definiteness correction}
Our proposed estimator \ref{proposed} is not guaranteed to be positive-definite. To correct this, we reshape our vector estimator back into a matrix and then identify the closest positive-definite matrix. \citet{higham1988computing} and \citet{huang2017calibration} showed that the projection of a $p \times p$ symmetric matrix $\bs{B}$ onto the space of positive semi-definite matrices is
\[
P_0(\bs{B})
=
\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert
=
\bs{Q}\text{diag}\{\max(\lambda_1,0),\max(\lambda_2,0),\ldots,\max(\lambda_p,0)\}\bs{Q}^\top,
\]
where $\Vert \cdot \Vert$ denotes the Frobenius norm, $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues. Projections in terms of other matrix norms are also possible.

To guarantee positive-definiteness, we follow \citet{huang2017calibration} and replace non-positive eigenvalues with a chosen positive value $c$ smaller than the least positive eigenvalue $\lambda_{\min}^+$, so that the corrected estimate is
\begin{equation}
  \label{near posdef}
  P_0(\bs{B})
  =
  \bs{Q}\text{diag}\{\max(\lambda_1,c),\max(\lambda_2,c),\ldots,\max(\lambda_p,c)\}\bs{Q}^\top.
\end{equation}
\citet{huang2017calibration} suggest $c_{\alpha}=10^{-\alpha}\lambda_{\min}^+$, where the parameter $\alpha$ is chosen to minimize $\Vert B - P_{c_{\alpha}}(B) \Vert + \alpha$ over a uniform partition of $\{\alpha_1,\ldots,\alpha_K\}$ of $[0,\alpha_K]$. In this paper we chose $K=20$ and $\alpha_K=10$.

\section{\label{numerical results}Numerical Results}
\subsection{\label{models}Models}

We considered five models for the population covariance matrix. For the first four settings, $\bs{\Sigma} = \text{diag}(\bs{s}) \bs{C} \text{diag}(\bs{s})$, where $\bs{C}$ is correlation matrix and $\bs{s}$ is a vector of standard deviations.
\begin{itemize}
\item Model 1. The standard deviations were independently generated from $\mathcal{U}(1,1.5)$ and the correlation matrix followed Model 2 of \citet{cai2011adaptive}:
  \[
  \bs{C}=
  \begin{pmatrix}&\bs{A}_1 & \bs{0} \\ &\bs{0} &\bs{I}_{p/2\times p/2}\end{pmatrix},
  \]
  where the $jk$th entry of $\bs{A}_1$ is $a_{jk} = \max(1- \vert j - k \vert / 10, 0)$. This setting modeled a sparse covariance matrix.

\item Model 2. The first $p / 2$ standard deviations equaled 1, the last $p / 2$ equaled 2, and the correlation matrix was
  \[
  \bs{C} = \begin{pmatrix} & \bs{C}_{11} & \bs{C}_{12} \\  & \bs{C}_{21} & \bs{C}_{22}\end{pmatrix},
  \]
  where $\bs{C}_{11}$ and $\bs{C}_{22}$ were $p/2 \times p/2$ compound symmetric matrices with correlation parameters 0.8 and 0.2, respectively, and $\bs{C}_{12}$ and $\bs{C}_{21}$ were $p/2 \times p/2$ matrices with entries equal to 0.4. This model was designed such that larger $\sigma_j$ and $\sigma_k$ tended to correspond to larger $r_{jk}$.

\item Model 3. The standard deviations were generated independently from $\mathcal{U}(1, 1.5)$ and $\bs{C}$ was a compound symmetric matrix with correlation parameter 0.7. This modeled a dense covariance matrix.

\item Model 4. This setting was the same as Model 3 except with correlation parameter 0.9. This high level of dependence tested the robustness of the pairwise composite likelihood estimator \ref{Gp hat}.

\item Model 5. With $\bs{U}$ a randomly generated orthogonal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4. This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
\end{itemize}

In each scenario, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$. We generated $200$ replicates and reported average errors under the following three norms, where $\hat{\bs{\Sigma}}$ is the estimated matrix with entries $\hat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
\begin{itemize}
\item Frobenius: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\hat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$, a version of \ref{frobenius risk},

\item Spectral: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\hat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\hat{\bs{\Sigma}} - \bs{\Sigma}$, and

\item Matrix $\ell_1$: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \hat{\sigma}_{jk} - \sigma_{jk} \vert$.
\end{itemize}

\subsection{\label{optimalK}Clustering-based exemplar algorithm}
We first studied the behavior of our $K$-means clustering-based exemplar algorithm for different $K$, described in Section \ref{implementation}. For a given $p$, we let $K = rp$ for different ratios ratios $r=2,1,0.5,0.25$. We compared these choices for $K$ to the full exemplar method. For all these estimators, we show the result after applying positive-definiteness correction.

\begin{figure}
\begin{center}
\centerline{\includegraphics[width=0.85\textwidth]{img/sim1_frobenius.pdf} }
\end{center}
\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.}
\label{fig:sim1_frobenius}
\end{figure}

\begin{table}
\begin{center}
\caption{\label{tab:sim1_time} Average running time for different ratios.}
%\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{lrrr}
\Hline
            & p=30 & p=100 & p=200 \\
\hline
Exemplar method   & 0.0748          & 4.2575         & 47.8616       \\
$K=2p$	      & 0.0445 	     & 1.3166	   & 12.2612         \\
$K=p$            & 0.0291         & 0.8744         & 8.0862         \\
$K=p/2$         & 0.0207         & 0.5861         & 5.3757         \\
$K=p/4$      & 0.0153         &0.4132          & 3.7415         \\
\hline
\end{tabular}
\end{center}
\end{table}

Figure \ref{fig:sim1_frobenius} presents the Frobenius norm error estimates from Model 1 to Model 5. Table \ref{tab:sim1_time} shows the running time only for Model 1, because the running time does not vary much across different models. The results show that different $K$ exhibit similar performance and are comparable to the full exemplar method. Letting $K = p$ seemed to provide a good balance between accuracy and speed, so we implement our proposed method with $K = p$ in the rest of this paper.

\subsection{\label{compared}Methods compared}

In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness. We compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:

\begin{itemize}
\item Sample: the sample covariance matrix.

\item Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \ref{linear model}.

\item QIS: the Quadratic-Inverse Shrinkage estimator of \citet{ledoit2019quadratic}, a recently developed nonlinear shrinkage method. QIS performs linear shrinkage on the sample eigenvalues of the covariance matrix in inverse eigenvalue space. A bandwidth parameter is required, which we choose following the paper's recommendation.

\item NERCOME: the Nonparametric Eigenvalue-Regularized COvariance Matrix Estimator of \citet{lam2016nonparametric}. This nonlinear shrinkage method randomly splits the samples into two groups, one for estimating eigenvectors and the other for estimating eigenvalues. Combining the estimates gives a matrix. Following the article, we repeated this procedure 50 times and took the final covariance matrix estimator to be the average of the individual matrices.

\item Adap: the adaptive thresholding method of \citep{cai2011adaptive} for sparse covariance matrices, which applies soft thresholding to entries of the sample covariance matrix. The threshold method is adaptive to the entry's variance and involves a tuning parameter. We fixed the parameter at 2, as recommended.
\end{itemize}

In addition to the above estimators, we also implemented the two following oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$.
\begin{itemize}
\item OracNonlin: the optimal rotation-invariant covariance estimator, defined in \citet{ledoit2019quadratic}, with $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$. The sample covariance, the linear shrinkage estimator of \citet{ledoit2004well}, and the nonlinear shrinkage estimators QIS and NERCOME are all rotation-invariant.

\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \ref{separable}. It equals our proposed estimator \ref{proposed} except with the true $G_p$ \ref{bayesian} instead of $\hat{G}_p$ \ref{Gp hat}. The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
\end{itemize}

\begin{figure}
\begin{center}
\centerline{  \includegraphics[width=0.85\textwidth]{img/sim2_frobenius.pdf}}
\end{center}
\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.}
\label{fig:sim2_frobenius}
\end{figure}

Figures \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators. Our MSG methods had the lowest or near-lowest errors across all settings except for Model 4 which has high correlations 0.9. This is not surprising because our method assumed independence of $\bs{A}_{jk}$.  In some cases, for example in Models 1 and 2, the improvement was substantial. Model 2 was especially interesting because the standard deviation and correlations were related. Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_p$ \ref{bayesian} and leverage it to provide much more accurate estimates. The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5. In every setting, correcting MSG for positive-definiteness never increased the risk and decreased the risk in some cases. We also did experiments for Spectral norm and Matrix $\ell_1$. The results for this two norms are very similar to Frobenius norm. One exception is Adap has the lowest error in terms of Matrix $\ell_1$ norm in Model 1 because of its sparsity. Though our estimator was motivated in terms of the Frobenius norm error, it performed extremely well in terms of the other two norms as well.

Finally, the simulations show that the class of separable estimators \ref{separable} proposed in this paper is fundamentally different from the class of rotation-invariant estimators, as the oracle optimal estimators in these two classes behave very differently. For example, the oracle separable estimator had vanishing risk in Model 2, while the oracle rotiation-invariant estimator does not. Separable estimators seemed better for Models 1 and 2 while rotation-invariant estimators were superior in Models 3 and 4. They seem comparable in Model 5.

\subsection{Data analysis}
\label{gene analysis}
Covariance matrix estimation is often used to reconstruct gene networks \citep{markowetz2007inferring}. We applied our MSG and the other covariance matrix estimators described in Section \ref{compared} to gene network estimation using data from a small round blue-cell tumor microarray experiment \citep{khan2001classification}, which was also studied by \citet{cai2011adaptive}. \citet{osareh2009classification} report the expression of 2308 genes from 63 samples from four groups: 12 neuroblastoma, 20 rhabdomyosarcoma, 8 Burkitt lymphoma, and 23 Ewing's sarcoma patients. We followed the same data preprocessing as \citet{cai2011adaptive} and sorted the genes in decreasing order according to their $F$-statistic
\begin{equation}
F = \frac{1}{k-1}\sum_{m=1}^kn_m(\bar{x}_m - \bar{x})^2 /   \frac{1}{n-k}\sum_{m=1}^k (n_m-1)\hat{\sigma}_m^2 
\end{equation}
where $k = 4$ is the number of patient categories, $n_m$, $\bar{x}_m$, and $\hat{\sigma}_k$ represent the sample size, sample mean, and sample variance of the gene's expression in the $m$th category, respectively, and $\bar{x}$ is the global mean. We proceeded with the top 40 genes and bottom 160 genes.

We applied various methods to estimate the covariance matrix of these 200 genes. To measure the accuracy of the estimators, we split the 63 samples into two subsets $\bs{X}_1$ and $\bs{X}_2$, ensuring that each subset consisted of the same number of subjects from each of the four disease groups. After centering the variables to have zero mean, we used $\bs{X}_1$ to calculate covariance matrix estimates and compared these to the sample covariance matrix $\bs{S}_2$ of $\bs{X}_2$, which served as a proxy for the unknown true covariance matrix. We measured the errors using the Frobenius, spectral, and matrix $\ell_1$ norms. We repeated this process 200 times.

Table \ref{tab:tab1} reports the average errors across the replications. Our MSG methods had the lowest average error. The positive-definiteness correction slightly reduced the risk as well. The linear shrinkage estimate was almost as accurate, but the other methods were much less accurate. These results suggest that our estimator can perform well in realistic settings, where the mean-zero multivariate normal distributional assumption on the data may not be met.

\begin{table}
\begin{center}
\caption{\label{tab:tab1} Average gene expression covariance matrix estimation errors. Bold entries highlight the smallest errors in each column.}
%\setlength{\tabcolsep}{3pt} % Default value: 6pt
\begin{tabular}{lrrr}
\Hline
            & Frobenius & Spectral & Matrix $\ell_1$ \\
\hline
MSG   & 24.07          & 12.89         & 47.12         \\
MSGCor & \textbf{23.90} & \textbf{12.87}& \textbf{46.63}         \\
Adap        & 28.81          & 17.20         & 50.71         \\
Linear      & 24.59         & 13.57         & 51.43         \\
QIS         & 28.66          & 15.94         & 59.30         \\
NERCOME     & 29.76          & 20.13         & 57.30         \\
Sample      & 28.39          & 16.10         & 57.90         \\
\hline

\end{tabular}
\end{center}
\end{table}

In addition to comparing the numerical accuracies, we also investigated whether our estimator gave qualitatively different gene networks compared to the other approaches. First, Figure \ref{network} illustrates the covariance matrices in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects. To avoid completely connected graphs, we sparsified the matrix estimates by thresholding the smaller entries of each matrix to zero. Since the adaptive thresholding method of \citet{cai2011adaptive} naturally produced a sparse estimated matrix, we thresholded the other matrix estimates to match the sparsity level of the \citet{cai2011adaptive} estimate.

\begin{figure}
\begin{center}
\centerline{ \includegraphics[width=0.85\textwidth]{img/network.pdf}}
\end{center}
\caption{Gene networks recovered by the different covariance matrix estimation methods.}
\label{network}
\end{figure}

The results show several interesting features. First, there appear to be two major clusters, which are disconnected in every estimated network except for the one produced by the adaptive thresholding approach. Second, the larger cluster appears to contain two sub-clusters, and this finer structure was only recovered by MSG and QIS, and to a lesser extent the linear shrinkage estimator and NERCOME. Finally, the nodes in the networks estimated by QIS and NERCOME appear to be clustered more tightly together compared to in the other networks. These observations suggest that MSG produces qualitatively different networks, in addition to lower estimation errors.

Finally, we also compared the estimated degrees of the genes in the different networks. For each estimated network, we ordered the 200 genes by degree and then selected the top 20\%, denoting this set $J_k$ for the $k$th network. For each pair of networks $k$ and $k'$, we calculated the similarity between their most connected genes using Jaccard index $\vert J_k \cap J_{k'} \vert / \vert J_k \cup J_{k'} \vert$. Figure \ref{top20} visualizes these similarities. Interestingly, however, among all estimators, they were also the most similar to the unbiased sample covariance matrix. Together with the above results, this indicates that MSG may simultaneously give the lowest error and, at least in terms of degree estimation, the most unbiased results.

\begin{figure}
\begin{center}
\centerline{\includegraphics[width=0.85\textwidth]{img/top20.pdf}}
\end{center}
\caption{Similarities of gene degrees between the estimated networks. Each number reports the Jaccard index between the top 20\% most connected genes of each pair of networks.}
\label{top20}
\end{figure}

\section{\label{sec:discussion}Discussion}
\label{discussion}

The class of separable covariance matrix estimators \ref{separable} that we proposed in this paper appears to be very promising. Many existing procedures already explicitly or implicitly target this class, and our proposed estimate \ref{proposed} of the optimal separable estimator outperforms a number of existing covariance matrix estimators. This is surprising because our approach vectorizes the matrix and therefore cannot take matrix structure, such as positive-definiteness, into account. This suggests that a vectorized approach combined with a positive-definiteness constraint may have improved performance. The resulting estimator would necessarily not be separable, because the estimate of the $jk$th entry would depend on more than just the $j$th and $k$th observed features, so the $g$-modeling estimation strategy is insufficient. More work is needed.

Though our estimator performs well in simulations and in real data, providing theoretical guarantees is difficult. In the standard mean vector estimation problem with $Y_i \sim N(\theta_i, 1)$, \citet{jiang2009general} showed that an empirical Bayes estimator based on a nonparametric maximum likelihood estimate of the prior on the $\theta_i$ can indeed asymptotically achieve the same risk as the oracle optimal separable estimator. However, this was in a simple model with a univariate prior distribution. \citet{saha2020nonparametric} extended these results to multivariate $\bs{Y}_i \sim N(\bs{\theta}_i, \bs{\Sigma}_i)$ with a multivariate prior on the $\bs{\theta}_i$, but assumed that the $\bs{Y}_i$ were independent. In contrast, our covariance matrix estimator is built from arbitrarily dependent $\bs{A}_{jk}$. These imposes significant theoretical difficulties that will require substantial work to address; we leave this for future research.

Finally, we have so far assumed that our data multivariate normal. To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \ref{Gp hat} and in our proposed estimator \ref{proposed}. If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.

\backmatter

\section*{Acknowledgments}
We thank Dr. Roger Koenker for his valuable comments.\vspace*{-8pt}

\bibliographystyle{biom} 
\bibliography{biblio}

\appendix
\section{}
\subsection{Proof of Proposition \ref{prop:linear}}

\begin{proof}
  We first rewrite the risk estimate $\hat{R}(\beta_S, \beta_I)$. Define $\bs{M}=(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$. Then the unbiased risk estimator can be re-written as
  \[
  \hat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M},
  \]
  where $\bs{Z} =(\bs{v_S}, \bs{v_I})$. Therefore
  \begin{align*}
    \hat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }(\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}\bs{v_S}-\bs{M}),\\
    &
    \hat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\hat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
  \end{align*}

  We will need to show $\hat{\mu} b_n^2 / d_n^2 = \hat{\beta}_I$ and $\hat{\beta}_I/\hat{\mu} + \hat{\beta}_S = 1$. Since
$$\bs{Z}^\top \bs{Z} = \begin{pmatrix}
s_{11} &\ldots &s_{pp} \\
u_{11} &\ldots &u_{pp} 
\end{pmatrix}\begin{pmatrix}
s_{11}  &u_{11} \\
\ldots&\ldots\\
s_{pp} &u_{pp} 
\end{pmatrix}=\begin{pmatrix}
\sum_{j,k=1}^ps_{jk}^2 & \sum_{j=1}^ps_{jj}\\
\sum_{j=1}^ps_{jj} & p
\end{pmatrix}$$
and
$\text{det}( \bs{Z}^\top \bs{Z} ) = p\sum_{j,k=1}^ps_{jk}^2 - (\sum_{j=1}^p s_{jj})^2 = p^3d_n^2$, it follows that
$$(\bs{Z}^\top \bs{Z})^{-1} = \frac{1}{p^3d_n^2}\begin{pmatrix}
p & -\sum_{j=1}^p s_{jj}\\
-\sum_{j=1}^p s_{jj} & \sum_{j,k=1}^ps_{jk}^2
\end{pmatrix},$$
and in addition
$$\bs{Z}^\top \bs{S} = \begin{pmatrix}
\sum_{j,k=1}^ps_{jk}^2\\
\sum_{j=1}^p s_{jj}
\end{pmatrix},
\quad 
\bs{Z^\top S}-\bs{M} =  \begin{pmatrix}
&\sum_{j,k=1}^ps_{jk}^2 - \hat{\Delta}_{jk}^2\\
&\sum_{j=1}^p s_{jj}
\end{pmatrix}.$$

Therefore
\begin{align*}
\hat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1} \bs{Z^\top S} \\
&= \frac{1}{p^3d_n^2}\begin{pmatrix}
p & -\sum_{j=1}^p s_{jj}\\
-\sum_{j=1}^p s_{jj} & \sum_{j,k=1}^ps_{jk}^2
\end{pmatrix}\begin{pmatrix}
\sum_{j,k=1}^ps_{jk}^2 - \hat{\Delta}_{jk}^2\\
\sum_{j=1}^p s_{jj}
\end{pmatrix}\\
&= \frac{1}{p^3d_n^2}\begin{pmatrix}
 p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
 (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)
\end{pmatrix}.
\end{align*}
The second component of $\hat{\bs{\beta}}$ equals $\hat{\beta}_I$, so
\begin{align*}
\hat{\beta}_I &= \frac{1}{p^3d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)\\
&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)/p^2 \} / d_n^2
=\hat{\mu}\frac{b_n^2}{d_n^2}.
\end{align*}
Furthermore,
\begin{align*}
\hat{\beta}_I /\hat{\mu} + \hat{\beta}_S &= \frac{1}{p^3d_n^2}p\sum_{j,k=1}^p \{s_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2\}\\
&=\frac{1}{p^3d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}
=
1.
\end{align*}
\end{proof}

\label{lastpage}
\end{document}

### 2021-08-24 msg.tex
changes of ref2 and part of ref1

@@
 
-\title[Empirical Bayes Covariance Matrix Estimation]{A Nonparametric Empirical Bayes Approach to Covariance Matrix Estimation}
+\title[Empirical Bayes Covariance Matrix Estimation]{A Compound Decision Approach to Covariance Matrix Estimation}
 
@@
 \label{linear model}
-\hat{\bs{\Sigma}}_{\text{LW}} = \hat{\rho}_{\text{LW}} \bs{S} + (1-\hat{\rho}_{\text{LW}})\hat{\mu} \bs{I}
+\widehat{\bs{\Sigma}}_{\text{LW}} = \widehat{\rho}_{\text{LW}} \bs{S} + (1-\widehat{\rho}_{\text{LW}})\widehat{\mu} \bs{I}
 \end{equation} 
-where $\hat{\mu} = \text{tr}(\bs{S})/p$, and $\hat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\hat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$.
+where $\widehat{\mu} = \text{tr}(\bs{S})/p$, and $\widehat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\widehat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$.
 Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \ref{frobenius risk}.
@@
 where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. It is straightforward to show that
-$$\hat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
-is an unbiased estimate of the risk \ref{frobenius risk}, where $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n$ is the sample variance of $s_{jk}$.
-The optimal estimator in this class can now be chosen by minimizing $\hat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
+$$\widehat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
+is an unbiased estimate of the risk \ref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n$ is the sample variance of $s_{jk}$.
+The optimal estimator in this class can now be chosen by minimizing $\widehat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
 It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}.
@@
   \label{prop:linear}
-  Define the estimator $\hat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\hat{\sigma}_{jk} = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, where $(\hat{\beta}_S, \hat{\beta}_I) = \argmin_{\beta_S, \beta_I} \hat{R}(\beta_S, \beta_I)$.
-Then $\hat{\bs{\Sigma}}_V$ = $\hat{\bs{\Sigma}}_{LW}$.
+  Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{R}(\beta_S, \beta_I)$.
+Then $\widehat{\bs{\Sigma}}_V$ = $\widehat{\bs{\Sigma}}_{LW}$.
 \end{prop}
@@
   \label{separable}
-  \mathcal{S} = \{\bs{\delta} : \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})\},
+  \mathcal{S} = \{\bs{\delta} : \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}), j\neq k,\quad 
+   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),j=1,\ldots,p)\},
 \end{equation}
@@
 In other words, rules in $\mathcal{S}$ estimate the $jk$th entry of the covariance matrix using a fixed function $t$ of only observations from features $j$ and $k$.
+The sufficient statistics of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is $2\times 2$ matrix
+$\begin{bmatrix}
+s_{jj} & s_{jk} \\
+s_{kj} & s_{kk} 
+\end{bmatrix}$.
 
@@
 To simplify notation, denote $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})^\top$ as $\bs{A}_{jk}$.
-The density of $f(\cdot \mid \bs{\eta}_{jk})$ of $\bs{A}_{jk}$ depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation.
+$f(\cdot \mid \bs{\eta}_{jk})$, the density of $\bs{A}_{jk}$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation.
 When $r_{jk} \ne 1$, $\bs{A}_{jk}$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
@@
 
-Now consider the Bayesian model
+Now consider the Bayesian model for non-diagonal entries
 \begin{equation}
-  \label{bayesian}
+  \label{bayesian nondiagonal}
   \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
   \quad
-  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
+  \bs{\eta} \sim G_{nd}(a,b, \gamma) =  \frac{2}{p(p-1)} \sum_{1\leq k<j\leq p} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
+\end{equation}
+and diagonal entries
+\begin{equation}
+  \label{bayesian diagonal}
+  \bs{A} \mid a \sim \widetilde{f}(\cdot \mid a),
+  \quad
+  a \sim G_d(a) =  \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a).
 \end{equation}
@@
   =\,&
-  \frac{1}{p^2} \sum_{j,k = 1}^p
-  \int \{t(\bs{A}) - \sigma_{jk} \}^2 f(\bs{A} \mid \bs{\eta}_{jk}) d\bs{A}\\
+  \frac{1}{p^2} \{\sum_{1\leq k<j\leq p}^p
+  \int \{t(\bs{A}) - \sigma_{jk} \}^2 f(\bs{A} \mid \bs{\eta}_{jk}) d\bs{A} 
+  +
+  \sum_{j=1}^p
+  \int \{\widetilde{t}(\bs{X}) - \sigma_{jj} \}^2 \widetilde{f}(\bs{X} \mid s_{jj}) d\bs{X}
+  \}\\
   =\,&
-  \int \int \{t(\bs{A}) - g(\bs{\eta})\}^2 f(\bs{A} \mid \bs{\eta}) dG_p(\bs{\eta}) d\bs{A}
-  =
-  \bb{E}[\{t(\bs{A}) -  g(\bs{\eta})\}^2],
+  \frac{p-1}{p}\int \int \{t(\bs{A}) - g(\bs{\eta})\}^2 f(\bs{A} \mid \bs{\eta}) dG_{nd}(\bs{\eta}) d\bs{A}
+  +
+  \frac{1}{p}\int \int \{\widetilde{t}(\bs{X}) - a^2\}^2 \widetilde{f}(\bs{X} \mid a) dG_{d}(a) d\bs{X} \\
+  =\,&
+   \frac{p-1}{p}\bb{E}[\{t(\bs{A}) -  g(\bs{\eta})\}^2]+ \frac{1}{p}\bb{E}[\{\widetilde{t}(\bs{X}) -  a^2\}^2],
 \end{align*}
 where $g(a, b, \gamma) = a b \gamma$ and the final expectation is the Bayes risk of estimating $\sigma_{jk}$.
-The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star(\bs{A}_{jk})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid \bs{A}\}$ minimizes the Bayes risk.
+The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star(\bs{A}_{jk})$ for $j\neq k$ and $\delta^\star_{jj}(\bs{X}) = \widetilde{t}^\star(\bs{X}_{\cdot j})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid \bs{A}\}$ and $\widetilde{t}^\star = \mathbb{E}\{g(a) \mid \bs{X}\}$ minimizes the Bayes risk.
 
@@
 We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the prior $G_p$.
-Under the Bayesian model \ref{bayesian}, and the working assumption that the $\bs{A}_{jk}$ are independent across $jk$, we estimate $G_p$ using
+Under the Bayesian model \ref{bayesian}, and the working assumption that the $\bs{A}_{jk}$ are independent across $jk$, we estimate $G_{nd}$ using
 \begin{equation}
-  \label{Gp hat}
-  \hat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
+  \label{Gnd hat}
+  \widehat{G}_{nd} = \argmax_{G \in \mathcal{G}_{nd}} \prod_{1\leq k<j<p} \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
 \end{equation}
-where $\mathcal{G}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$.
-Of course, the $\bs{A}_{jk}$ are not independent, so $\hat{G}_p$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
-Using $\hat{G}_p$, we estimate the vectorized $\bs{\Sigma}$ using
+
+estimate $G_d$ using
+\begin{equation}
+  \label{Gd hat}
+  \widehat{G}_d = \argmax_{G \in \mathcal{G}_d} \prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG(a),
+\end{equation}
+where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $\mathcal{G}_{d}$ is the family of all distributions supported on $\mathbb{R}$.
+Of course, the $\bs{A}_{jk}$ are not independent, so $\widehat{G}_p$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
+Using $\widehat{G}_p$, we estimate the vectorized $\bs{\Sigma}$ using $ \widehat{\bs{\delta}}(\bs{X})
+  =
+  (\widehat{t}(\bs{A}_{21}), \ldots,  \widehat{t}(\bs{A}_{p,p-1}),\widehat{t}(\bs{X}_{\cdot 1}),\ldots,\widehat{t}(\bs{X}_{\cdot p}))$, where 
 \begin{equation}
   \label{proposed}
-  \hat{\bs{\delta}}(\bs{X})
-  =
-  (\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
+  \widehat{t}(\bs{A}_{jk}) = \frac{\int g(\bs{\eta}) f(\bs{A}_{jk} \mid \bs{\eta})d\widehat{G}_{nd}(\bs{\eta})}{\int f(\bs{A}_{jk} \mid \bs{\eta}) d\widehat{G}_{nd}(\bs{\eta})},
   \quad 
-  \hat{t}(\bs{A}) = \frac{\int g(\bs{\eta}) f(\bs{A} \mid \bs{\eta})d\hat{G}_p(\bs{\eta})}{\int f(\bs{A} \mid \bs{\eta}) d\hat{G}_p(\bs{\eta})}.
+  \widehat{t}(\bs{X}_{\cdot j}) = \frac{\int a^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a)d\widehat{G}_d(a)}{\int \widetilde{f}(\bs{X}_{\cdot j} \mid a) d\widehat{G}_d(a)}.
 \end{equation}
-The $\hat{t}$ estimates the Bayes rule $t^\star$ and $\hat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.
+The $\widehat{t}$ estimates the Bayes rule $t^\star$ and $\widehat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.
 
@@
 In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}.
-Theoretical justification of our proposed $\hat{\bs{\delta}}$ is difficult and is discussed in Section \ref{discussion}.
-Nevertheless, our numerical results in Section \ref{numerical results} show that in practice, our $\hat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.
+Theoretical justification of our proposed $\widehat{\bs{\delta}}$ is difficult and is discussed in Section \ref{discussion}.
+Nevertheless, our numerical results in Section \ref{numerical results} show that in practice, our $\widehat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.
 
@@
 
-Calculating the estimated prior $\hat{G}_p$ \ref{Gp hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
+Calculating the estimated prior $\widehat{G}_p$ \ref{Gp hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
 \citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points.
@@
 Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$.
-We can then use the EM algorithm to estimate the masses $\hat{\bs{w}}=\{\hat{w}_{1},\ldots, \hat{w}_{D}\}$ at those points via the iteration
+We can then use the EM algorithm to estimate the masses $\widehat{\bs{w}}=\{\widehat{w}_{1},\ldots, \widehat{w}_{D}\}$ at those points via the iteration
 \[
-\hat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_l)}
+\widehat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\widehat{w}_{\tau}^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\widehat{w}_l^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_l)}
 \]
@@
 \[
-\hat{\bs{\delta}}(\bs{X})
-=
-(\hat{t}(\bs{A}_{11}), \ldots, \hat{t}(\bs{A}_{pp})),
+\widehat{t}(\bs{A}_{jk}) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{A} \mid \bs{\xi}_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{A} \mid \bs{\xi}_\tau) \widehat{w}_{\tau}},
 \quad
-\hat{t}(\bs{A}) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{A} \mid \bs{\xi}_\tau) \hat{w}_{\tau}}.
+\widehat{t}(\bs{X}_{\cdot j}) = \frac{ \sum_{\tau=1}^D a_{\tau}^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D \widetilde{f}(\bs{X}_{\cdot j}  \mid a_\tau) \widehat{w}_{\tau}}.
 \]
@@
 However, the fact that $G_p$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$.
-Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\hat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$.
+Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\widehat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$.
 This reduces the size of the support set, but even in this case the computation complexity grows like $O(p^2)$.
@@
 
+In our implementation, we use the density function of $\bs{S}_{jk}=\begin{bmatrix}
+s_{jj} & s_{jk} \\
+s_{kj} & s_{kk} 
+\end{bmatrix}$ instead of $\bs{A}_{jk}$ since it is the sufficient statistics of $\bs{A}_{jk}$ and easier to calculate. Based on the property of sufficient statistics, 
+\begin{equation}
+\frac{f(\bs{A}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{A}_{jk}\mid \bs{\xi}_l)} = \frac{f(\bs{S}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{S}_{jk}\mid \bs{\xi}_l)}
+\end{equation}
+So it is equivalent to use $f(\bs{A}_{jk}\mid \bs{\xi}_{\tau})$ in calculation.
 \subsection{\label{posdef}Positive definiteness correction}
@@
 where $\Vert \cdot \Vert$ denotes the Frobenius norm, $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues.
-Projections in terms of other matrix norms are also possible.
 
@@
 In each scenario, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$.
-We generated $200$ replicates and reported average errors under the following three norms, where $\hat{\bs{\Sigma}}$ is the estimated matrix with entries $\hat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
+We generated $200$ replicates and reported average errors under the following three norms, where $\widehat{\bs{\Sigma}}$ is the estimated matrix with entries $\widehat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
 \begin{itemize}
-\item Frobenius: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\hat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$, a version of \ref{frobenius risk},
+\item Frobenius: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\widehat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$, a version of \ref{frobenius risk},
 
-\item Spectral: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\hat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\hat{\bs{\Sigma}} - \bs{\Sigma}$, and
+\item Spectral: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\widehat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\widehat{\bs{\Sigma}} - \bs{\Sigma}$, and
 
-\item Matrix $\ell_1$: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \hat{\sigma}_{jk} - \sigma_{jk} \vert$.
+\item Matrix $\ell_1$: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \widehat{\sigma}_{jk} - \sigma_{jk} \vert$.
 \end{itemize}
@@
 In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness.
+$K$-means clustering is applied on sample grids with $K=p$.
 We compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:
@@
 \item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \ref{separable}.
-It equals our proposed estimator \ref{proposed} except with the true $G_p$ \ref{bayesian} instead of $\hat{G}_p$ \ref{Gp hat}.
+It equals our proposed estimator \ref{proposed} except with the true $G_p$ \ref{bayesian} instead of $\widehat{G}_p$ \ref{Gp hat}.
 The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
@@
 \end{center}
-\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.}
+\caption{Average Frobenius norm errors over 200 replications.
+The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.
+In MSG, $K$-means clustering is applied with $K=p$}
 \label{fig:sim2_frobenius}
@@
 
-Figures \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators.
+Figure \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators.
 Our MSG methods had the lowest or near-lowest errors across all settings except for Model 4 which has high correlations 0.9.
@@
 \citet{osareh2009classification} report the expression of 2308 genes from 63 samples from four groups: 12 neuroblastoma, 20 rhabdomyosarcoma, 8 Burkitt lymphoma, and 23 Ewing's sarcoma patients.
+In MSG, the clustering parameter $K$ is still set as $p=200$.
 We followed the same data preprocessing as \citet{cai2011adaptive} and sorted the genes in decreasing order according to their $F$-statistic
 \begin{equation}
-F = \frac{1}{k-1}\sum_{m=1}^kn_m(\bar{x}_m - \bar{x})^2 /   \frac{1}{n-k}\sum_{m=1}^k (n_m-1)\hat{\sigma}_m^2 
+F = \frac{1}{k-1}\sum_{m=1}^kn_m(\overline{x}_m - \overline{x})^2 /   \frac{1}{n-k}\sum_{m=1}^k (n_m-1)\widehat{\sigma}_m^2 
 \end{equation}
-where $k = 4$ is the number of patient categories, $n_m$, $\bar{x}_m$, and $\hat{\sigma}_k$ represent the sample size, sample mean, and sample variance of the gene's expression in the $m$th category, respectively, and $\bar{x}$ is the global mean.
+where $k = 4$ is the number of patient categories, $n_m$, $\overline{x}_m$, and $\widehat{\sigma}_k$ represent the sample size, sample mean, and sample variance of the gene's expression in the $m$th category, respectively, and $\overline{x}$ is the global mean.
 We proceeded with the top 40 genes and bottom 160 genes.
@@
 
+\section*{Supporting Information}
+Web Appendices, Tables, and Figures referenced in Section 3 are available with this paper at the Biometrics website on Wiley Online Library.
+
 \appendix
@@
 \begin{proof}
-  We first rewrite the risk estimate $\hat{R}(\beta_S, \beta_I)$.
-Define $\bs{M}=(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$.
+  We first rewrite the risk estimate $\widehat{R}(\beta_S, \beta_I)$.
+Define $\bs{M}=(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$.
 Then the unbiased risk estimator can be re-written as
   \[
-  \hat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M},
+  \widehat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M},
   \]
@@
   \begin{align*}
-    \hat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }(\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}\bs{v_S}-\bs{M}),\\
+    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }(\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}\bs{v_S}-\bs{M}),\\
     &
-    \hat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\hat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
+    \widehat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\widehat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
   \end{align*}
 
-  We will need to show $\hat{\mu} b_n^2 / d_n^2 = \hat{\beta}_I$ and $\hat{\beta}_I/\hat{\mu} + \hat{\beta}_S = 1$. Since
+  We will need to show $\widehat{\mu} b_n^2 / d_n^2 = \widehat{\beta}_I$ and $\widehat{\beta}_I/\widehat{\mu} + \widehat{\beta}_S = 1$. Since
 $$\bs{Z}^\top \bs{Z} = \begin{pmatrix}
@@
 \bs{Z^\top S}-\bs{M} =  \begin{pmatrix}
-&\sum_{j,k=1}^ps_{jk}^2 - \hat{\Delta}_{jk}^2\\
+&\sum_{j,k=1}^ps_{jk}^2 - \widehat{\Delta}_{jk}^2\\
 &\sum_{j=1}^p s_{jj}
@@
 \begin{align*}
-\hat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1} \bs{Z^\top S} \\
+\widehat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1} \bs{Z^\top S} \\
 &= \frac{1}{p^3d_n^2}\begin{pmatrix}
@@
 \end{pmatrix}\begin{pmatrix}
-\sum_{j,k=1}^ps_{jk}^2 - \hat{\Delta}_{jk}^2\\
+\sum_{j,k=1}^ps_{jk}^2 - \widehat{\Delta}_{jk}^2\\
 \sum_{j=1}^p s_{jj}
@@
 &= \frac{1}{p^3d_n^2}\begin{pmatrix}
- p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
- (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)
+ p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
+ (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)
 \end{pmatrix}.
 \end{align*}
-The second component of $\hat{\bs{\beta}}$ equals $\hat{\beta}_I$, so
+The second component of $\widehat{\bs{\beta}}$ equals $\widehat{\beta}_I$, so
 \begin{align*}
-\hat{\beta}_I &= \frac{1}{p^3d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)\\
-&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)/p^2 \} / d_n^2
-=\hat{\mu}\frac{b_n^2}{d_n^2}.
+\widehat{\beta}_I &= \frac{1}{p^3d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)\\
+&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)/p^2 \} / d_n^2
+=\widehat{\mu}\frac{b_n^2}{d_n^2}.
 \end{align*}
@@
 \begin{align*}
-\hat{\beta}_I /\hat{\mu} + \hat{\beta}_S &= \frac{1}{p^3d_n^2}p\sum_{j,k=1}^p \{s_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2\}\\
+\widehat{\beta}_I /\widehat{\mu} + \widehat{\beta}_S &= \frac{1}{p^3d_n^2}p\sum_{j,k=1}^p \{s_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2\}\\
 &=\frac{1}{p^3d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}

### 2021-09-20 msg.tex

@@
 However, in modern problems the number of features can be of the same order as or exceed the sample size, and the standard sample covariance matrix estimator behaves poorly in this regime.
-To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrix.
-These can roughly be divided into two groups, according to whether they impose assumptions about the structure of population covariance matrix.
+To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrces.
+These can roughly be divided into two groups, according to whether they impose assumptions of the structure of population covariance matrix.
 
@@
 Another class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension.
-Other common structured methods assume that the covariance matrix is banding \citep{li2017estimation} or Toeplitz matrix \citep{liu2017covariance}.
+Other common structured methods assume that the covariance matrix is banded \citep{li2017estimation} or Toeplitz matrix \citep{liu2017covariance}.
 
-In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet can still outperform the sample covariance matrix.
+In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have less estimation error comparing to the sample covariance matrix.
 A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix.
@@
 Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance.
-However, they modify only the sample eigenvalues and not the sample eigenvectors.
+However, as the class of rotation-equivariant estimators \citep{stein1975estimation, stein1986lectures}, nonlinear shrinkage estimators modify only the sample eigenvalues and keep the sample eigenvectors.
 It is known that sample eigenvectors are not consistent estimators of population eigenvectors when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}.
@@
 Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis.
+Our approach is similar to \citet{dey2018corshrink}, which also applied empirical Bayes shrinkage method to estimate correlation matrix.
+Different from our method, their corShrink method has different target and they do not shrink standard deviations.
 
 The article is organized as follows.
-In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach.
-In Section \ref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset.
-Finally, Section \ref{sec:discussion} concludes with a discussion.
+In Section \eqref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach.
+In Section \eqref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset.
+Finally, Section \eqref{sec:discussion} concludes with a discussion.
 Our procedure is implemented in the R package \verb|cole|, available on GitHub.
@@
 Our proposed approach is motivated by two observations.
-First, \ref{frobenius risk} shows that estimating $\bs{\Sigma}$ under Frobenius risk is equivalent to simultaneously estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under a loss function that aggregates errors across components.
+First, \eqref{frobenius risk} shows that estimating $\bs{\Sigma}$ under Frobenius risk is equivalent to simultaneously estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under a loss function that aggregates errors across components.
 Second, this type of vector estimation problem has been well-studied in the compound decision literature.
@@
 where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$.
-A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is squared error loss \citep{johnstone2017gaussian}.
+A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.
 
@@
 We propose to apply some of these recent vector estimation ideas to covariance matrix estimation.
-Clearly, covariance matrix estimation under the Frobenius risk \ref{frobenius risk} can be viewed as a compound decision problem.
+Clearly, covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can be viewed as a compound decision problem.
 Furthermore, some existing covariance matrix estimation procedures can already be interpreted as taking a vector approach.
-The sample covariance matrix $\bs{S}$, for instance, can be thought of as estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$ using maximum likelihood.
+The sample covariance matrix $\bs{S}=\frac{1}{n}\bs{X}^{T}\bs{X}$, for instance, can be thought of as estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$ using maximum likelihood.
 Less trivially, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective.
@@
 where $\widehat{\mu} = \text{tr}(\bs{S})/p$, and $\widehat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\widehat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$.
-Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \ref{frobenius risk}.
+Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \eqref{frobenius risk}.
 We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
@@
 \end{equation}
-where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. It is straightforward to show that
-$$\widehat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
-is an unbiased estimate of the risk \ref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n$ is the sample variance of $s_{jk}$.
-The optimal estimator in this class can now be chosen by minimizing $\widehat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
+where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. As proved in appendix, we can show that
+$$URE(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
+is an unbiased estimator of the risk \eqref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n(n-1)$ is an unbiased estimator of variance of $s_{jk}$.
+The optimal estimator in this class can now be chosen by minimizing $URE(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
 It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}.
-Proposition \ref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \ref{linear model}.
+Proposition \eqref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \eqref{linear model}.
 \begin{prop}
@@
   Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{R}(\beta_S, \beta_I)$.
-Then $\widehat{\bs{\Sigma}}_V$ = $\widehat{\bs{\Sigma}}_{LW}$.
+Then the coefficients $\widehat{\beta}_S$, $\widehat{\beta}_I $ and $\widehat{\rho}_{\text{LW}}$, $(1-\widehat{\rho}_{\text{LW}})\widehat{\mu}$ are very closed.
+
   \end{prop}
@@
   \label{separable}
-  \mathcal{S} = \{\bs{\delta} : \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}), j\neq k,\quad 
-   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),j=1,\ldots,p)\},
+  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
+   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
 \end{equation}
@@
 In other words, rules in $\mathcal{S}$ estimate the $jk$th entry of the covariance matrix using a fixed function $t$ of only observations from features $j$ and $k$.
-The sufficient statistics of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is $2\times 2$ matrix
-$\begin{bmatrix}
+The sufficient statistics of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is $2\times 2$ matrix of $S$ denoted by $\bs{S}_{jk}=
+\begin{bmatrix}
 s_{jj} & s_{jk} \\
@@
 We propose to search for the optimal estimator within $\mathcal{S}$.
-This is sensible because $\mathcal{S}$ includes the sample covariance $(s_{11}, \ldots, s_{pp})^\top$, the class of linear estimators \ref{linear class} used by \citet{ledoit2004well}, which can be expressed as
-\[
-\delta_{jk}(\bs{X}) = \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n + \beta_I I(\bs{X}_{\cdot j} = \bs{X}_{\cdot k}),
-\]
+This is sensible because $\mathcal{S}$ includes the sample covariance $(s_{11}, \ldots, s_{pp})^\top$, the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
+\begin{align*}
+\delta_{jk}(\bs{X})&= \delta_{kj}(\bs{X})= \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n, \, 1\leq k<j\leq p\\
+\delta_{jj}(\bs{X})&=\beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot j} / n+\beta_I, \, j=1,\ldots,p
+\end{align*}
 and the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}.
-Therefore the optimal separable estimator $\bs{\delta}^\star$ that minimizes the scaled squared Frobenius risk \ref{frobenius risk} will perform at least as well as these three estimators, and may perform better.
+Therefore the optimal separable estimator $\bs{\delta}^\star$ that minimizes the scaled squared Frobenius risk \eqref{frobenius risk} will perform at least as well as these three estimators, and may perform better.
 Targeting the optimal separable rule is standard in the compound decision literature \citep{zhang2003compound}.
@@
 We take a similar approach here.
-To simplify notation, denote $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})^\top$ as $\bs{A}_{jk}$.
-$f(\cdot \mid \bs{\eta}_{jk})$, the density of $\bs{A}_{jk}$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation.
-When $r_{jk} \ne 1$, $\bs{A}_{jk}$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
+$f(\cdot \mid \bs{\eta}_{jk})$, the density of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation.
+When $j\neq k$, $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
 \[
@@
 \]
-When $r_{jk} = 1$, $\bs{A}_{jk}$ consists of mean-zero univariate normals with variances $\sigma_j^2$.
 
@@
   \label{bayesian nondiagonal}
-  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
+  (\bs{X},\bs{X}^{'}) \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
   \quad
@@
   \label{bayesian diagonal}
-  \bs{A} \mid a \sim \widetilde{f}(\cdot \mid a),
+  \bs{X} \mid a \sim \widetilde{f}(\cdot \mid a),
   \quad
@@
 \end{equation}
-By the fundamental theorem of compound decisions \citep{robbins1951asymptotically, jiang2009general}, this is closely related to the vectorized covariance matrix estimation problem under Frobenius risk \ref{frobenius risk}: for any separable $\bs{\delta} \in \mathcal{S}$ \ref{separable}, the Frobenius risk can be written as
+the $G_d$ is the marginal density for the first parameter of $G_{nd}$
+\begin{equation}
+\label{eq:marginal}
+\int dG_{nd}(a,b,\gamma)dbd\gamma = dG_d(a)
+\end{equation}
+By the fundamental theorem of compound decisions \citep{robbins1951asymptotically, jiang2009general}, this is closely related to the vectorized covariance matrix estimation problem under Frobenius risk \eqref{frobenius risk}.
+Then we have the following proposition.
+
+\begin{prop}\label{prop:bayes risk}
+For any $\bs{\delta} \in \mathcal{S}$ \eqref{separable}, the Frobenius risk can be written as the convex combination of Bayesian risks with respect to priors $G_{nd}$ and $G_d$.
 \begin{align*}
@@
   =\,&
-  \frac{1}{p^2} \{\sum_{1\leq k<j\leq p}^p
-  \int \{t(\bs{A}) - \sigma_{jk} \}^2 f(\bs{A} \mid \bs{\eta}_{jk}) d\bs{A} 
-  +
+  \frac{1}{p^2} \{2\sum_{1\leq k<j\leq p}^p
+  \int \{t(\bs{X},\bs{X}^{\prime}) - \sigma_{jk} \}^2 f(t(\bs{X},\bs{X}^{\prime}) \mid \bs{\eta}_{jk}) d\bs{X}d\bs{X}^{\prime}\\
+  +\,&
   \sum_{j=1}^p
@@
   =\,&
-  \frac{p-1}{p}\int \int \{t(\bs{A}) - g(\bs{\eta})\}^2 f(\bs{A} \mid \bs{\eta}) dG_{nd}(\bs{\eta}) d\bs{A}
-  +
+  \frac{p-1}{p}\int \int \{t(\bs{X},\bs{X}^{\prime}) - g(\bs{\eta})\}^2 f((\bs{X},\bs{X}^{\prime})\mid \bs{\eta}) dG_{nd}(\bs{\eta}) d\bs{X}d\bs{X}^{\prime}\\
+  +\,&
   \frac{1}{p}\int \int \{\widetilde{t}(\bs{X}) - a^2\}^2 \widetilde{f}(\bs{X} \mid a) dG_{d}(a) d\bs{X} \\
   =\,&
-   \frac{p-1}{p}\bb{E}[\{t(\bs{A}) -  g(\bs{\eta})\}^2]+ \frac{1}{p}\bb{E}[\{\widetilde{t}(\bs{X}) -  a^2\}^2],
+   \frac{p-1}{p}\bb{E}[\{t(\bs{X},\bs{X}^{\prime}) -  g(\bs{\eta})\}^2]+ \frac{1}{p}\bb{E}[\{\widetilde{t}(\bs{X}) -  a^2\}^2],
 \end{align*}
 where $g(a, b, \gamma) = a b \gamma$ and the final expectation is the Bayes risk of estimating $\sigma_{jk}$. 
-The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star(\bs{A}_{jk})$ for $j\neq k$ and $\delta^\star_{jj}(\bs{X}) = \widetilde{t}^\star(\bs{X}_{\cdot j})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid \bs{A}\}$ and $\widetilde{t}^\star = \mathbb{E}\{g(a) \mid \bs{X}\}$ minimizes the Bayes risk.
+\end{prop}
+The decision rule which minimizes the Frobenius risk depends on all the $\sigma_j$ and $r_{jk}$.
+The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star((\bs{X}_{\cdot j},\bs{X}_{\cdot k}))$ for $j> k$ and $\delta^\star_{jj}(\bs{X}) = \widetilde{t}^\star(\bs{X}_{\cdot j})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid (\bs{X},\bs{X}^{'})\}$ and $\widetilde{t}^\star = \mathbb{E}\{a^2 \mid \bs{X}\}$ minimizes the Bayes risk.
 
 Based on this result, we propose the following empirical Bayes procedure.
-We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the prior $G_p$.
-Under the Bayesian model \ref{bayesian}, and the working assumption that the $\bs{A}_{jk}$ are independent across $jk$, we estimate $G_{nd}$ using
+We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the priors $G_{nd}$ and $G_d$.
+Under the Bayesian model \eqref{bayesian}, and the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent across $jk$, we estimate $G_{nd}$ using
 \begin{equation}
   \label{Gnd hat}
-  \widehat{G}_{nd} = \argmax_{G \in \mathcal{G}_{nd}} \prod_{1\leq k<j<p} \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
+  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}_{nd}} \{\prod_{1\leq k\leq j<p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\} 
+  \{\prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG_d(a)\},
 \end{equation}
-
-estimate $G_d$ using
-\begin{equation}
-  \label{Gd hat}
-  \widehat{G}_d = \argmax_{G \in \mathcal{G}_d} \prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG(a),
-\end{equation}
-where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $\mathcal{G}_{d}$ is the family of all distributions supported on $\mathbb{R}$.
-Of course, the $\bs{A}_{jk}$ are not independent, so $\widehat{G}_p$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
-Using $\widehat{G}_p$, we estimate the vectorized $\bs{\Sigma}$ using $ \widehat{\bs{\delta}}(\bs{X})
-  =
-  (\widehat{t}(\bs{A}_{21}), \ldots,  \widehat{t}(\bs{A}_{p,p-1}),\widehat{t}(\bs{X}_{\cdot 1}),\ldots,\widehat{t}(\bs{X}_{\cdot p}))$, where 
+where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $G_d$ is determined by $G_{nd}$ as indicated in \eqref{eq:marginal}.
+Of course, the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are not independent, so $\widehat{G}_{nd}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
+Using $\widehat{G}_{nd}$ and $\widehat{G}_d$, we estimate the vectorized $\bs{\Sigma}$ using
+$$ \widehat{\bs{\delta}}(\bs{X})=
+  (\widehat{t}(\bs{X}_{\cdot 2},\bs{X}_{\cdot 1}), \ldots,  \widehat{t}(\bs{X}_{\cdot p},\bs{X}_{\cdot (p-1)}),\widehat{t}(\bs{X}_{\cdot 1}),\ldots,\widehat{t}(\bs{X}_{\cdot p}))$$
+  , where 
 \begin{equation}
   \label{proposed}
-  \widehat{t}(\bs{A}_{jk}) = \frac{\int g(\bs{\eta}) f(\bs{A}_{jk} \mid \bs{\eta})d\widehat{G}_{nd}(\bs{\eta})}{\int f(\bs{A}_{jk} \mid \bs{\eta}) d\widehat{G}_{nd}(\bs{\eta})},
+  \widehat{t}(\bs{X}_{\cdot j},\bs{X}_{\cdot k}) = \frac{\int g(\bs{\eta}) f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta})d\widehat{G}_{nd}(\bs{\eta})}{\int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) d\widehat{G}_{nd}(\bs{\eta})},
   \quad 
@@
 \end{equation}
-The $\widehat{t}$ estimates the Bayes rule $t^\star$ and $\widehat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.
+The $\widehat{t}$ estimates the Bayes rule $t^\star$, $\widetilde{t}^\star$ and $\widehat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.
 
 Our proposed procedure is an example of what \citet{efron2014two} calls $g$-modeling, an approach to empirical Bayes problems that proceeds by modeling the prior.
-A major advantage of nonparametric estiation of the prior is that it allows the data itself to determine how best to shrink the estimator.
+A major advantage of nonparametric estimation of the prior is that it allows the data itself to determine how best to shrink the estimator.
+This kind of procedure is also adopted in \citet{dey2018corshrink}.
 In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}.
-Theoretical justification of our proposed $\widehat{\bs{\delta}}$ is difficult and is discussed in Section \ref{discussion}.
-Nevertheless, our numerical results in Section \ref{numerical results} show that in practice, our $\widehat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.
+Theoretical justification of our proposed $\widehat{\bs{\delta}}$ is difficult and is discussed in Section \eqref{discussion}.
+Nevertheless, our numerical results in Section \eqref{numerical results} show that in practice, our $\widehat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.
 
@@
 
-Calculating the estimated prior $\widehat{G}_p$ \ref{Gp hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
+Calculating the estimated prior $\widehat{G}_{nd}$ \eqref{Gnd hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
 \citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points.
@@
 We can then use the EM algorithm to estimate the masses $\widehat{\bs{w}}=\{\widehat{w}_{1},\ldots, \widehat{w}_{D}\}$ at those points via the iteration
-\[
-\widehat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\widehat{w}_{\tau}^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\widehat{w}_l^{(k-1)} f(\bs{A}_{jk} \mid \bs{\xi}_l)}
-\]
+\begin{equation}
+\label{eq:iteration}
+\widehat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\widehat{w}_{\tau}^{(k-1)} f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\widehat{w}_l^{(k-1)} f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)}
+\end{equation}
 over $k$.
 Early stopping of the EM algorithm can be useful \citep{koenker2019comment}, and more sophisticated convex optimization procedures can be used as well \citep{koenker2014convex}.
-Our proposed estimator \ref{proposed} then becomes
+Our proposed estimator \eqref{proposed} then becomes
 \[
-\widehat{t}(\bs{A}_{jk}) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{A} \mid \bs{\xi}_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{A} \mid \bs{\xi}_\tau) \widehat{w}_{\tau}},
+\widehat{t}(\bs{X}_j,\bs{X}_k) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_\tau) \widehat{w}_{\tau}},
 \quad
@@
 Ideally, the grid points should be chosen to densely cover the parameter space.
-However, the fact that $G_p$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$.
+However, the fact that $G_{nd}$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$.
 Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\widehat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$.
@@
 Let $s_j$, $s_k$, and $\gamma_{jk}$ be the sample variances and correlation between the $j$th and $k$th covariates.
-We first apply $K$-means clustering to identify $K$ clusters among the $p (p-1) / 2$ off-diagonal sample points $(s_j,s_k,\gamma_{jk})$ and $\lceil{K^{1/2}}\rceil$ clusters among the $p$ diagonal sample points $(s_j,s_j,1)$.
-We then use the $K+\lceil{K^{1/2}}\rceil$ cluster centroids as our support points.
-We cluster the off- and on-diagonal observations separately to ensure that the support points $\bs{\xi}_\tau$ are such that $f(\bs{A}_{jk} \mid \bs{\xi}_\tau) \ne 0$ when both $j = k$ and $j \ne k$.
-Figure \ref{fig:sim1_frobenius} shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table \ref{tab:sim1_time} shows that they can be significantly faster.
+We first apply $K$-means clustering to identify $K$ clusters among the $p (p-1) / 2$ lower triangular off-diagonal sample points $(s_j,s_k,\gamma_{jk})$ and find their symmetric points by exchanging their first two dimensions.
+We then use the $2K$ cluster centroids as our support points.
+Figure \ref{fig:sim1_frobenius} shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table \eqref{tab:sim1_time} shows that they can be significantly faster.
 
@@
 s_{kj} & s_{kk} 
-\end{bmatrix}$ instead of $\bs{A}_{jk}$ since it is the sufficient statistics of $\bs{A}_{jk}$ and easier to calculate. Based on the property of sufficient statistics, 
+\end{bmatrix}$ instead of $(\bs{X}_j,\bs{X}_k)$ since it is the sufficient statistics of $(\bs{X}_j,\bs{X}_k)$ with respect to $\bs{C}_{jk}$. Based on the property of sufficient statistics, 
 \begin{equation}
-\frac{f(\bs{A}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{A}_{jk}\mid \bs{\xi}_l)} = \frac{f(\bs{S}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{S}_{jk}\mid \bs{\xi}_l)}
+\frac{f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_{\tau})}{f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)} = \frac{f(\bs{S}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{S}_{jk}\mid \bs{\xi}_l)}
 \end{equation}
-So it is equivalent to use $f(\bs{A}_{jk}\mid \bs{\xi}_{\tau})$ in calculation.
+it is equivalent to use $f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_{\tau})$ instead to simplify the calculation.
+Although we assume $\bs{X}$ has mean of zero theoretically, in practice, we usually do not know the true means.
+In such case, we can centralize each $\bs{X}_{\cdot j}$ by their sample means.
+
 \subsection{\label{posdef}Positive definiteness correction}
-Our proposed estimator \ref{proposed} is not guaranteed to be positive-definite.
+Our proposed estimator \eqref{proposed} is not guaranteed to be positive-definite.
 To correct this, we reshape our vector estimator back into a matrix and then identify the closest positive-definite matrix.
@@
 \section{\label{numerical results}Numerical Results}
+In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness.
+In our EM algorithm, the maximum number of iterations \eqref{eq:iteration} is 200 and it early stops when $\frac{L(\bs{w}^{(k+1)})-L(\bs{w}^{(k)})}{L(\bs{w}^{(k)})}\leq 10^{-4}$, where $L(\bs{w}^{(k)}) = \sum_{1\leq k\leq j\leq p}\log f(\bs{X}_{\cdot j},\bs{X}_{\cdot k}|\bs{w}^{(k)})$ is composite log likelihood of the observations.
 \subsection{\label{models}Models}
@@
 This setting was the same as Model 3 except with correlation parameter 0.9.
-This high level of dependence tested the robustness of the pairwise composite likelihood estimator \ref{Gp hat}.
+This high level of dependence tested the robustness of the pairwise composite likelihood estimator \eqref{Gp hat}.
 
@@
 This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
+
+\item Model 6.
+With $\bs{U}$ a randomly generated orthogonal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p -3$ entries equaled 1.
 \end{itemize}
@@
 In each scenario, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$.
-We generated $200$ replicates and reported average errors under the following three norms, where $\widehat{\bs{\Sigma}}$ is the estimated matrix with entries $\widehat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
-\begin{itemize}
-\item Frobenius: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\widehat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$, a version of \ref{frobenius risk},
-
-\item Spectral: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\widehat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\widehat{\bs{\Sigma}} - \bs{\Sigma}$, and
-
-\item Matrix $\ell_1$: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \widehat{\sigma}_{jk} - \sigma_{jk} \vert$.
-\end{itemize}
+Although $\bb{E}\bs{X}=0$ in our setting, we assume it is unknown and use $\bs{S}=\frac{1}{n-1}(\bs{X}-\bs{\bar{X}})^\top(\bs{X}-\bs{\bar{X}})$.
+We generated $200$ replicates and reported median errors under the following Frobenius norms, where $\widehat{\bs{\Sigma}}$ is the estimated matrix with entries $\widehat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
+$$\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\widehat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$$
+which is a version of \eqref{frobenius risk}.
+%\item Spectral: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\widehat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\widehat{\bs{\Sigma}} - \bs{\Sigma}$, and
+%
+%\item Matrix $\ell_1$: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \widehat{\sigma}_{jk} - \sigma_{jk} \vert$.
 
 \subsection{\label{optimalK}Clustering-based exemplar algorithm}
-We first studied the behavior of our $K$-means clustering-based exemplar algorithm for different $K$, described in Section \ref{implementation}.
+We first studied the behavior of our $K$-means clustering-based exemplar algorithm for different $K$, described in Section \eqref{implementation}.
 For a given $p$, we let $K = rp$ for different ratios ratios $r=2,1,0.5,0.25$.
@@
 \end{center}
-\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.}
+\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, Orth and Spiked panels correspond to Models 1 through 5, respectively.}
 \label{fig:sim1_frobenius}
@@
 
-Figure \ref{fig:sim1_frobenius} presents the Frobenius norm error estimates from Model 1 to Model 5.
-Table \ref{tab:sim1_time} shows the running time only for Model 1, because the running time does not vary much across different models.
+Figure \ref{fig:sim1_frobenius} presents the Frobenius norm error estimates from Model 1 to Model 6.
+Table \eqref{tab:sim1_time} shows the running time only for Model 1, because the running time does not vary much across different models.
 The results show that different $K$ exhibit similar performance and are comparable to the full exemplar method.
@@
 
-In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness.
-$K$-means clustering is applied on sample grids with $K=p$.
-We compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:
+In this subsection, we compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:
 
@@
 
-\item Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \ref{linear model}.
+\item Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \eqref{linear model}.
 
@@
 We fixed the parameter at 2, as recommended.
+
+\item CorShrink: Empirical Bayes shrinkage estimation of correlation matrix \citep{dey2018corshrink}.
 \end{itemize}
@@
 
-\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \ref{separable}.
-It equals our proposed estimator \ref{proposed} except with the true $G_p$ \ref{bayesian} instead of $\widehat{G}_p$ \ref{Gp hat}.
+\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \eqref{separable}.
+It equals our proposed estimator \eqref{proposed} except with the true $G_p$ \eqref{bayesian} instead of $\widehat{G}_p$ \eqref{Gp hat}.
 The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
@@
 Model 2 was especially interesting because the standard deviation and correlations were related.
-Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_p$ \ref{bayesian} and leverage it to provide much more accurate estimates.
+Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_{nd}$ \eqref{bayesian} and leverage it to provide much more accurate estimates.
 The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5.
@@
 
-Finally, the simulations show that the class of separable estimators \ref{separable} proposed in this paper is fundamentally different from the class of rotation-invariant estimators, as the oracle optimal estimators in these two classes behave very differently.
+Finally, the simulations show that the class of separable estimators \eqref{separable} proposed in this paper is fundamentally different from the class of rotation-invariant estimators, as the oracle optimal estimators in these two classes behave very differently.
 For example, the oracle separable estimator had vanishing risk in Model 2, while the oracle rotiation-invariant estimator does not.
@@
 Covariance matrix estimation is often used to reconstruct gene networks \citep{markowetz2007inferring}.
-We applied our MSG and the other covariance matrix estimators described in Section \ref{compared} to gene network estimation using data from a small round blue-cell tumor microarray experiment \citep{khan2001classification}, which was also studied by \citet{cai2011adaptive}.
+We applied our MSG and the other covariance matrix estimators described in Section \eqref{compared} to gene network estimation using data from a small round blue-cell tumor microarray experiment \citep{khan2001classification}, which was also studied by \citet{cai2011adaptive}.
 \citet{osareh2009classification} report the expression of 2308 genes from 63 samples from four groups: 12 neuroblastoma, 20 rhabdomyosarcoma, 8 Burkitt lymphoma, and 23 Ewing's sarcoma patients.
@@
 
-Table \ref{tab:tab1} reports the average errors across the replications.
+Table \eqref{tab:tab1} reports the average errors across the replications.
 Our MSG methods had the lowest average error.
@@
 
-The class of separable covariance matrix estimators \ref{separable} that we proposed in this paper appears to be very promising.
-Many existing procedures already explicitly or implicitly target this class, and our proposed estimate \ref{proposed} of the optimal separable estimator outperforms a number of existing covariance matrix estimators.
+The class of separable covariance matrix estimators \eqref{separable} that we proposed in this paper appears to be very promising.
+Many existing procedures already explicitly or implicitly target this class, and our proposed estimate \eqref{proposed} of the optimal separable estimator outperforms a number of existing covariance matrix estimators.
 This is surprising because our approach vectorizes the matrix and therefore cannot take matrix structure, such as positive-definiteness, into account.
@@
 \citet{saha2020nonparametric} extended these results to multivariate $\bs{Y}_i \sim N(\bs{\theta}_i, \bs{\Sigma}_i)$ with a multivariate prior on the $\bs{\theta}_i$, but assumed that the $\bs{Y}_i$ were independent.
-In contrast, our covariance matrix estimator is built from arbitrarily dependent $\bs{A}_{jk}$.
+In contrast, our covariance matrix estimator is built from arbitrarily dependent $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$.
 These imposes significant theoretical difficulties that will require substantial work to address; we leave this for future research.
 
-Finally, we have so far assumed that our data multivariate normal.
-To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \ref{Gp hat} and in our proposed estimator \ref{proposed}.
+Finally, we have so far assumed that our data are multivariate normal.
+To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gp hat} and in our proposed estimator \eqref{proposed}.
 If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.
@@
 \section{}
-\subsection{Proof of Proposition \ref{prop:linear}}
+\subsection{Proof of Proposition \eqref{prop:linear}}
 
 \begin{proof}
+The estimator in \citet{ledoit2004well} is
+\begin{equation} 
+\label{eq:lw}
+\bs{\Sigma}_{LW} = \frac{d_n^2-b_n^2}{d_n^2}\bs{S}+ \frac{b_n^2}{d_n^2}\hat{\mu}\bs{I}
+\end{equation}
+where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{n^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.
+
   We first rewrite the risk estimate $\widehat{R}(\beta_S, \beta_I)$.
@@
   \[
-  \widehat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M},
+ p^2 \widehat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
   \]
@@
   \begin{align*}
-    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }(\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}\bs{v_S}-\bs{M}),\\
+    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\widehat{R}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
     &
@@
 
-  We will need to show $\widehat{\mu} b_n^2 / d_n^2 = \widehat{\beta}_I$ and $\widehat{\beta}_I/\widehat{\mu} + \widehat{\beta}_S = 1$. Since
+  We will need to show $\widehat{\mu} b_n^2 / d_n^2 \to \widehat{\beta}_I$ and $\widehat{\beta}_I/\widehat{\mu} + \widehat{\beta}_S = 1$. Since
 $$\bs{Z}^\top \bs{Z} = \begin{pmatrix}
@@
 and
-$\text{det}( \bs{Z}^\top \bs{Z} ) = p\sum_{j,k=1}^ps_{jk}^2 - (\sum_{j=1}^p s_{jj})^2 = p^3d_n^2$, it follows that
-$$(\bs{Z}^\top \bs{Z})^{-1} = \frac{1}{p^3d_n^2}\begin{pmatrix}
+$\text{det}( \bs{Z}^\top \bs{Z} ) = p\sum_{j,k=1}^ps_{jk}^2 - (\sum_{j=1}^p s_{jj})^2 = p^2d_n^2$, it follows that
+$$(\bs{Z}^\top \bs{Z})^{-1} = \frac{1}{p^2d_n^2}\begin{pmatrix}
 p & -\sum_{j=1}^p s_{jj}\\
@@
 and in addition
-$$\bs{Z}^\top \bs{S} = \begin{pmatrix}
+$$\bs{Z}^\top \bs{v_S}= \begin{pmatrix}
 \sum_{j,k=1}^ps_{jk}^2\\
@@
 \quad 
-\bs{Z^\top S}-\bs{M} =  \begin{pmatrix}
+\bs{Z^\top \bs{v_S}}-\bs{M} =  \begin{pmatrix}
 &\sum_{j,k=1}^ps_{jk}^2 - \widehat{\Delta}_{jk}^2\\
@@
 \begin{align*}
-\widehat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1} \bs{Z^\top S} \\
-&= \frac{1}{p^3d_n^2}\begin{pmatrix}
+\widehat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z^\top \bs{v_S}}-\bs{M} ) \\
+&= \frac{1}{p^2d_n^2}\begin{pmatrix}
 p & -\sum_{j=1}^p s_{jj}\\
@@
 \end{pmatrix}\begin{pmatrix}
-\sum_{j,k=1}^ps_{jk}^2 - \widehat{\Delta}_{jk}^2\\
+\sum_{j,k=1}^p(s_{jk}^2 - \widehat{\Delta}_{jk}^2)\\
 \sum_{j=1}^p s_{jj}
 \end{pmatrix}\\
-&= \frac{1}{p^3d_n^2}\begin{pmatrix}
+&= \frac{1}{p^2d_n^2}\begin{pmatrix}
  p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
@@
 \begin{align*}
-\widehat{\beta}_I &= \frac{1}{p^3d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)\\
-&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)/p^2 \} / d_n^2
-=\widehat{\mu}\frac{b_n^2}{d_n^2}.
+\widehat{\beta}_I &= \frac{1}{p^2d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)\\
+&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)/p \} / d_n^2
+=\widehat{\mu}\frac{b_n^2}{d_n^2}\frac{n}{n-1}.
 \end{align*}
+which is very closed to the coefficient of $\bs{I}$ in \eqref{eq:lw}
 Furthermore,
 \begin{align*}
-\widehat{\beta}_I /\widehat{\mu} + \widehat{\beta}_S &= \frac{1}{p^3d_n^2}p\sum_{j,k=1}^p \{s_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2\}\\
-&=\frac{1}{p^3d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}
+\widehat{\beta}_I /\widehat{\mu} + \widehat{\beta}_S &= \frac{1}{p^2d_n^2}\sum_{j,k=1}^p \{ps_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2\}\\
+&=\frac{1}{p^2d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}
 =
@@
 
+\subsection{Proof of Proposition \eqref{prop:bayes risk}}
+\begin{proof}
+Using the fact that when $\bb{E}\bs{X}=\bs{0}$, $\bb{E}s_{jk} = \sigma_{jk}$ for all $j,k=1,\ldots,p$
+\begin{align*}
+R(\beta_{S},\beta_{I}) &=\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} \{ (\beta_{S}s_{jk}+\beta_Iu_{jk}-\sigma_{jk})^2 \} \\
+& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} [s_{jk}-\sigma_{jk} - \{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}]^2\\
+& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E}[(s_{jk}-\sigma_{jk})^2+ \bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2-2(1-\beta_{S})\bb{E}s_{jk}(s_{jk}-\sigma_{jk})]\\
+& =\frac{1}{p^2}\sum_{j,k=1}^p [(2\beta_{S}-1)\text{Var}(s_{jk})+\bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2] \\
+& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\widehat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2]
+\end{align*}
+\end{proof}
 \label{lastpage}

### 2021-09-27 msg.tex

@@
 where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. As proved in appendix, we can show that
-$$URE(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
-is an unbiased estimator of the risk \eqref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n(n-1)$ is an unbiased estimator of variance of $s_{jk}$.
-The optimal estimator in this class can now be chosen by minimizing $URE(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
+$$\textsc{RE}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
+is an approximately unbiased estimator of the risk \eqref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ is an approximately unbiased estimator of variance of $s_{jk}$.
+The optimal estimator in this class can now be chosen by minimizing $RE(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
 It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}.
-Proposition \eqref{prop:linear} shows that this subspace shrinkage estimator is identical to the \citet{ledoit2004well} estimator \eqref{linear model}.
+Proposition \eqref{prop:linear} shows that this subspace shrinkage estimator is closely related to the \citet{ledoit2004well} estimator \eqref{linear model}.
 \begin{prop}
 \label{prop:linear}
-  Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{R}(\beta_S, \beta_I)$.
-Then the coefficients $\widehat{\beta}_S$, $\widehat{\beta}_I $ and $\widehat{\rho}_{\text{LW}}$, $(1-\widehat{\rho}_{\text{LW}})\widehat{\mu}$ are very closed.
-
+Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{RE}(\beta_S, \beta_I)$.
+Then the estimator \eqref{linear model} can be expressed as
+$$\widehat{\bs{\Sigma}}_{\text{LW}} = (\widehat{\beta}_S)_+ \bs{S} + \min\{\widehat{\beta}_I, \hat{\mu}\} \bs{I}$$.
+where $\hat{\mu} = \frac{1}{p}\sum_{j=1}^p s_{jj}$
 \end{prop}
-
+The first coefficient, which is constraint to be nonnegative, is an analog of nonnegative James-Stein estimator.
 \subsection{\label{sec:proposed}Proposed estimator}
@@
 We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the priors $G_{nd}$ and $G_d$.
-Under the Bayesian model \eqref{bayesian}, and the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent across $jk$, we estimate $G_{nd}$ using
+Under the Bayesian model \eqref{bayesian nondiagonal} and \eqref{bayesian diagonal}, and the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent across $jk$, we estimate $G_{nd}$ using
 \begin{equation}
@@
 This setting was the same as Model 3 except with correlation parameter 0.9.
-This high level of dependence tested the robustness of the pairwise composite likelihood estimator \eqref{Gp hat}.
+This high level of dependence tested the robustness of the pairwise composite likelihood estimator \eqref{Gnd hat}.
 
 \item Model 5.
-With $\bs{U}$ a randomly generated orthogonal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4.
+With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4.
 This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
@@
 \item Model 6.
-With $\bs{U}$ a randomly generated orthogonal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p -3$ entries equaled 1.
+With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p -3$ entries equaled 1.
 \end{itemize}
@@
 \hline
-Exemplar method   & 0.0748          & 4.2575         & 47.8616       \\
-$K=2p$	      & 0.0445 	     & 1.3166	   & 12.2612         \\
-$K=p$            & 0.0291         & 0.8744         & 8.0862         \\
-$K=p/2$         & 0.0207         & 0.5861         & 5.3757         \\
-$K=p/4$      & 0.0153         &0.4132          & 3.7415         \\
+Exemplar method   & 0.1357          & 6.0338        & 91.8942       \\
+$K=2p$	      & 0.0539 	     & 1.1883	   & 10.7687         \\
+$K=p$            & 0.0357         & 0.7852         & 7.0807         \\
+$K=p/2$         & 0.0222        & 0.5231         & 4.6736         \\
+$K=p/4$      & 0.0158         &0.3694          & 3.2749         \\
 \hline
@@
 
-\item OracMSG: the optimal covariance estimator in the class of separable estimators $\mathcal{S}$ \eqref{separable}.
-It equals our proposed estimator \eqref{proposed} except with the true $G_p$ \eqref{bayesian} instead of $\widehat{G}_p$ \eqref{Gp hat}.
-The adaptive thresholding method of \citet{cai2011adaptive} also targets a separable estimator.
+\item OracMSG: It equals our proposed estimator \eqref{proposed} except sample grid points are cluster centers of true parameters $(\sigma_j,\sigma_k,r_{jk})$.
 \end{itemize}
@@
 Model 2 was especially interesting because the standard deviation and correlations were related.
-Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_{nd}$ \eqref{bayesian} and leverage it to provide much more accurate estimates.
+Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_{nd}$ \eqref{bayesian nondiagonal} and leverage it to provide much more accurate estimates.
 The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5.
@@
 \hline
-MSG   & 24.07          & 12.89         & 47.12         \\
-MSGCor & \textbf{23.90} & \textbf{12.87}& \textbf{46.63}         \\
-Adap        & 28.81          & 17.20         & 50.71         \\
-Linear      & 24.59         & 13.57         & 51.43         \\
-QIS         & 28.66          & 15.94         & 59.30         \\
-NERCOME     & 29.76          & 20.13         & 57.30         \\
-Sample      & 28.39          & 16.10         & 57.90         \\
+MSG   & 24.13 (2.54)         & 13.14(2.96)         & 46.53(8.49)         \\
+MSGCor & \textbf{24.04(2.56)} & \textbf{13.12(2.94)}& \textbf{46.00(8.67)}         \\
+Adap        & 28.93(2.03)          & 17.54(3.85)         & 49.66(8.47)         \\
+Linear      & 24.49(2.53)         & 13.65(3.14)         & 49.46(9.04)         \\
+QIS         & 28.76(2.79)          & 16.09(3.35)         & 58.73(9.74)         \\
+NERCOME     & 24.52(2.52)          & 13.19(3.07)         & 48.27(9.67)         \\
+Sample      & 28.43(2.54)          & 16.21(3.15)         & 56.08(9.89)         \\
 \hline
@@
 Finally, we have so far assumed that our data are multivariate normal.
-To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gp hat} and in our proposed estimator \eqref{proposed}.
+To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gnd hat} and in our proposed estimator \eqref{proposed}.
 If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.
@@
 \label{eq:lw}
-\bs{\Sigma}_{LW} = \frac{d_n^2-b_n^2}{d_n^2}\bs{S}+ \frac{b_n^2}{d_n^2}\hat{\mu}\bs{I}
+\bs{\Sigma}_{LW} = (\frac{d_n^2-b_n^2}{d_n^2})_+\bs{S}+ \min\{1,\frac{b_n^2}{d_n^2}\}\hat{\mu}\bs{I}
 \end{equation}
-where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{n^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.
+where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{pn^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.
 
-  We first rewrite the risk estimate $\widehat{R}(\beta_S, \beta_I)$.
+  We first rewrite the risk estimate $\widehat{RE}(\beta_S, \beta_I)$.
 Define $\bs{M}=(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$.
@@
   \[
- p^2 \widehat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
+ p^2 \widehat{RE}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
   \]
@@
   \begin{align*}
-    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\widehat{R}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
+    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\widehat{RE}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
     &
@@
 &=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)/p \} / d_n^2
-=\widehat{\mu}\frac{b_n^2}{d_n^2}\frac{n}{n-1}.
+=\widehat{\mu}\frac{b_n^2}{d_n^2}.
 \end{align*}
@@
 & =\frac{1}{p^2}\sum_{j,k=1}^p [(2\beta_{S}-1)\text{Var}(s_{jk})+\bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2] \\
-& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\widehat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2]
+& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\frac{n}{n-1}\widehat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2]
 \end{align*}
+The last equality is based on the fact that $\text{Var}(s_{jk}) = \frac{1}{n(n-1)}\sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2$.
 \end{proof}

### 2021-10-02 response_reviews.tex
add support and reponse to review

\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\def\bs{\boldsymbol}
\def\bb{\mathbb}

\title{}

\begin{document}

\section{Comments to the editor}
\begin{enumerate}
\item \emph{As a journal, Biometrics presents papers that focus on the development of new methods and results of use in biosciences. As of now, the application seems more an afterthought, the use of a convenient (easily available) dataset to fit the scope of the journal. I believe that the authors should motivate their approach from the gene expression data or other bio-related applications starting from the Introduction.}
\item \emph{It is not really intuitively clear why the vectorization works, despite the approach not taking into consideration the structure of the covariance matrix. Also, the manuscript proposes a series of ad-hoc expedients to implement the method. See the clustering-based exemplar algorithm and replacing non-positive eigenvalues with a chosen positive value $c$ in the positive-definite correction. These expedients do not seem to be completely justified by the proposed separable compound-decision theory framework, and their effects on the estimation are unclear. A referee also commented on the impossibility to follow closely the implementation of the EM algorithm. In short, while I appreciate that the theoretical justification for the asymptotic efficiency of the proposed vectorization is not clear, the current implementation reads more like a clever algorithm than a well-founded statistical method. The authors should justify all the steps of the algorithm and provide an assessment of their contribution to its overall performance.}

\item \emph{I agree with one of the referees that the title should be re-thought, since compound decision theory or compound decisions appear quite relevant to the approach, more than the empirical Bayes aspect of the paper.}\\
We agree that compound decision theory should be more emphasized than empirical Bayes in our paper. Following your suggestions, we have changed the title from "A Empirical Bayes Approach to Covariance Matrix Estimation" to "A Compound Decision Approach to Covariance Matrix Estimation".

\item \emph{on page 4, it appears to me that the choice of the risk function $\hat{R}$ does not correspond to the Frobenius risk. The rationale for choosing this risk function should be provided. It somehow appears the choice is dictated by Ledoit et al (2004) and Proposition 1, and not by independent first principles. If so, it should be explicitly stated before.}\\
We squared and scaled the Frobenius norm $\|A\|_F = \sqrt{\sum_{j,k=1}^pa_{jk}^2}$ to $\frac{1}{p^2}\sum_{j,k=1}^p a_{jk}^2$ in order to apply fundamental theorem of compound decisions, for which the risk function should be scaled by the number of parameters because the empirical distribution has mass probability of $\frac{1}{d}$ where $d$ is the number of parameters.\\
The transformation of Frobenius norm will not change the value of $(\widehat{\beta}_S,\widehat{\beta}_I)$ since it is determined by minimizing $RE(\beta_S,\beta_I)$.
\item \emph{Figures 1 and 2 do not allow to appreciate uncertainty. Perhaps a table would be better.}\\
We add the tables of the medians and uncertainty of Frobenius errors in supporting information. The uncertainty of Frobenius error is measured by the difference of 75\% quantile and 25\% quantile in 200 replicates. We use this range to measure the uncertainty because the errors of most methods are not following normal distribution. 
\item \emph{The manuscript requires careful proofreading and improvements. in the structure of the presentation and the visualization of the results. Non-exhaustive examples are reported below:}

\emph{- p.1 “To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrix” $\to$ matrices}

It has been fixed.

\emph{- p. 1 “about the structure of population covariance matrix” $\to$ of the population covariance matrix}

It has been fixed.

\emph{- p. 2 “Other common structured methods assume that the covariance matrix is banding” $\to$ banded?}

It has been edited to 'banded'.

\emph{- p. 2 “yet can still outperform the sample covariance matrix” What does this mean?}

It means that these unstructured model have been shown to have lower estimation risk comparing to sample covariance matrix. For example, in \cite{ledoit2004}, the numerical result shows that their linear shrinkage method has much lower risk than sample covariance matrix. To make it clear, we have changed to "yet still have less estimation error comparing to the sample covariance matrix".

\emph{- p. 3 “is squared error loss”$\to$ is the squared error loss}

It has been fixed.

\emph{- p. 4 “Now consider the problem of estimating the vectorized $\boldsymbol{\Sigma}$ under risk 1” $\to$ risk (1) Equations are typically referenced between brackets. The same issue appears in many other points where equations are referenced throughout the manuscript.}

We used to use \textbackslash ref to reference equation, we have changed it to \textbackslash eqref so that they are referenced between brackets.

\emph{- p. 8 “A major advantage of nonparametric estiation of the prior”$\to$ estimation}

It has been fixed.

\emph{- page 8, “Figure 1 shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table 1 shows that they can be significantly faster” However, at this point Figure 1 refers to models that have not yet been presented. Also, the reference to Table 1 anticipates point made in Section 3.2}

We have moved the setting of the experiment to the start of Section 3 Numerical Results.

\emph{- p. 16 “Finally, we have so far assumed that our data multivariate normal” $\to$ are normal}

It has been fixed.

\emph{- In Equations (1) and (2) the arguments of the risk functions are $\theta$ and $\delta$. However, on page 5 the arguments of $\hat{R}$ are $\beta_{S}$, and $\beta_{I}$.}

We have changed the notation $\hat{R}$ to $\text{RE}$ which means risk estimator and this differentiates it from $R$. 
\end{enumerate}

\section{Comments to Referee1}
Thank you so much for the detailed and inspiring suggestions on our paper. We learnt a lot through them. We agree with your comments and address them in the updated draft version. Please find the following responses for more details.
\subsection{comments to section "CorShrink"}
After carefully reviewed \cite{dey2018}, we think it is very related to our paper and it is valuable to make the comparison.

As mentioned in Section 1, our method is quite similar to \cite{dey2018}. Both of us used empirical Bayes estimator and nonparametric estimation of prior distribution. The difference is that our method aim to estimate covariance matrix which requires to also shrink standard deviations.

We also compared our two methods by numerical experiments which are shown in supporting information. We designed two simulations to compare the performance on correlation matrix estimation and covariance matrix estimation. In the first simulation, we compare covariance matrix estimation. For corShrink, $\widehat{\bs{\Sigma}} = \text{diag}(\widehat{SD}) \widehat{\bs{R}} \text{diag}(\widehat{SD})$, where $\widehat{SD}$ are sample standard deviations. In the second simulation, $\bs{\widehat{R}} = \text{diag}(1/\widehat{SD})\widehat{\bs{\Sigma}}\text{diag}(1/\widehat{SD})$, where $\widehat{SD}$ are standard deviations derived from $\widehat{\bs{\Sigma}}$. For each setting, we take $n$ as 20 and 100.

\subsection{comments to section "Separable rules, symmetry and the diagonal"}
\begin{equation}\label{R1}
\widehat{\Sigma} = (\widehat{\sigma}_{jk})_{1\leq j,k\leq p},\quad \widehat{\sigma}_{jk} = t(X_{\cdot,j}, X_{\cdot,k}), \quad t:\bb{R}^n\times \bb{R}^n\to \bb{R}
\end{equation}
\begin{enumerate}
\item \emph{While reading the paper I was a bit confused by the fact that (R1) treats diagonal entries and off-diagonal in the same way. It would feel more natural to me to define a separable estimator as in (R1) for $j\neq k$, and as follows on the diagonal:}
$$\hat{\sigma}_{jj} = \tilde{t}(X_{\cdot j}),\quad \tilde{t}:\bb{R}^n \to \bb{R}$$
We have changed the rule to 
\begin{equation}
  \label{separable}
  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
\end{equation}
\emph{In the Multivariate Gaussian case this is not an issue, since $X_{\cdot j}\neq X_{\cdot k}$ for $j\neq k$ with probability 1. But the proposed approach would also apply with minor modifications to e.g., discrete measurements (with known likelihood) in which case $X_{\cdot j}\neq X_{\cdot k}$ could happen with non-zero probability.} \\
We agree that it is not very precise to use $I(\bs{X}_{\cdot j}=\bs{X}_{\cdot k})$ to differentiate on-diagonal entries and off-diagonal entries. So we changed separable rule to the above class of rules.

\emph{Furthermore, while this issue is ignored for most of the paper, diagonal and off-diagonal entries are treated in a different way at the end of Section 2.3 (in the context of the clustering based algorithm). I believe it would be valuable to make the above distinction more explicit!}
\item \emph{A related question I had concerns the symmetry of the estimator. Does the estimator satisfy the symmetry property $\hat{\sigma}_{jk} =\hat{\sigma}_{kj}$? }

Yes, we only calculate estimates of lower triangular matrix of $\Sigma$, i.e., $\hat{\sigma}_{jk}$ where $j\geq k$. Then the upper triangular matrix estimation is calculated by the symmetry of covariance matrix $\sigma_{kj} = \sigma_{jk}$. Correspondingly, we should fix the Bayesian model in Section2.2 from
\begin{equation}
  \label{bayesian}
  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
to 
\begin{equation}
  \label{bayesian nondigaonal}
 (\bs{X}_{\cdot j},\bs{X}_{\cdot k})\mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
  \quad
  \bs{\eta} \sim G_{nd}(a,b, \gamma) =  \frac{2}{p(p-1)} \sum_{1\leq k< j \leq p} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
\end{equation}
and
\begin{equation}
  \label{bayesian nondigaonal}
  \bs{X}_{\cdot j} \mid a \sim \widetilde{f}(\cdot \mid a),
  \quad
  a \sim G_{d}(a) =  \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a).
\end{equation}
, as well as the proposed estimator from
\begin{equation}
  \label{Gp hat}
  \widehat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
\end{equation}
to
\begin{equation}
  \label{Gp hat revision}
  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}} \prod_{1\leq k< j \leq p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\prod_{j=1}^p \int \widetilde{f}(\bs{X}_{\cdot j} \mid a) dG_{d}(a),
\end{equation}
\emph{Similarly, the reader may wonder whether the estimation procedure enforces symmetry in the prior $G$ in (6) with respect to $\sigma_j,\sigma_k$.}

In correspondence, we modified our sample exemplar algorithm, we first clusters sample points $(s_j,s_k, r_{jk})$ of lower triangular part of sample covariance matrix and get the $K$ centroids $(a,b,\gamma)$, then find their symmetric points $(b,a,\gamma)$. We also updated EM algorithm to guarantee each $(a,b,\gamma)$ and $(b,a,\gamma)$ has the same weight. 
\item \emph{The paper emphasizes the Multivariate Gaussian problem. In that case, the sample covariance is sufficient, and so, it would seem more natural to me to define a separable rule, as one that satisfies,
$$\hat{\sigma}_{jk} = \bar{t}(X_{\cdot j}^TX_{\cdot k}),\quad \bar{t}: \bb{R}\to \bb{R}_{+}$$
instead of (R1). I am wondering why the authors chose to use (R1) instead.}

$X_{\cdot j}^TX_{\cdot k}$ is sufficient statistics of the single entry $\sigma_{jk}$. In fact, for $(\bs{X}_{\cdot,j}, \bs{X}_{\cdot,k})$, the sufficient statistics for $\eta$ is $\bs{S}_{jk} = \begin{pmatrix} s_{jj}&s_{jk} \\  s_{kj}&s_{kk} \end{pmatrix}$. We also tried density based on single entries $s_{jk}$ before but the performance is not as good as using $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$.
\item \emph{If I understand correctly, one of the benefits of estimator (3) is that it works without specific assumptions on the noise distribution. Instead the approach here appears to more heavily rely on a known likelihood (and specifically a multivariate Gaussian likelihood). It would be great if some simulations could be conducted under misspecification, wherein the data is not multivariate Gaussian, but the approach is applied assuming multivariate Gaussian data.}

This is a good point. To measure the performance of our method on nonGaussian distribution, we did some simulations on negative binomial distribution. We first generate independent distributed $Y$ and then let $X = L Y$, where L is the Cholesky decomposition of $\Sigma$. The simulation results are presented in supporting information.
\end{enumerate}

\subsection{comments to section "Compound decision theory or empirical Bayes?"}
\begin{enumerate}
\item \emph{The result following equation (6) could be presented separately as a Theorem (or Proposition). It should be highlighted that part of the statement is that the estimator in (R1) that optimizes (R2) depends on all the $\sigma_{j}$ and $\rho_{jk}$.}

We rewrote the equation after (6) as a separate proposition to show the equivalence between the Frobenius risk and Bayes risk.
\item \emph{The term “compound decision theory” or “compound decisions” etc. deserves to be in the title!}

Yes, we agree that "compound decision" is more deserved to be highlighted in the paper so we replaced "nonparametric empirical Bayes" with "compound decision" in the title.
\item \emph{On the other hand, I would downemphasize the empirical Bayes aspect of the paper (and my preference would be to even remove it from the title). To me, an empirical Bayes analysis (especially since here the goal is to perform well, in a frequentist sense, in terms of (R2)) has the following flavor: “Let us mimic what a Bayesian would do, when we do not know the true prior, and in a way that we can get frequentist guarantees as well”. However, I do not think a Bayesian would start approaching this problem by first positing (6) and then using (8)2. Instead a more natural Bayesian approach would be to set a prior for the full covariance matrix $\Sigma$, e.g., as in Berger et al. [2020] and references therein. I would restrict the discussion of empirical Bayes to just saying that model (6) appears formally when optimizing over separable rules (R1). To approximate the unknown optimal rule, one follows an empirical Bayesian “Generalized Maximum Likelihood” approach with respect to the pairwise composite likelihood.3}
\end{enumerate}

\subsection{comments to section "other remarks"}
\begin{enumerate}
\item \emph{It would be of great service to the community if the authors could provide code implementing the method and reproducing the results!}

We put our code of methods and simulation in supporting information.
\item \emph{Related to the above point, based on the description in the manuscript it would not be possible for me to fully reproduce the results. For example, does the actual implementation use the EM algorithm or an interior point convex solver? If the EM algorithm is used, how many steps are taken or what is the termination criterion? And why is the EM algorithm used (since for these empirical Bayes problem it typically converges extremely slowly compared to interior point solvers, as carefully demonstrated by Koenker and Mizera [2014])? On the other hand, if an interior point solver is used, which one?}

In our implementation, we used EM algorithm to estimate the prior distribution. We added the related details to Section 3 Numerical Results. The maximum number of iterations is set as 200. Iteration early stops when the criterion $\frac{L(\bs{w}^{(k+1)}) - L(\bs{w}^{(k)})}{L(\bs{w}^{(k)})}\leq 10^{-4}$ is met. We also tried interior point, but it always failed to find the optimal solution. The possible reason is the density $f(\cdot|a,b,\gamma)$ has very different scales, from $10^{-22}$ to $2$ and this makes interior point method difficult to solve.
\item \emph{First paragraph of page 3, “they modify only the sample eigenvalues and not the sample eigenvectors”: The situation here is a bit more nuanced, since e.g., sometimes using the sample eigenvectors is provably the optimal thing to do (e.g., if one constrains themselves to rotationally invariant estimators as in Bun et al. [2016] and references therein)}

The sentence "However, they modify only the sample eigenvalues and not the sample eigenvectors." is not very appropriate so we changed it to "However, as rotation-equivariant estimators (Stein, 1975, 1986), shrinkage estimators modify only the sample eigenvalues and keep the sample eigenvectors.".
\item \emph{Page 5, “it is straightforward to show that”: The “straightforward” could be replaced by a short proof in the appendix.}

We added the proof in the appendix.

\emph{Also could it be that the risk estimate is not exactly unbiased (because of dividing by $n$ and not $n-1$ in the definition of $\hat{\Delta}_{jk}^2$)?}
You're right. We realized that the previous defined $\hat{\Delta}_{jk}^2$ is not unbiased estimator of Frobenius risk of $\bs{\delta}$ (1). So we changed the text to describe it as an "approximately unbiased risk estimator" since the factor $\frac{n-1}{n}$ is closed to 1. It is also shown in appendix B.

\emph{Where is the sample covariance $\bs{S}$ defined?}

In the updated manuscript, $\bs{S}$ is defined as $\bs{S} = \frac{1}{n}\bs{X}^\top \bs{X}$ as it is first mentioned in page 4, section 2.1.

\item \emph{Section 2.5, “Projections in terms of other matrix norms are also possible”. Is this actually being done in the implementation?}

No. We did not implement it in terms of other norms.

\emph{Else perhaps remove this sentence.}

It has been removed.
\end{enumerate}
\section{Comments to Referee2}
\begin{enumerate}
\item \emph{In the numerical experiments and the real data analysis, what was the size of the grid that was used? I may have missed it but I did not see it mentioned in the caption for figure 2.}\\
In section 3.3, methods compared and section 3.4, data analysis, we used clustering-based exemplar algorithm, where we find $K$ clusters of all off-diagonal sample points. $K$ is set to be $p$ to balance accuracy and speed according to result of section 3.2.
\item \emph{In the numerical experiments, the dimension p goes as far as 200. I believe it will be useful to consider a setting where $n = 100$ and $p = 1000$.} \\
We did the simulation for $n=100$ and $p=1000$ using the different covariance matrix models in Section 3.1. Please see the supporting information for the simulation results.  
\item \emph{Continuing on the numerical experiments, models 1 to 5 impose various structures on $\bs{\Sigma}$. It will be interesting to see the performance of the proposed method if a spiked covariance structure is imposed on $\bs{\Sigma}$.} \\
We added the spiked model as the sixth matrix models in Section 3.1. Spiked covariance matrix is generated by $\Sigma = \bs{U}\text{diag}(\bs{l})\bs{U}^\top$, where $\bs{U}$ is a randomly generated orthogonal unit matrix, $\bs{l} = \{4,3,2,1,\ldots,1\}$.
\item \emph{Typo on page 6: "The density of $f(\cdot \mid \bs{\eta}_{jk})$ of ...".}\\
Thanks for poiting this out. We've fixed this sentence as "$f(\cdot \mid \bs{\eta}_{jk})$, the density of $\bs{A}_{jk}$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, ...".
\end{enumerate}
\begin{thebibliography}{9}
\bibitem{dey2018}
Dey, K. K., \& Stephens, M. (2018). CorShrink: Empirical Bayes shrinkage estimation of correlations, with applications. bioRxiv, 368316.
\bibitem{ledoit2004}
Ledoit, O., \& Wolf, M. (2004). A well-conditioned estimator for large-dimensional covariance matrices. Journal of multivariate analysis, 88(2), 365-411.

\end{thebibliography}
\end{document}

### 2021-11-02 response_reviews.tex
edited intro and compound decision formulation

@@
 \documentclass[12pt]{article}
+\usepackage[margin=1in]{geometry}
 \usepackage{amsmath}
@@
 
-\title{}
+\title{Response to reviews of BIOM2021236M}
+\author{}
+\date{}
 
 \begin{document}
+\maketitle
 
-\section{Comments to the editor}
+\section{Response to the Associate Editor}
+
 \begin{enumerate}
@@
 Please find the following responses for more details.
-\subsection{comments to section "CorShrink"}
+
+\subsection{Comments to section "CorShrink"}
 After carefully reviewed \cite{dey2018}, we think it is very related to our paper and it is valuable to make the comparison.
@@
 
-\subsection{comments to section "Separable rules, symmetry and the diagonal"}
+\subsection{Comments to section "Separable rules, symmetry and the diagonal"}
 \begin{equation}\label{R1}
@@
 
-\subsection{comments to section "Compound decision theory or empirical Bayes?"}
+\subsection{Comments to section "Compound decision theory or empirical Bayes?"}
 \begin{enumerate}
@@
 \end{document}
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2021-11-02 msg.tex
edited intro and compound decision formulation

@@
 \usepackage{graphicx}
-%\usepackage{subcaption}
+% \usepackage{subcaption}
+\usepackage{xcolor}
 
@@
 
-\title[Empirical Bayes Covariance Matrix Estimation]{A Compound Decision Approach to Covariance Matrix Estimation}
+\def\blue{\textcolor{blue}}
+
+\title[Coumpound Decision Covariance Matrix Estimation]{A Compound Decision Approach to Covariance Matrix Estimation}
 
@@
 \begin{abstract}
-  We propose an empirical Bayes method to estimate high-dimensional covariance matrices.
-Our procedure centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem.
-Drawing from the compound decision theory literature, we introduce a new class of decision rules that generalizes several existing procedures.
-We then use a nonparametric empirical Bayes $g$-modeling approach to estimate the oracle optimal rule in that class.
-This allows us to let the data itself determine how best to shrink the estimator, rather than shrinking in a pre-determined direction such as toward a diagonal matrix.
-Simulation results and a gene expression network analysis shows that our approach can outperform a number of state-of-the-art proposals in a wide range of settings, sometimes substantially.
+  Covariance matrix estimation is a fundamental statistical task in many applications, but the sample covariance matrix is sub-optimal when the sample size is comparable to or less than the number of features.
+Such high-dimensional settings are common in modern genomics, where covariance matrix estimation is frequently employed as a method for inferring gene networks.
+To achieve estimation accuracy in these settings, existing methods typically either assume that the population covariance matrix has some particular structure, for example sparsity, or apply shrinkage to better estimate the population eigenvalues.
+In this paper, we propose a new approach to estimating high-dimensional covariance matrices that can have better performance than existing methods.
+We first frame covariance matrix estimation as a compound decision problem.
+This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class.
+Simulation results and gene network inference in a single cell RNA-seq dataset show that our approach is comparable to or can outperform a number of state-of-the-art proposals in a wide range of settings.
 \end{abstract}
@@
 
-Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications, such as portfolio management \citep{ledoit2003improved}, genomics \citep{schafer2005shrinkage}, and array signal processing \citep{abramovich2001locally}.
-However, in modern problems the number of features can be of the same order as or exceed the sample size, and the standard sample covariance matrix estimator behaves poorly in this regime.
-To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrces.
+Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications.
+However, in modern problems the number of features can be of the same order as or exceed the sample size.
+This high-dimensional setting is especially common in genomics, where covariance matrices are used to model gene networks but the number of genes can be much larger than the number of biological replicates \citep{schafer2005shrinkage, markowetz2007inferring}.
+Accurate gene network estimation has become even more important with the advent of single-cell RNA-sequencing (scRNA-seq) technology, where interest now centers on inferring cell- or cell type-specific gene networks \citep{aibar2017scenic, dai2019cell, stegle2015computational}.
+For example, in Section \ref{gene analysis} we study cell type-specific gene network inference using scRNA-seq data from honey bee brains.
+\blue{Need to do analysis}
+
+A common approach to estimating gene expression networks is to use the standard sample covariance or corrlation matrices \citep{langfelder2008wgcna, zhang2005general}.
+However, these estimators behave poorly in the high-dimensional regime.
+To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrices.
 These can roughly be divided into two groups, according to whether they impose assumptions of the structure of population covariance matrix.
 
-Structured methods make structural assumptions about the population covariance matrix.
-One class models the population covariance matrix as sparse.
-The most common method to address this problem is thresholding \citep{rothman2009generalized, cai2011adaptive}.
-Penalized likelihood methods \citep{xue2012positive} can also estimate large-scale sparse covariance matrix by penalizing a log-likelihood function.
-Another class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension.
-Other common structured methods assume that the covariance matrix is banded \citep{li2017estimation} or Toeplitz matrix \citep{liu2017covariance}.
+Structured methods make assumptions about the form of the population covariance matrix.
+One popular class of methods assumes that the matrix is sparse, or has many zero entries.
+The most common strategy in this class is to threshold the entries of the sample covariance \citep{rothman2009generalized, cai2011adaptive}, but penalized likelihood methods \citep{xue2012positive} have also been used.
+A second class of methods assume the data arise from a factor model \citep{fan2008high}, so that the covariance matrix has low intrinsic dimension.
+Other common structured methods assume that the covariance matrix is banded \citep{li2017estimation} or Toeplitz \citep{liu2017covariance}.
 
-In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have less estimation error comparing to the sample covariance matrix.
+In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have lower estimation error than the sample covariance matrix.
 A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix.
 More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}.
-These shrink the eigenvalues of the sample covariance matrix toward clusters.
-Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global mean.
+These shrink the eigenvalues of the sample covariance matrix in a data-adaptive fashion.
+Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global means.
 
 Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance.
-However, as the class of rotation-equivariant estimators \citep{stein1975estimation, stein1986lectures}, nonlinear shrinkage estimators modify only the sample eigenvalues and keep the sample eigenvectors.
-It is known that sample eigenvectors are not consistent estimators of population eigenvectors when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}.
-This suggests that there may exist a class of unstructured estimators that can outperform nonlinear shrinkage.
+However, most existing nonlinear shrinkage estimators belong to the class of rotation-equivariant estimators \citep{bun2016rotational, stein1975estimation, stein1986lectures}, which modify the sample eigenvalue estimates but still use the sample eigenvectors.
+On the other hand, sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}.
+This suggests that certain classes of non-rotationally invariant estimators may outperform existing unstructured covariance matrix estimators.
 
-%In this paper, we propose a new covariance matrix estimator which is approximately optimal among all separable estimators.
+% In this paper, we propose a new covariance matrix estimator which is approximately optimal among all separable estimators.
 Our method is motivated by the similarity between covariance estimation and vector mean estimation.
@@
 
-Here we propose a new unstructured estimator for high-dimensional covariance matrices.
-Our approach centers on vectorizing the covariance matrix and treating matrix estimation as a vector estimation problem.
-We do this because it allows us to use a nonparametric empirical Bayes shrinkage procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex}.
+Here we propose a new class of unstructured non-rotationally invariant estimators for high-dimensional covariance matrices.
+Our approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}; see Section \ref{sec:compound}.
+This framing motivates us to vectorize the covariance matrix and solve the resulting vector estimation problem using a nonparametric empirical Bayes procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex, efron2019bayes}.
 We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator.
 Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis.
-Our approach is similar to \citet{dey2018corshrink}, which also applied empirical Bayes shrinkage method to estimate correlation matrix.
-Different from our method, their corShrink method has different target and they do not shrink standard deviations.
+Our compound decision approach is conceptually similar the empirical Bayes approach taken by \citet{dey2018corshrink} to estimate a high-dimensional correlation matrix.
+We discuss in detail the connections with their work in Section \ref{sec:proposed}.
+\blue{Need to add to numerical results}
 
@@
 Our procedure is implemented in the R package \verb|cole|, available on GitHub.
+\blue{Is this implemented in the package yet?}
 
-\section{\label{sec:method}Method}
-\subsection{\label{sec:compound}Compound decision problem formulation}
-Suppose we have $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$.
-The purpose of this paper is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$ that minimizes the scaled squared Frobenius risk
+\section{\label{sec:compound}Compound decision problem formulation}
+\subsection{\label{sec:background}Background}
+
+Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$.
+A common measure of estimation performance is the scaled squared Frobenius risk
 \begin{equation}
@@
 where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.
+This paper constructs a new non-rotationally invariant $\bs{\delta}$ that in simulations and a data analysis has comparable or lower risk \eqref{frobenius risk} compared to existing methods.
 
-Our proposed approach is motivated by two observations.
-First, \eqref{frobenius risk} shows that estimating $\bs{\Sigma}$ under Frobenius risk is equivalent to simultaneously estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under a loss function that aggregates errors across components.
-Second, this type of vector estimation problem has been well-studied in the compound decision literature.
-Thus, recent advances in vector estimation may be profitably applied to covariance matrices.
-
-We first briefly review compound decision problems.
-Introduced by \citet{robbins1951asymptotically}, these problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$.
+Our proposed approach is motivated by the observation that minimizing \eqref{frobenius risk} is a type of compound decision problem, which we briefly review in this section.
+Introduced by \citet{robbins1951asymptotically}, compound decision problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$.
 Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk
@@
 
-A key property of compound decision problems is that while a given $Y_i$ seems to offer no information about any specific $\theta_j$ when $j \ne i$, borrowing information across all components of $\bs{Y}$ to estimate $\bs{\theta}$ is superior to estimating each $\theta_i$ using the corresponding $Y_i$ alone.
+A key property of compound decision problems is that while a given $Y_i$ or $\theta_i$ seems like it should offer no help for estimating $\theta_j$ when $j \ne i$, in fact borrowing information across all indices $i = 1, \ldots, n$ to estimate $\theta_j$ is superior to estimating each $\theta_j$ using the corresponding $Y_j$ alone.
 A classical example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$.
-It is well-known that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$.
+\citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$.
 A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.
 
-We propose to apply some of these recent vector estimation ideas to covariance matrix estimation.
-Clearly, covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can be viewed as a compound decision problem.
-Furthermore, some existing covariance matrix estimation procedures can already be interpreted as taking a vector approach.
-The sample covariance matrix $\bs{S}=\frac{1}{n}\bs{X}^{T}\bs{X}$, for instance, can be thought of as estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$ using maximum likelihood.
+The discussion above shows that covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can be viewed as a compound decision problem.
+Furthermore, one special consequence of choosing \eqref{frobenius risk} as the risk measure is that the matrix estimation problem becomes equivalent to a vector estimation problem, specifically the problem of estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under average squared error loss.
+This interpretation is conceptually very similar to the classical Gaussian sequence problem.
+
+We therefore propose to apply modern ideas for Gaussian sequence estimation to covariance matrix estimation.
+One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite.
+Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}.
+As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive semidefinite.
+This will in fact be true of one of the estimators we propose in Section \ref{sec:proposed}.
+To resolve this issue, we project the result estimate into the space of positive-definite matrices; see Section \ref{posdef}.
+
+\subsection{\label{sec:connections}Connections to existing work}
+
+Treating the covariance matrix estimation problem as a vector estimation compound decision problem may seem unintuitive, but many existing matrix estimation methods can also be interpreted as carrying out vector estimation.
+The simplest example is the sample covariance matrix $\bs{S} = \frac{1}{n}\bs{X}^{T}\bs{X}$, which takes this form because the $\bs{X}$ are assumed to have mean zero.
+This can be thought of as estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$ using maximum likelihood.
 Less trivially, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective.
-Their adaptive thresholding method is a version of the soft thresholding method of \citet{donoho1995adapting}, which was originally developed to estimate a sparse mean vector in the Gaussian sequence problem.
+Their adaptive thresholding method applies a version of the soft thresholding method of \citet{donoho1995adapting}, originally developed to estimate a sparse mean vector in the Gaussian sequence problem, to each entry of the sample covariance matrix.
 
-Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be interpreted as a vector estimator. The estimator is defined as
+Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as a solution to a vector estimation problem where the estimand is the $p^2 \times 1$ vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$.
+Their estimator is defined as
 \begin{equation}
 \label{linear model}
-\widehat{\bs{\Sigma}}_{\text{LW}} = \widehat{\rho}_{\text{LW}} \bs{S} + (1-\widehat{\rho}_{\text{LW}})\widehat{\mu} \bs{I}
+\hat{\bs{\Sigma}}_{\text{LW}} = \left(1 - \frac{b_n^2}{d_n^2}\right) \bs{S} + \frac{b_n^2}{d_n^2}\hat{\mu} \bs{I}
 \end{equation} 
-where $\widehat{\mu} = \text{tr}(\bs{S})/p$, and $\widehat{\rho}_{\text{LW}} = 1 - b_n^2 / d_n^2$ for $d_n^2=\|\bs{S}-\widehat{\mu} \bs{I} \|_F^2$ and $b_n^2 = \sum_{i=1}^n \| \bs{X}_i \bs{X}_i^\top - \bs{S} \|_F^2 / n^2$.
-Now consider the problem of estimating the vectorized $\bs{\Sigma}$ under risk \eqref{frobenius risk}.
-We restrict attention to decision rules that estimate each component $\sigma_{jk}$ using
+for $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, and $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$, where $\Vert \cdot \Vert_F$ denotes the Frobenius norm.
+
+To see this, we first restrict attention to the class of linear decision rules for estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$, which we define to take the form
 \begin{equation}
   \label{linear class}
-  \beta_S s_{jk} + \beta_I u_{jk},
+  \delta_{jk}(\bs{X}) = \beta_S s_{jk} + \beta_I u_{jk}
 \end{equation}
-where $s_{jk}$ is the $jk$th entry of $\bs{S}$, $u_{jk}$ is the $jk$th entry of $\bs{I}$, and the class is indexed by the parameters $(\beta_S, \beta_I)$. As proved in appendix, we can show that
-$$\textsc{RE}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \widehat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ]$$
-is an approximately unbiased estimator of the risk \eqref{frobenius risk}, where $\widehat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ is an approximately unbiased estimator of variance of $s_{jk}$.
-The optimal estimator in this class can now be chosen by minimizing $RE(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
-It can be shown that this is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion}.
-Proposition \eqref{prop:linear} shows that this subspace shrinkage estimator is closely related to the \citet{ledoit2004well} estimator \eqref{linear model}.
+for some parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of $\bs{I}$.
+In other words, the estimate for $\sigma_{jk}$ is linear in $s_{jk}$.
+Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}, and the following proposition shows that it is possible to construct a good estimate of this risk for every fixed value of $\beta_S$ and $\beta_I$.
+
+\begin{prop}
+  \label{prop:Rhat}
+  Define $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ and
+  \begin{equation}
+    \label{eq:Rhat}
+    \hat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ].
+  \end{equation}
+  If $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$, then
+  \[
+    \lim_{n \rightarrow 0} \{ \bb{E} \hat{R}(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta}) \} = 0
+  \]
+  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}.
+\end{prop}
+The condition on $\bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ requires that the variances of the entries of $\bs{S}$ do not grow too quickly as $n$ grows.
+It is also required by \citet{ledoit2004well} and is implied by their Lemma 3.1.
+
+Proposition \ref{prop:Rhat} shows that $\hat{R}(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}.
+It is then reasonable to estimate the best estimator in this class by minimizing $\hat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
+Proposition \ref{prop:linear} shows that this estimator is closely related to the \citet{ledoit2004well} estimator \eqref{linear model}.
+
 \begin{prop}
 \label{prop:linear}
-Define the estimator $\widehat{\bs{\Sigma}}_V$ such that its $jk$th entry obeys $\widehat{\sigma}_{jk} = \widehat{\beta}_S s_{jk} + \widehat{\beta}_I u_{jk}$, where $(\widehat{\beta}_S, \widehat{\beta}_I) = \argmin_{\beta_S, \beta_I} \widehat{RE}(\beta_S, \beta_I)$.
-Then the estimator \eqref{linear model} can be expressed as
-$$\widehat{\bs{\Sigma}}_{\text{LW}} = (\widehat{\beta}_S)_+ \bs{S} + \min\{\widehat{\beta}_I, \hat{\mu}\} \bs{I}$$.
-where $\hat{\mu} = \frac{1}{p}\sum_{j=1}^p s_{jj}$
+Let $\hat{\beta}_S$ and $\hat{\beta}_I$ denote the minimizers of $\hat{R}(\beta_S, \beta_I)$ and define the linear decision rule
+\begin{equation}
+  \label{eq:opt_linear}
+  \hat{\delta}_{jk}(\bs{X}) = \max(\hat{\beta}_S, 0) s_{jk} + \min\{\hat{\beta}_I, \hat{\mu}\} u_{jk}.
+\end{equation}
+Then $\hat{\delta}_{jk}(\bs{X}) = \hat{\sigma}_{\text{LW}jk}$, where $\hat{\sigma}_{\text{LW}jk}$ is the $jk$th entry of $\bs{\Sigma}_{\text{LW}}$ \eqref{linear model}.
 \end{prop}
-The first coefficient, which is constraint to be nonnegative, is an analog of nonnegative James-Stein estimator.
+
+Proposition \ref{prop:linear} therefore shows that the \citet{ledoit2004well} covariance matrix estimator can essentially be obtained as the solution to a pure vector estimation problem, without considering the matrix structure of the target vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$.
+There are interesting connections between the estimator in Proposition \ref{prop:linear} and classical estimators of a vector of normal means.
+First, instead of \eqref{eq:opt_linear}, consider the estimator $\tilde{\delta}_{jk}(\bs{X}) = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, which directly plugs the estimated $\hat{\beta}_S$ and $\hat{\beta}_I$ into \eqref{linear class}.
+It can be shown that  $\tilde{\delta}_{jk}(\bs{X})$ is analogous to the Efron-Morris estimator for the Gaussian sequence problem \citep{efron1973stein}.
+In other words, it can be shown that $\tilde{\delta}(\bs{X})$ is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion, stigler19901988}.
+A major difference between $\hat{\delta}_{jk}$ \eqref{eq:opt_linear} and $\tilde{\delta}_{jk}$ is that $\hat{\beta}_S$ in the latter is replaced by $\max(\hat{\beta}_S, 0)$ in the former.
+This is equivalent to how the shrinkage factor in the classical James-Stein estimator is replaced by its positive part in the positive-part James-Stein estimator \citep{baranchik1964multiple}.
+
+\section{\label{sec:method}Method}
 \subsection{\label{sec:proposed}Proposed estimator}
 
-The previous section argues that treating covariance matrix estimation as a vector estimation problem can be a fruitful strategy, but discusses only estimators linear in $s_{jk}$.
-We propose to consider a larger class, the class of so-called separable rules.
-In the standard compound decision problem of estimating $\bs{\theta}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ \citep{robbins1951asymptotically}.
-Here we generalize this to the problem of estimating a vectorized matrix.
-For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, define the class of separable rules
+Section \ref{sec:compound} argues that treating covariance matrix estimation as a vector estimation problem can be a fruitful strategy, as evidenced by the linear estimator exhibited in Proposition \ref{prop:linear}.
+This suggests that estimators that are nonlinear in $s_{jk}$ may have better performance.
+
+We propose to consider an even larger class of estimators, the class of so-called separable rules.
+In the standard compound decision problem of estimating $\bs{\theta}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ for some function $t$ that does not depend on the index $i$ \citep{robbins1951asymptotically}.
+Here we generalize this idea to the problem of estimating a vectorized matrix.
+For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, we define the class of separable rules to be
 \begin{equation}
@@
 where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature.
-In other words, rules in $\mathcal{S}$ estimate the $jk$th entry of the covariance matrix using a fixed function $t$ of only observations from features $j$ and $k$.
-The sufficient statistics of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is $2\times 2$ matrix of $S$ denoted by $\bs{S}_{jk}=
-\begin{bmatrix}
-s_{jj} & s_{jk} \\
-s_{kj} & s_{kk} 
-\end{bmatrix}$.
+In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $\tilde{t}$ observations of the corresponding feature and off-diagonal entries using a function $t$ of observations from the two corresponding features, where $t$ and $\tilde{t}$ do not depend on the indices $j$ and $k$.
+Furthermore, we enforce rule in $\mathcal{S}$ to give symmetric estimates of the off-diagonal entries.
+
+% The sufficient statistics of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is $2\times 2$ matrix of $S$ denoted by $\bs{S}_{jk}=
+% \begin{bmatrix}
+% s_{jj} & s_{jk} \\
+% s_{kj} & s_{kk} 
+% \end{bmatrix}$.
+
+This class of separable rules is reasonable to consider because it includes several common covariance estimators, including the sample covariance $(s_{11}, \ldots, s_{pp})^\top$, the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}, and the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
+\begin{align*}
+t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}) &= \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n, \, 1\leq k<j\leq p,\\
+\tilde{t}(\bs{X}_{\cdot j})&=\beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot j} / n+\beta_I, \, j=1,\ldots,p.
+\end{align*}
+Furthermore, whereas these three existing estimators are ultimately all functions of the $s_{jk}$, the class $\mathcal{S}$ allows for much more general rules that can be any function of the observations $\bs{X}_{\cdot j}$ and $\bs{X}_{\cdot k}$.
 
 We propose to search for the optimal estimator within $\mathcal{S}$.
-This is sensible because $\mathcal{S}$ includes the sample covariance $(s_{11}, \ldots, s_{pp})^\top$, the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
-\begin{align*}
-\delta_{jk}(\bs{X})&= \delta_{kj}(\bs{X})= \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n, \, 1\leq k<j\leq p\\
-\delta_{jj}(\bs{X})&=\beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot j} / n+\beta_I, \, j=1,\ldots,p
-\end{align*}
-and the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}.
-Therefore the optimal separable estimator $\bs{\delta}^\star$ that minimizes the scaled squared Frobenius risk \eqref{frobenius risk} will perform at least as well as these three estimators, and may perform better.
+The optimal separable estimator $\bs{\delta}^\star$ that minimizes the scaled squared Frobenius risk \eqref{frobenius risk} over all rules in $\mathcal{S}$ will perform at least as well as the three estimators mentioned above, and may perform better.
 Targeting the optimal separable rule is standard in the compound decision literature \citep{zhang2003compound}.
@@
   \label{Gnd hat}
-  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}_{nd}} \{\prod_{1\leq k\leq j<p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\} 
+  \hat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}_{nd}} \{\prod_{1\leq k\leq j<p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\} 
   \{\prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG_d(a)\},
@@
 where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $G_d$ is determined by $G_{nd}$ as indicated in \eqref{eq:marginal}.
-Of course, the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are not independent, so $\widehat{G}_{nd}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
-Using $\widehat{G}_{nd}$ and $\widehat{G}_d$, we estimate the vectorized $\bs{\Sigma}$ using
-$$ \widehat{\bs{\delta}}(\bs{X})=
-  (\widehat{t}(\bs{X}_{\cdot 2},\bs{X}_{\cdot 1}), \ldots,  \widehat{t}(\bs{X}_{\cdot p},\bs{X}_{\cdot (p-1)}),\widehat{t}(\bs{X}_{\cdot 1}),\ldots,\widehat{t}(\bs{X}_{\cdot p}))$$
+Of course, the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are not independent, so $\hat{G}_{nd}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
+Using $\hat{G}_{nd}$ and $\hat{G}_d$, we estimate the vectorized $\bs{\Sigma}$ using
+$$ \hat{\bs{\delta}}(\bs{X})=
+  (\hat{t}(\bs{X}_{\cdot 2},\bs{X}_{\cdot 1}), \ldots,  \hat{t}(\bs{X}_{\cdot p},\bs{X}_{\cdot (p-1)}),\hat{t}(\bs{X}_{\cdot 1}),\ldots,\hat{t}(\bs{X}_{\cdot p}))$$
   , where 
@@
   \label{proposed}
-  \widehat{t}(\bs{X}_{\cdot j},\bs{X}_{\cdot k}) = \frac{\int g(\bs{\eta}) f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta})d\widehat{G}_{nd}(\bs{\eta})}{\int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) d\widehat{G}_{nd}(\bs{\eta})},
+  \hat{t}(\bs{X}_{\cdot j},\bs{X}_{\cdot k}) = \frac{\int g(\bs{\eta}) f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta})d\hat{G}_{nd}(\bs{\eta})}{\int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) d\hat{G}_{nd}(\bs{\eta})},
   \quad 
-  \widehat{t}(\bs{X}_{\cdot j}) = \frac{\int a^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a)d\widehat{G}_d(a)}{\int \widetilde{f}(\bs{X}_{\cdot j} \mid a) d\widehat{G}_d(a)}.
+  \hat{t}(\bs{X}_{\cdot j}) = \frac{\int a^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a)d\hat{G}_d(a)}{\int \widetilde{f}(\bs{X}_{\cdot j} \mid a) d\hat{G}_d(a)}.
 \end{equation}
-The $\widehat{t}$ estimates the Bayes rule $t^\star$, $\widetilde{t}^\star$ and $\widehat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.
+The $\hat{t}$ estimates the Bayes rule $t^\star$, $\widetilde{t}^\star$ and $\hat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.
 
@@
 In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}.
-Theoretical justification of our proposed $\widehat{\bs{\delta}}$ is difficult and is discussed in Section \eqref{discussion}.
-Nevertheless, our numerical results in Section \eqref{numerical results} show that in practice, our $\widehat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.
+Theoretical justification of our proposed $\hat{\bs{\delta}}$ is difficult and is discussed in Section \eqref{discussion}.
+Nevertheless, our numerical results in Section \eqref{numerical results} show that in practice, our $\hat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.
 
@@
 
-Calculating the estimated prior $\widehat{G}_{nd}$ \eqref{Gnd hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
+Calculating the estimated prior $\hat{G}_{nd}$ \eqref{Gnd hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
 \citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points.
@@
 Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$.
-We can then use the EM algorithm to estimate the masses $\widehat{\bs{w}}=\{\widehat{w}_{1},\ldots, \widehat{w}_{D}\}$ at those points via the iteration
+We can then use the EM algorithm to estimate the masses $\hat{\bs{w}}=\{\hat{w}_{1},\ldots, \hat{w}_{D}\}$ at those points via the iteration
 \begin{equation}
 \label{eq:iteration}
-\widehat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\widehat{w}_{\tau}^{(k-1)} f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\widehat{w}_l^{(k-1)} f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)}
+\hat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)}
 \end{equation}
@@
 \[
-\widehat{t}(\bs{X}_j,\bs{X}_k) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_\tau) \widehat{w}_{\tau}},
+\hat{t}(\bs{X}_j,\bs{X}_k) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_\tau) \hat{w}_{\tau}},
 \quad
-\widehat{t}(\bs{X}_{\cdot j}) = \frac{ \sum_{\tau=1}^D a_{\tau}^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a_\tau) \widehat{w}_{\tau}}{ \sum_{\tau=1}^D \widetilde{f}(\bs{X}_{\cdot j}  \mid a_\tau) \widehat{w}_{\tau}}.
+\hat{t}(\bs{X}_{\cdot j}) = \frac{ \sum_{\tau=1}^D a_{\tau}^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D \widetilde{f}(\bs{X}_{\cdot j}  \mid a_\tau) \hat{w}_{\tau}}.
 \]
@@
 However, the fact that $G_{nd}$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$.
-Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\widehat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$.
+Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\hat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$.
 This reduces the size of the support set, but even in this case the computation complexity grows like $O(p^2)$.
@@
 Although $\bb{E}\bs{X}=0$ in our setting, we assume it is unknown and use $\bs{S}=\frac{1}{n-1}(\bs{X}-\bs{\bar{X}})^\top(\bs{X}-\bs{\bar{X}})$.
-We generated $200$ replicates and reported median errors under the following Frobenius norms, where $\widehat{\bs{\Sigma}}$ is the estimated matrix with entries $\widehat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
-$$\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\widehat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$$
+We generated $200$ replicates and reported median errors under the following Frobenius norms, where $\hat{\bs{\Sigma}}$ is the estimated matrix with entries $\hat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
+$$\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\hat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$$
 which is a version of \eqref{frobenius risk}.
-%\item Spectral: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\widehat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\widehat{\bs{\Sigma}} - \bs{\Sigma}$, and
+%\item Spectral: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\hat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\hat{\bs{\Sigma}} - \bs{\Sigma}$, and
 %
-%\item Matrix $\ell_1$: $\Vert \widehat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \widehat{\sigma}_{jk} - \sigma_{jk} \vert$.
+%\item Matrix $\ell_1$: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \hat{\sigma}_{jk} - \sigma_{jk} \vert$.
 
@@
 \begin{equation}
-F = \frac{1}{k-1}\sum_{m=1}^kn_m(\overline{x}_m - \overline{x})^2 /   \frac{1}{n-k}\sum_{m=1}^k (n_m-1)\widehat{\sigma}_m^2 
+F = \frac{1}{k-1}\sum_{m=1}^kn_m(\overline{x}_m - \overline{x})^2 /   \frac{1}{n-k}\sum_{m=1}^k (n_m-1)\hat{\sigma}_m^2 
 \end{equation}
-where $k = 4$ is the number of patient categories, $n_m$, $\overline{x}_m$, and $\widehat{\sigma}_k$ represent the sample size, sample mean, and sample variance of the gene's expression in the $m$th category, respectively, and $\overline{x}$ is the global mean.
+where $k = 4$ is the number of patient categories, $n_m$, $\overline{x}_m$, and $\hat{\sigma}_k$ represent the sample size, sample mean, and sample variance of the gene's expression in the $m$th category, respectively, and $\overline{x}$ is the global mean.
 We proceeded with the top 40 genes and bottom 160 genes.
@@
 \section{}
+
+\subsection{Proof of Proposition \eqref{prop:Rhat}}
+\begin{proof}
+Using the fact that when $\bb{E}\bs{X}=\bs{0}$, $\bb{E}s_{jk} = \sigma_{jk}$ for all $j,k=1,\ldots,p$.
+Therefore for the class of linear decision rules \eqref{linear class}, the scaled Frobenius risk \eqref{frobenius risk} equals
+\begin{align*}
+R(\bs{\Sigma}, \bs{\delta}) &=\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} \{ (\beta_{S}s_{jk}+\beta_Iu_{jk}-\sigma_{jk})^2 \} \\
+& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} [s_{jk}-\sigma_{jk} - \{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}]^2\\
+& =\frac{1}{p^2}\sum_{j,k=1}^p [\bb{E}\{(s_{jk}-\sigma_{jk})^2\}+ \bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2-2(1-\beta_{S})\bb{E}\{s_{jk}(s_{jk}-\sigma_{jk})\}]\\
+& =\frac{1}{p^2}\sum_{j,k=1}^p [(2\beta_{S}-1)\text{Var}(s_{jk})+\bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2] \\
+& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\frac{n}{n-1}\hat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2],
+\end{align*}
+with $\hat{\Delta}_{jk}$ defined in Proposition \ref{prop:linear}. Therefore
+\begin{align*}
+  \bb{E} \hat{R}(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta})
+  =\,&
+       \frac{-1}{n - 1} (2\beta_S-1) \frac{1}{p^2} \sum_{j,k=1}^{p}\bb{E}\hat{\Delta}_{jk}^2
+       =
+       \frac{-1}{n} (2\beta_S-1) \frac{1}{p^2} \sum_{j,k=1}^{p} \text{Var}(s_{jk}) \\
+  =\,&
+       \frac{-1}{n} (2\beta_S-1) \frac{1}{p^2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2
+  \rightarrow 0,
+\end{align*}
+where the last result follows because by assumption, $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$.
+\end{proof}
+
 \subsection{Proof of Proposition \eqref{prop:linear}}
@@
 
-  We first rewrite the risk estimate $\widehat{RE}(\beta_S, \beta_I)$.
-Define $\bs{M}=(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$.
+  We first rewrite the risk estimate $\hat{R}(\beta_S, \beta_I)$.
+Define $\bs{M}=(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$.
 Then the unbiased risk estimator can be re-written as
   \[
- p^2 \widehat{RE}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
+ p^2 \hat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
   \]
@@
   \begin{align*}
-    \widehat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\widehat{RE}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
+    \hat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\hat{R}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
     &
-    \widehat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\widehat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
+    \hat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\hat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
   \end{align*}
 
-  We will need to show $\widehat{\mu} b_n^2 / d_n^2 \to \widehat{\beta}_I$ and $\widehat{\beta}_I/\widehat{\mu} + \widehat{\beta}_S = 1$. Since
+  We will need to show $\hat{\mu} b_n^2 / d_n^2 \to \hat{\beta}_I$ and $\hat{\beta}_I/\hat{\mu} + \hat{\beta}_S = 1$. Since
 $$\bs{Z}^\top \bs{Z} = \begin{pmatrix}
@@
 \bs{Z^\top \bs{v_S}}-\bs{M} =  \begin{pmatrix}
-&\sum_{j,k=1}^ps_{jk}^2 - \widehat{\Delta}_{jk}^2\\
+&\sum_{j,k=1}^ps_{jk}^2 - \hat{\Delta}_{jk}^2\\
 &\sum_{j=1}^p s_{jj}
@@
 \begin{align*}
-\widehat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z^\top \bs{v_S}}-\bs{M} ) \\
+\hat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z^\top \bs{v_S}}-\bs{M} ) \\
 &= \frac{1}{p^2d_n^2}\begin{pmatrix}
@@
 \end{pmatrix}\begin{pmatrix}
-\sum_{j,k=1}^p(s_{jk}^2 - \widehat{\Delta}_{jk}^2)\\
+\sum_{j,k=1}^p(s_{jk}^2 - \hat{\Delta}_{jk}^2)\\
 \sum_{j=1}^p s_{jj}
@@
 &= \frac{1}{p^2d_n^2}\begin{pmatrix}
- p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
- (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)
+ p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
+ (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)
 \end{pmatrix}.
 \end{align*}
-The second component of $\widehat{\bs{\beta}}$ equals $\widehat{\beta}_I$, so
+The second component of $\hat{\bs{\beta}}$ equals $\hat{\beta}_I$, so
 \begin{align*}
-\widehat{\beta}_I &= \frac{1}{p^2d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)\\
-&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2)/p \} / d_n^2
-=\widehat{\mu}\frac{b_n^2}{d_n^2}.
+\hat{\beta}_I &= \frac{1}{p^2d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)\\
+&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)/p \} / d_n^2
+=\hat{\mu}\frac{b_n^2}{d_n^2}.
 \end{align*}
@@
 \begin{align*}
-\widehat{\beta}_I /\widehat{\mu} + \widehat{\beta}_S &= \frac{1}{p^2d_n^2}\sum_{j,k=1}^p \{ps_{jk}^2 -p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\widehat{\Delta}_{jk}^2\}\\
+\hat{\beta}_I /\hat{\mu} + \hat{\beta}_S &= \frac{1}{p^2d_n^2}\sum_{j,k=1}^p \{ps_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2\}\\
 &=\frac{1}{p^2d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}
@@
 \subsection{Proof of Proposition \eqref{prop:bayes risk}}
-\begin{proof}
-Using the fact that when $\bb{E}\bs{X}=\bs{0}$, $\bb{E}s_{jk} = \sigma_{jk}$ for all $j,k=1,\ldots,p$
-\begin{align*}
-R(\beta_{S},\beta_{I}) &=\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} \{ (\beta_{S}s_{jk}+\beta_Iu_{jk}-\sigma_{jk})^2 \} \\
-& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} [s_{jk}-\sigma_{jk} - \{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}]^2\\
-& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E}[(s_{jk}-\sigma_{jk})^2+ \bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2-2(1-\beta_{S})\bb{E}s_{jk}(s_{jk}-\sigma_{jk})]\\
-& =\frac{1}{p^2}\sum_{j,k=1}^p [(2\beta_{S}-1)\text{Var}(s_{jk})+\bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2] \\
-& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\frac{n}{n-1}\widehat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2]
-\end{align*}
-The last equality is based on the fact that $\text{Var}(s_{jk}) = \frac{1}{n(n-1)}\sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2$.
-\end{proof}
+
 \label{lastpage}
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2021-11-07 msg.tex
update images, weight iteration and some typos

@@
 Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis.
-Our compound decision approach is conceptually similar the empirical Bayes approach taken by \citet{dey2018corshrink} to estimate a high-dimensional correlation matrix.
+Our compound decision approach is conceptually similar to the empirical Bayes approach taken by \citet{dey2018corshrink} to estimate a high-dimensional correlation matrix.
 We discuss in detail the connections with their work in Section \ref{sec:proposed}.
@@
 where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature.
-In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $\tilde{t}$ observations of the corresponding feature and off-diagonal entries using a function $t$ of observations from the two corresponding features, where $t$ and $\tilde{t}$ do not depend on the indices $j$ and $k$.
+In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $\tilde{t}$ of observations of the corresponding feature and off-diagonal entries using a function $t$ of observations from the two corresponding features, where $t$ and $\tilde{t}$ do not depend on the indices $j$ and $k$.
 Furthermore, we enforce rule in $\mathcal{S}$ to give symmetric estimates of the off-diagonal entries.
@@
 The decision rule which minimizes the Frobenius risk depends on all the $\sigma_j$ and $r_{jk}$.
-The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star((\bs{X}_{\cdot j},\bs{X}_{\cdot k}))$ for $j> k$ and $\delta^\star_{jj}(\bs{X}) = \widetilde{t}^\star(\bs{X}_{\cdot j})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid (\bs{X},\bs{X}^{'})\}$ and $\widetilde{t}^\star = \mathbb{E}\{a^2 \mid \bs{X}\}$ minimizes the Bayes risk.
+The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star((\bs{X}_{\cdot j},\bs{X}_{\cdot k}))$ for $j> k$ and $\delta^\star_{jj}(\bs{X}) = \widetilde{t}^\star(\bs{X}_{\cdot j})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid (\bs{X},\bs{X}^{'})\}$ and $\widetilde{t}^\star = \mathbb{E}\{a^2 \mid \bs{X}\}$ minimize the Bayes risk.
 
@@
   \hat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}_{nd}} \{\prod_{1\leq k\leq j<p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\} 
-  \{\prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG_d(a)\},
+  \{\prod_{j = 1}^p \int \widetilde{f}(\bs{X}_{\cdot j} \mid a) dG_d(a)\},
 \end{equation}
@@
 
-Calculating the estimated prior $\hat{G}_{nd}$ \eqref{Gnd hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
-\citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points.
+Calculating the estimated prior $\hat{G}_{nd}$ \eqref{Gnd hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}_{nd}$ supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$.
+\citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p(p+1)/2$ points.
 The EM algorithm has traditionally been used to estimate the locations of the support points and the masses at those points \citep{laird1978nonparametric}, but this is a difficult nonconvex optimization problem.
@@
 Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$.
+The support grid points are constructed symmetrically, satisfying $\bs{\xi}_{\tau+D/2}=(b_{\tau},a_{\tau},\gamma_{\tau})$ for $\tau=1,\ldots,D/2$.
+Since the empirical distribution of diagonal entries $G_{d}$ is symmetric about the first two parameters, so we would to like enforce the weights $w_{\tau}$ also have symmetric property.
 We can then use the EM algorithm to estimate the masses $\hat{\bs{w}}=\{\hat{w}_{1},\ldots, \hat{w}_{D}\}$ at those points via the iteration
-\begin{equation}
+\begin{align*}
 \label{eq:iteration}
-\hat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)}
-\end{equation}
-over $k$.
+%\hat{w}_{\tau}^{(k)} = \frac{2}{p(p-1)}\sum_{1\leq k< j \leq p} \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)}
+\hat{w}_{\tau}^{(k)} = \hat{w}_{\tau+D/2}^{(k)}  &= \frac{2}{p(p-1)} [\sum_{1\leq k< j \leq p} \frac{\hat{w}_{\tau}^{(k-1)} \{f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau})+f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau+D/2})\}}{\sum_{l=1}^{D/2}\hat{w}_l^{(k-1)} \{f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)+f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_{l+D/2})\}} \\
+&+\sum_{j=1}^p \frac{\hat{w}_{\tau}^{(k-1)} \{\widetilde{f}(\bs{X}_j \mid a_{\tau})+\widetilde{f}(\bs{X}_j \mid a_{\tau+D/2})\}}{\sum_{l=1}^{D/2}\hat{w}_{l}^{(k-1)} \{\widetilde{f}(\bs{X}_j \mid a_{l})+\widetilde{f}(\bs{X}_j \mid a_{l+D/2})\} }]
+\end{align*}
+over $k$ and the iteration guarantees symmetry.
 Early stopping of the EM algorithm can be useful \citep{koenker2019comment}, and more sophisticated convex optimization procedures can be used as well \citep{koenker2014convex}.
@@
 s_{kj} & s_{kk} 
-\end{bmatrix}$ instead of $(\bs{X}_j,\bs{X}_k)$ since it is the sufficient statistics of $(\bs{X}_j,\bs{X}_k)$ with respect to $\bs{C}_{jk}$. Based on the property of sufficient statistics, 
+\end{bmatrix}$ instead of $(\bs{X}_j,\bs{X}_k)$ since it is the sufficient statistic of $(\bs{X}_j,\bs{X}_k)$ with respect to $\bs{C}_{jk}$. Based on the property of sufficient statistics, 
 \begin{equation}
@@
 
-We considered five models for the population covariance matrix.
+We considered six models for the population covariance matrix to explore the behavior of our method in different settings.
 For the first four settings, $\bs{\Sigma} = \text{diag}(\bs{s}) \bs{C} \text{diag}(\bs{s})$, where $\bs{C}$ is correlation matrix and $\bs{s}$ is a vector of standard deviations.
 \begin{itemize}
-\item Model 1. The standard deviations were independently generated from $\mathcal{U}(1,1.5)$ and the correlation matrix followed Model 2 of \citet{cai2011adaptive}:
+\item Model 1 (Sparse). The standard deviations were independently generated from $\mathcal{U}(1,1.5)$ and the correlation matrix followed Model 2 of \citet{cai2011adaptive}:
   \[
@@
 
-\item Model 2. The first $p / 2$ standard deviations equaled 1, the last $p / 2$ equaled 2, and the correlation matrix was
+\item Model 2 (Block). The first $p / 2$ standard deviations equaled 1, the last $p / 2$ equaled 2, and the correlation matrix was
   \[
@@
 
-\item Model 3.
+\item Model 3 (Dense).
 The standard deviations were generated independently from $\mathcal{U}(1, 1.5)$ and $\bs{C}$ was a compound symmetric matrix with correlation parameter 0.7.
@@
 
-\item Model 4.
+\item Model 4 (Dense2).
 This setting was the same as Model 3 except with correlation parameter 0.9.
@@
 
-\item Model 5.
-With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4.
-This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
+\item Model 5 (Orth).
+With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues independently generated from $\mathcal{U}(1, 4)$.
+This referred to simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
 
-\item Model 6.
+\item Model 6 (Spiked).
 With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p -3$ entries equaled 1.
@@
 \end{center}
-\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, Orth and Spiked panels correspond to Models 1 through 5, respectively.}
+\caption{Median Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, Orth and Spiked panels correspond to Models 1 through 6, respectively.}
 \label{fig:sim1_frobenius}
@@
 \end{center}
-\caption{Average Frobenius norm errors over 200 replications.
-The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.
+\caption{Median Frobenius norm errors over 200 replications.
+The Sparse, Block, Dense, Dense2, Orth and Spiked panels correspond to Models 1 through 6, respectively.
 In MSG, $K$-means clustering is applied with $K=p$}
@@
 Figure \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators.
-Our MSG methods had the lowest or near-lowest errors across all settings except for Model 4 which has high correlations 0.9.
-This is not surprising because our method assumed independence of $\bs{A}_{jk}$.
-In some cases, for example in Models 1 and 2, the improvement was substantial.
+From Model 1 to Model 4 when the prior distributions of $G_{nd}$ are simple, our MSG methods had the lowest errors across all settings.
+For Model 4 which has high correlations 0.9, the priority is less obvious.
+This is not surprising because our method assumed independence of $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$.
+In some cases, the improvement was substantial.
 Model 2 was especially interesting because the standard deviation and correlations were related.
 Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_{nd}$ \eqref{bayesian nondiagonal} and leverage it to provide much more accurate estimates.
-The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5.
+The nonlinear shrinkage estimators outperformed MSG in Model 5 and Model 6.
+From the plot, we can observe that CorShrink, of which the target is to estimate the correlation matrix, also had competitive performance in all settings.
 In every setting, correcting MSG for positive-definiteness never increased the risk and decreased the risk in some cases.
@@
 Finally, the simulations show that the class of separable estimators \eqref{separable} proposed in this paper is fundamentally different from the class of rotation-invariant estimators, as the oracle optimal estimators in these two classes behave very differently.
-For example, the oracle separable estimator had vanishing risk in Model 2, while the oracle rotiation-invariant estimator does not.
-Separable estimators seemed better for Models 1 and 2 while rotation-invariant estimators were superior in Models 3 and 4.
+For example, the oracle separable estimator had vanishing risk in Model 2, 3 and 4, while the oracle rotiation-invariant estimator does not.
+Separable estimators seemed better for Models 1 to 4 while rotation-invariant estimators were superior in Models 6.
 They seem comparable in Model 5.

### 2021-12-01 response_reviews.tex
revised for biometrics to respond to reviewer comments

@@
 \usepackage{lscape}
+
+\usepackage{xr}
+\externaldocument{msg}
 
@@
 
+Thank you for your very valuable comments. We respond to each point below.
+
 \begin{enumerate}
@@
 I believe that the authors should motivate their approach from the gene expression data or other bio-related applications starting from the Introduction.}
+
+  We have revised our introduction and have also completely redone our data analysis to study a question about gene network inference that arose during some of our previous collaborative work.
+We were involved in an analysis of bulk RNA-seq expression from mouse brain in a study of the neurogenomic response to social intrusion (Saul et al., 2017).
+We have now used some of these data to study gene networks in different regions of the mouse brain; \textbf{see Section 5}.
+
 \item \emph{It is not really intuitively clear why the vectorization works, despite the approach not taking into consideration the structure of the covariance matrix.
@@
 
-\item \emph{I agree with one of the referees that the title should be re-thought, since compound decision theory or compound decisions appear quite relevant to the approach, more than the empirical Bayes aspect of the paper.}\\
-We agree that compound decision theory should be more emphasized than empirical Bayes in our paper.
-Following your suggestions, we have changed the title from "A Empirical Bayes Approach to Covariance Matrix Estimation" to "A Compound Decision Approach to Covariance Matrix Estimation".
+  The success of our vectorization perspective is indeed surprising. In our revision we have made several attempts to contextualize this finding.
+  \begin{enumerate}
+
+  \item In \textbf{Section 2.1} we now note that measuring the success of covariance matrix estimation using Frobenius risk does not incentivize positive-definiteness.
+In other words, there can exist estimators of $\bs{\Sigma}$ that have low Frobenius risk but which are not positive-definite.
+So the fact that vectorization works is in part a consequence of the risk function.
+
+  \item In \textbf{Section 2.2} we note that several existing shrinkage methods can already be interpreted as vectorized methods.
+The sample covariance matrix, the linear shrinkage method of Ledoit and Wolf (2004), and the adaptive thresholding method of Cai and Liu (2011).
+The adapative thresholding method is also not guaranteed to be positive-definite.
+
+  \item In \textbf{Section 4.4} we now study one possible reason why vectorization may work better than existing rotationally invariant procedures that shrink only eigenvalues but use sample eigenvectors in their estimators.
+Since sample eigenvectors can be poor estimators of population eigenvectors in high-dimensions, we hypothesized that our vectorized approach would be better in simulation settings where the sample and population eigenvectors were more different.
+This appears to be the case; \textbf{see Figure 3}.
+  \end{enumerate}
+
+  We agree that projecting our final estimator on to the space of positive-definite matrices is not warranted by the compound decision formalism.
+Indeed, our uncorrected estimator still works very well in our simulation settings, and furthermore the (uncorrected) oracle vectorized estimator can be dramatically better than the oracle rotationally invariant estimator; \textbf{see Figure 2}.
+Nevertheless, it does appear that in finite samples, positive-definiteness correction can improve performance.
+We mention in the \textbf{Discussion} that a non-separable class of estimators that can incorporate this constraint is worth exploring in the future.
+
+  In our new \textbf{Section 3.3} we have also clarified our EM updates and have added details about iimplementation.
+
+  In short, in our paper we explore a new perspective on covariance matrix estimation, a vectorization perspective founded on a compound decision theoretic interpretation of the Frobenius risk.
+We define a new class of estimators, derive the optimal vectorized estimator within this class, and estimate the optimal estimator using approaches from the compound decision literature.
+Finally, we argue that these estimators work well for covariance matrices whose sample and population eigenvectors are very different.
+
+\item \emph{I agree with one of the referees that the title should be re-thought, since compound decision theory or compound decisions appear quite relevant to the approach, more than the empirical Bayes aspect of the paper.}
+
+  We have changed our title following your suggestions.
 
@@
 It somehow appears the choice is dictated by Ledoit et al (2004) and Proposition 1, and not by independent first principles.
-If so, it should be explicitly stated before.}\\
-We squared and scaled the Frobenius norm $\|A\|_F = \sqrt{\sum_{j,k=1}^pa_{jk}^2}$ to $\frac{1}{p^2}\sum_{j,k=1}^p a_{jk}^2$ in order to apply fundamental theorem of compound decisions, for which the risk function should be scaled by the number of parameters because the empirical distribution has mass probability of $\frac{1}{d}$ where $d$ is the number of parameters.\\
-The transformation of Frobenius norm will not change the value of $(\widehat{\beta}_S,\widehat{\beta}_I)$ since it is determined by minimizing $RE(\beta_S,\beta_I)$.
-\item \emph{Figures 1 and 2 do not allow to appreciate uncertainty. Perhaps a table would be better.}\\
-We add the tables of the medians and uncertainty of Frobenius errors in supporting information.
-The uncertainty of Frobenius error is measured by the difference of 75\% quantile and 25\% quantile in 200 replicates.
-We use this range to measure the uncertainty because the errors of most methods are not following normal distribution.
+If so, it should be explicitly stated before.}
+
+  We have clarified in \textbf{Section 2.2} that the choice $\hat{R}$ follows directly from first principles.
+  \begin{enumerate}
+  \item In (4) we restrict ourselves to the class of linear vectorized estimators of $\bs{\Sigma}$.
+
+  \item In Proposition 1 we show that $\hat{R}$ is an asymptotically unbiased estimate of the Frobenius risk for any fixed $\bs{\Sigma}$ and any fixed linear estimator.
+
+  \item A natural procedure is to minimize the empirical risk $\hat{R}$ over the class of linear estimators.
+
+  \item In Proposition 2 we show that the estimator obtained after empirical risk minimization is very closely related to the Ledoit and Wolf (2004) estimator.
+  \end{enumerate}
+
+\item \emph{Figures 1 and 2 do not allow to appreciate uncertainty. Perhaps a table would be better.}
+
+  We now give exact numerical results for all simulations in our \textbf{Supporting Information}.
+
 \item \emph{The manuscript requires careful proofreading and improvements. in the structure of the presentation and the visualization of the results. Non-exhaustive examples are reported below:}
@@
 
-It has been fixed.
+This has been fixed.
 
@@
 
-It has been edited to 'banded'.
+This has been edited to ``banded''.
 
@@
 It means that these unstructured model have been shown to have lower estimation risk comparing to sample covariance matrix.
-For example, in \cite{ledoit2004}, the numerical result shows that their linear shrinkage method has much lower risk than sample covariance matrix.
-To make it clear, we have changed to "yet still have less estimation error comparing to the sample covariance matrix".
+To make it clear, we have changed to ``yet still have less estimation error comparing to the sample covariance matrix''.
 
@@
 
-We used to use \textbackslash ref to reference equation, we have changed it to \textbackslash eqref so that they are referenced between brackets.
+This has been fixed.
 
@@
 
-We have moved the setting of the experiment to the start of Section 3 Numerical Results.
+This has been fixed.
 
@@
 
-We have changed the notation $\hat{R}$ to $\text{RE}$ which means risk estimator and this differentiates it from $R$. 
+We now use $\hat{R}_L$ to denote the risk estimate for the class of linear decision rules. We also make the distinction between $\hat{R}_L$ and $R$ more explicit in \textbf{Proposition 1}.
 \end{enumerate}
 
-\section{Comments to Referee1}
-Thank you so much for the detailed and inspiring suggestions on our paper.
-We learnt a lot through them.
-We agree with your comments and address them in the updated draft version.
-Please find the following responses for more details.
+\section{Response to Referee 1}
 
-\subsection{Comments to section "CorShrink"}
-After carefully reviewed \cite{dey2018}, we think it is very related to our paper and it is valuable to make the comparison.
+Thank you for your positive comments, and your careful reading and extremely helpful suggestions. We respond to them below.
 
-As mentioned in Section 1, our method is quite similar to \cite{dey2018}.
-Both of us used empirical Bayes estimator and nonparametric estimation of prior distribution.
-The difference is that our method aim to estimate covariance matrix which requires to also shrink standard deviations.
+\subsection{Response to referee section ``CorShrink''}
 
-We also compared our two methods by numerical experiments which are shown in supporting information.
-We designed two simulations to compare the performance on correlation matrix estimation and covariance matrix estimation.
-In the first simulation, we compare covariance matrix estimation.
-For corShrink, $\widehat{\bs{\Sigma}} = \text{diag}(\widehat{SD}) \widehat{\bs{R}} \text{diag}(\widehat{SD})$, where $\widehat{SD}$ are sample standard deviations.
-In the second simulation, $\bs{\widehat{R}} = \text{diag}(1/\widehat{SD})\widehat{\bs{\Sigma}}\text{diag}(1/\widehat{SD})$, where $\widehat{SD}$ are standard deviations derived from $\widehat{\bs{\Sigma}}$.
-For each setting, we take $n$ as 20 and 100.
+Thank you so much for bringing our attention to Dey and Stephens (2018). It is indeed extremely closely related to our work. We now discuss connections to their work in several places.
 
-\subsection{Comments to section "Separable rules, symmetry and the diagonal"}
-\begin{equation}\label{R1}
-\widehat{\Sigma} = (\widehat{\sigma}_{jk})_{1\leq j,k\leq p},\quad \widehat{\sigma}_{jk} = t(X_{\cdot,j}, X_{\cdot,k}), \quad t:\bb{R}^n\times \bb{R}^n\to \bb{R}
-\end{equation}
+\begin{enumerate}
+\item In the last paragraphs of \textbf{Section 3.2} we note that the main difference between our approaches is that Dey and Stephens (2018) do not shrink the variances while we do, as our $\hat{t}_d$ estimates $\sigma_j$ using its supposed posterior expectation assuming a prior distribution of $\hat{G}_d$.
+Furthermore, our Proposition \ref{prop:bayes risk} shows that our shrinkage approach can be interpreted as viewing the triples $(\sigma_j, \sigma_k, r_{jk})$ as if they were drawn from the trivariate prior $G_{od}$, which allows for possible dependencies between $r_{jk}$ and the $\sigma_j$.
+For example, model 2 in our simulations in Section \ref{numerical results} considers a covariance matrix where entries with larger standard deviations also tend to have larger correlations.
+In contrast to the method of Dey and Stephens (2018), our approach can learn this structure if it exists and take advantage of it for more accurate estimation.
+We demonstrate this in Model 2 in our simulations.
+
+\item We have added CorShrink to our simulation studies for estimating both covariance matrices (\textbf{Figure \ref{fig:sim1_frobenius}}) and correlation matrices (\textbf{Table 3 in the Supporting Information}).
+We found that for estimating covariance matrices, our approach could outperform CorShrink in many cases, likely because it was able to accurately shrink the standard deviations.
+In the \textbf{Supporting Information} we also compare MSG and MSGCor to CorShrink for estimating correlation instead of covariance matrices.
+We find that in that case, CorShrink performs slightly better than MSG and MSGCor except in our Model 2, likely because the additional flexibility that our method trades off lower bias for higher variance.
+
+\end{enumerate}
+
+\subsection{Response to referee section ``Separable rules, symmetry and the diagonal''}
+
 \begin{enumerate}
 \item \emph{While reading the paper I was a bit confused by the fact that (R1) treats diagonal entries and off-diagonal in the same way.
-It would feel more natural to me to define a separable estimator as in (R1) for $j\neq k$, and as follows on the diagonal:}
-$$\hat{\sigma}_{jj} = \tilde{t}(X_{\cdot j}),\quad \tilde{t}:\bb{R}^n \to \bb{R}$$
-We have changed the rule to 
-\begin{equation}
-  \label{separable}
+It would feel more natural to me to define a separable estimator as in (R1) for $j\neq k$ \ldots In the Multivariate Gaussian case this is not an issue, since $X_{\cdot j}\neq X_{\cdot k}$ for $j\neq k$ with probability 1.
+But the proposed approach would also apply with minor modifications to e.g., discrete measurements (with known likelihood) in which case $X_{\cdot j}\neq X_{\cdot k}$ could happen with non-zero probability.
+Furthermore, while this issue is ignored for most of the paper, diagonal and off-diagonal entries are treated in a different way at the end of Section 2.3 (in the context of the clustering based algorithm).
+I believe it would be valuable to make the above distinction more explicit!}
+
+  Thank you for this excellent suggestion. We completely agree and have now redefined our class of separable estimators to be
+  \[
   \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
    \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
-\end{equation}
-\emph{In the Multivariate Gaussian case this is not an issue, since $X_{\cdot j}\neq X_{\cdot k}$ for $j\neq k$ with probability 1.
-But the proposed approach would also apply with minor modifications to e.g., discrete measurements (with known likelihood) in which case $X_{\cdot j}\neq X_{\cdot k}$ could happen with non-zero probability.} \\
-We agree that it is not very precise to use $I(\bs{X}_{\cdot j}=\bs{X}_{\cdot k})$ to differentiate on-diagonal entries and off-diagonal entries.
-So we changed separable rule to the above class of rules.
+ \]
+ so that the on- and off-diagonal entries are treated separately.
+In \textbf{Section \ref{sec:class}} we derive the optimal estimator in this class.
+This estimator turns out to be the same as the estimator we actually implemented in our previous draft, but redefining $\mathcal{S}$ makes this much more rigorous.
 
-\emph{Furthermore, while this issue is ignored for most of the paper, diagonal and off-diagonal entries are treated in a different way at the end of Section 2.3 (in the context of the clustering based algorithm).
-I believe it would be valuable to make the above distinction more explicit!}
-\item \emph{A related question I had concerns the symmetry of the estimator. Does the estimator satisfy the symmetry property $\hat{\sigma}_{jk} =\hat{\sigma}_{kj}$? }
+\item \emph{A related question I had concerns the symmetry of the estimator.
+Does the estimator satisfy the symmetry property $\hat{\sigma}_{jk} =\hat{\sigma}_{kj}$?
+Similarly, the reader may wonder whether the estimation procedure enforces symmetry in the prior $G$ in (6) with respect to $\sigma_j,\sigma_k$.}
 
-Yes, we only calculate estimates of lower triangular matrix of $\Sigma$, i.e., $\hat{\sigma}_{jk}$ where $j\geq k$.
-Then the upper triangular matrix estimation is calculated by the symmetry of covariance matrix $\sigma_{kj} = \sigma_{jk}$.
-Correspondingly, we should fix the Bayesian model in Section2.2 from
-\begin{equation}
-  \label{bayesian}
-  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
-  \quad
-  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
-\end{equation}
-to 
-\begin{equation}
-  \label{bayesian nondigaonal}
- (\bs{X}_{\cdot j},\bs{X}_{\cdot k})\mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
-  \quad
-  \bs{\eta} \sim G_{nd}(a,b, \gamma) =  \frac{2}{p(p-1)} \sum_{1\leq k< j \leq p} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
-\end{equation}
-and
-\begin{equation}
-  \label{bayesian nondigaonal}
-  \bs{X}_{\cdot j} \mid a \sim \widetilde{f}(\cdot \mid a),
-  \quad
-  a \sim G_{d}(a) =  \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a).
-\end{equation}
-, as well as the proposed estimator from
-\begin{equation}
-  \label{Gp hat}
-  \widehat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
-\end{equation}
-to
-\begin{equation}
-  \label{Gp hat revision}
-  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}} \prod_{1\leq k< j \leq p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\prod_{j=1}^p \int \widetilde{f}(\bs{X}_{\cdot j} \mid a) dG_{d}(a),
-\end{equation}
-\emph{Similarly, the reader may wonder whether the estimation procedure enforces symmetry in the prior $G$ in (6) with respect to $\sigma_j,\sigma_k$.}
+  In our new implementation we now explicitly enforce symmetry.
+We construct the support points of our estimated discrete prior to be symmetric and we also force the estimated masses at these points to be symmetric as well.
+See \textbf{Section \ref{implementation}} for details.
 
-In correspondence, we modified our sample exemplar algorithm, we first clusters sample points $(s_j,s_k, r_{jk})$ of lower triangular part of sample covariance matrix and get the $K$ centroids $(a,b,\gamma)$, then find their symmetric points $(b,a,\gamma)$.
-We also updated EM algorithm to guarantee each $(a,b,\gamma)$ and $(b,a,\gamma)$ has the same weight.
 \item \emph{The paper emphasizes the Multivariate Gaussian problem.
@@
 
-$X_{\cdot j}^TX_{\cdot k}$ is sufficient statistics of the single entry $\sigma_{jk}$.
-In fact, for $(\bs{X}_{\cdot,j}, \bs{X}_{\cdot,k})$, the sufficient statistics for $\eta$ is $\bs{S}_{jk} = \begin{pmatrix} s_{jj}&s_{jk} \\  s_{kj}&s_{kk} \end{pmatrix}$.
-We also tried density based on single entries $s_{jk}$ before but the performance is not as good as using $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$.
+We chose our definition of a separable estimator because it is more general than $\bar{t}(\bs{X}_{\cdot j}^\top \bs{X}_{\cdot k})$.
+We actually initially tried this class of estimators but found that its performance was worse than the estimators we proposed in the manuscript.
+Intuitively, the reason is that the statistic $X_{\cdot j}^\top X_{\cdot j}$ is not sufficient for the parameter $\sigma_{jk}$.
+
 \item \emph{If I understand correctly, one of the benefits of estimator (3) is that it works without specific assumptions on the noise distribution.
@@
 
-This is a good point.
-To measure the performance of our method on nonGaussian distribution, we did some simulations on negative binomial distribution.
-We first generate independent distributed $Y$ and then let $X = L Y$, where L is the Cholesky decomposition of $\Sigma$.
-The simulation results are presented in supporting information.
+  We now study the performance of our proposed method when the multivariate Gaussian assumption is violated.
+\textbf{Tables 4 and 5 in the Supporting Information} show that our procedures can still provide relatively accurate estimates for uniform or negative binomial data.
+
 \end{enumerate}
 
-\subsection{Comments to section "Compound decision theory or empirical Bayes?"}
+\subsection{Response to referee section ``Compound decision theory or empirical Bayes?''}
 \begin{enumerate}
@@
 
-We rewrote the equation after (6) as a separate proposition to show the equivalence between the Frobenius risk and Bayes risk.
-\item \emph{The term “compound decision theory” or “compound decisions” etc. deserves to be in the title!}
+  We have followed your suggestion and added \textbf{Proposition \ref{prop:bayes risk}} along with a discussion after it highlighting the fact that the optimal estimator depends on all the $\sigma_{j}$ and $\rho_{jk}$.
 
-Yes, we agree that "compound decision" is more deserved to be highlighted in the paper so we replaced "nonparametric empirical Bayes" with "compound decision" in the title.
+\item \emph{The term ``compound decision theory'' or ``compound decisions'' etc. deserves to be in the title!}
+
+  Thank you for the suggestion! We have revised the title accordingly.
+
 \item \emph{On the other hand, I would downemphasize the empirical Bayes aspect of the paper (and my preference would be to even remove it from the title).
-To me, an empirical Bayes analysis (especially since here the goal is to perform well, in a frequentist sense, in terms of (R2)) has the following flavor: “Let us mimic what a Bayesian would do, when we do not know the true prior, and in a way that we can get frequentist guarantees as well”.
+To me, an empirical Bayes analysis (especially since here the goal is to perform well, in a frequentist sense, in terms of (R2)) has the following flavor: ``Let us mimic what a Bayesian would do, when we do not know the true prior, and in a way that we can get frequentist guarantees as well''.
 However, I do not think a Bayesian would start approaching this problem by first positing (6) and then using (8)2.
@@
 I would restrict the discussion of empirical Bayes to just saying that model (6) appears formally when optimizing over separable rules (R1).
-To approximate the unknown optimal rule, one follows an empirical Bayesian “Generalized Maximum Likelihood” approach with respect to the pairwise composite likelihood.3}
+To approximate the unknown optimal rule, one follows an empirical Bayesian ``Generalized Maximum Likelihood'' approach with respect to the pairwise composite likelihood.}
+
+  This is an excellent point and we have done exactly as you suggested. We now only refer to empirical Bayes ideas as a method of estimating the optimal compound decision rule.
+
 \end{enumerate}
 
-\subsection{comments to section "other remarks"}
+\subsection{Response to section ``Other remarks''}
 \begin{enumerate}
@@
 
-We put our code of methods and simulation in supporting information.
+  We now note at the \textbf{end of the Introduction} that our code is available on Github and give the link.
+
 \item \emph{Related to the above point, based on the description in the manuscript it would not be possible for me to fully reproduce the results.
@@
 
-In our implementation, we used EM algorithm to estimate the prior distribution.
-We added the related details to Section 3 Numerical Results.
-The maximum number of iterations is set as 200.
-Iteration early stops when the criterion $\frac{L(\bs{w}^{(k+1)}) - L(\bs{w}^{(k)})}{L(\bs{w}^{(k)})}\leq 10^{-4}$ is met.
-We also tried interior point, but it always failed to find the optimal solution.
-The possible reason is the density $f(\cdot|a,b,\gamma)$ has very different scales, from $10^{-22}$ to $2$ and this makes interior point method difficult to solve.
-\item \emph{First paragraph of page 3, “they modify only the sample eigenvalues and not the sample eigenvectors”: The situation here is a bit more nuanced, since e.g., sometimes using the sample eigenvectors is provably the optimal thing to do (e.g., if one constrains themselves to rotationally invariant estimators as in Bun et al.
+  We have revised our description of our implementation in \textbf{Section \ref{implementation}} to make our EM algorithm more clear.
+We aso note that instead of EM, we attempted to maximize the composite likelihood \eqref{Gnd hat} using the interior point solver MOSEK, as advocated by Koenker and Mizera (2014), but ran into computational difficulties because for even moderate sample sizes $n$, the values of the likelihoods we needed to calculate could be extremely small at certain support points.
+These small values caused problems for the R wrapper for MOSEK.
+Our implementation of the EM algorithm was more robust to these small density values.
+
+\item \emph{First paragraph of page 3, ``they modify only the sample eigenvalues and not the sample eigenvectors'': The situation here is a bit more nuanced, since e.g., sometimes using the sample eigenvectors is provably the optimal thing to do (e.g., if one constrains themselves to rotationally invariant estimators as in Bun et al.
 [2016] and references therein)}
 
-The sentence "However, they modify only the sample eigenvalues and not the sample eigenvectors." is not very appropriate so we changed it to "However, as rotation-equivariant estimators (Stein, 1975, 1986), shrinkage estimators modify only the sample eigenvalues and keep the sample eigenvectors.".
-\item \emph{Page 5, “it is straightforward to show that”: The “straightforward” could be replaced by a short proof in the appendix.}
+  Thank you for this reference.
+We now explicitly distinguish our proposed method from rotationally invariant ones and note in our \textbf{Introduction} that while rotationally invariate estimators perform extremely well, Bun et al.
+(2016) showed that the eigenvectors of any rotationally invariant estimate must be the same as those of the sample covariance matrix.
+This can be problematic because sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate  (Mestre, 2008).
+This suggests that unstructured but non-rotationally invariant covariance matrix estimators may be worth exploring, as these can modify both the sample eigenvectors as well as the eigenvalues.
+Our contribution is to propose and study one such class of non-rotationally invariant estimators.
 
-We added the proof in the appendix.
+  Incidentally, our new simulations in \textbf{Figure \ref{fig:eigenvectors}} suggest that the simulation settings where our estimators can outperform existing methods are exactly those where the sample and population eigenvectors differ most.
 
-\emph{Also could it be that the risk estimate is not exactly unbiased (because of dividing by $n$ and not $n-1$ in the definition of $\hat{\Delta}_{jk}^2$)?}
-You're right.
-We realized that the previous defined $\hat{\Delta}_{jk}^2$ is not unbiased estimator of Frobenius risk of $\bs{\delta}$ (1).
-So we changed the text to describe it as an "approximately unbiased risk estimator" since the factor $\frac{n-1}{n}$ is closed to 1.
-It is also shown in appendix B.
+\item \emph{Page 5, ``it is straightforward to show that'': The ``straightforward'' could be replaced by a short proof in the appendix.}
 
-\emph{Where is the sample covariance $\bs{S}$ defined?}
+  We have rewritten this as \textbf{Proposition \ref{prop:Rhat}} and have added a proof to the Supporting Information.
 
-In the updated manuscript, $\bs{S}$ is defined as $\bs{S} = \frac{1}{n}\bs{X}^\top \bs{X}$ as it is first mentioned in page 4, section 2.1.
+\item \emph{Also could it be that the risk estimate is not exactly unbiased (because of dividing by $n$ and not $n-1$ in the definition of $\hat{\Delta}_{jk}^2$)?}
 
-\item \emph{Section 2.5, “Projections in terms of other matrix norms are also possible”. Is this actually being done in the implementation?}
+  You are absolutely right, thank you for catching this. We have changed our text to mention that the risk estimate is only asymptotically unbaised, and we show this in the Supporting Information.
 
-No. We did not implement it in terms of other norms.
+\item \emph{Where is the sample covariance $\bs{S}$ defined?}
 
-\emph{Else perhaps remove this sentence.}
+  We now define $\bs{S}$ in the first paragraph of \textbf{Section \ref{sec:connections}}.
 
-It has been removed.
+\item \emph{Section 2.5, ``Projections in terms of other matrix norms are also possible''. Is this actually being done in the implementation?}
+
+  We did not implement this and so have removed this comment from the manuscript.
+
+\item \emph{Else perhaps remove this sentence.}
+
+This has been removed.
+
 \end{enumerate}
-\section{Comments to Referee2}
+
+\section{Response to Referee 2}
+
+Thank you for your positive comments. We respond to them below.
+
 \begin{enumerate}
-\item \emph{In the numerical experiments and the real data analysis, what was the size of the grid that was used? I may have missed it but I did not see it mentioned in the caption for figure 2.}\\
-In section 3.3, methods compared and section 3.4, data analysis, we used clustering-based exemplar algorithm, where we find $K$ clusters of all off-diagonal sample points.
-$K$ is set to be $p$ to balance accuracy and speed according to result of section 3.2.
-\item \emph{In the numerical experiments, the dimension p goes as far as 200. I believe it will be useful to consider a setting where $n = 100$ and $p = 1000$.} \\
-We did the simulation for $n=100$ and $p=1000$ using the different covariance matrix models in Section 3.1. Please see the supporting information for the simulation results.  
+\item \emph{In the numerical experiments and the real data analysis, what was the size of the grid that was used? I may have missed it but I did not see it mentioned in the caption for figure 2.}
+
+  We have now added a separate \textbf{Section \ref{optimalK}} that specifies the choice of grid size for our proposed method.
+
+\item \emph{In the numerical experiments, the dimension p goes as far as 200. I believe it will be useful to consider a setting where $n = 100$ and $p = 1000$.}
+
+  We have added a simulation with $p = 1000$ in the \textbf{Supporting Information} and in \textbf{Section \ref{sec:accuracies}} note that while our methods perform well in our Models 1 through 4, other methods are better in Models 5 and 6 for large $p$.
+This is likely due to the fact that our class of estimators may not be be optimal in these settings because the population eigenvectors in these models appear relatively easy to estimate; \textbf{see the last paragraph of Section \ref{sec:accuracies} for details}.
+
 \item \emph{Continuing on the numerical experiments, models 1 to 5 impose various structures on $\bs{\Sigma}$.
-It will be interesting to see the performance of the proposed method if a spiked covariance structure is imposed on $\bs{\Sigma}$.} \\
-We added the spiked model as the sixth matrix models in Section 3.1.
-Spiked covariance matrix is generated by $\Sigma = \bs{U}\text{diag}(\bs{l})\bs{U}^\top$, where $\bs{U}$ is a randomly generated orthogonal unit matrix, $\bs{l} = \{4,3,2,1,\ldots,1\}$.
-\item \emph{Typo on page 6: "The density of $f(\cdot \mid \bs{\eta}_{jk})$ of ...".}\\
-Thanks for poiting this out. We've fixed this sentence as "$f(\cdot \mid \bs{\eta}_{jk})$, the density of $\bs{A}_{jk}$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, ...".
+It will be interesting to see the performance of the proposed method if a spiked covariance structure is imposed on $\bs{\Sigma}$.}
+
+  We have added a spiked covariance matrix as model 6 in our new simulations.
+As mentioned in our response above, we found that models 5 and 6 constitute settings where our methods may not work as well as existing approaches.
+
+\item \emph{Typo on page 6: "The density of $f(\cdot \mid \bs{\eta}_{jk})$ of ...".}
+
+  This has been fixed.
 \end{enumerate}
-\begin{thebibliography}{9}
-\bibitem{dey2018}
-Dey, K. K., \& Stephens, M. (2018). CorShrink: Empirical Bayes shrinkage estimation of correlations, with applications. bioRxiv, 368316.
-\bibitem{ledoit2004}
-Ledoit, O., \& Wolf, M. (2004). A well-conditioned estimator for large-dimensional covariance matrices. Journal of multivariate analysis, 88(2), 365-411.
 
-\end{thebibliography}
 \end{document}

### 2021-12-01 msg.tex
revised for biometrics to respond to reviewer comments

--- 2021-11-02 msg.tex
@@
 \usepackage{xcolor}
+\usepackage{url}
 
@@
 To achieve estimation accuracy in these settings, existing methods typically either assume that the population covariance matrix has some particular structure, for example sparsity, or apply shrinkage to better estimate the population eigenvalues.
-In this paper, we propose a new approach to estimating high-dimensional covariance matrices that can have better performance than existing methods.
+In this paper, we study a new approach to estimating high-dimensional covariance matrices.
 We first frame covariance matrix estimation as a compound decision problem.
 This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class.
-Simulation results and gene network inference in a single cell RNA-seq dataset show that our approach is comparable to or can outperform a number of state-of-the-art proposals in a wide range of settings.
+Simulation results and gene network inference in an RNA-seq experiment in mouse show that our approach is comparable to or can outperform a number of state-of-the-art proposals, particularly when the sample eigenvectors are poor estimates of the population eigenvectors.
 \end{abstract}
@@
 This high-dimensional setting is especially common in genomics, where covariance matrices are used to model gene networks but the number of genes can be much larger than the number of biological replicates \citep{schafer2005shrinkage, markowetz2007inferring}.
-Accurate gene network estimation has become even more important with the advent of single-cell RNA-sequencing (scRNA-seq) technology, where interest now centers on inferring cell- or cell type-specific gene networks \citep{aibar2017scenic, dai2019cell, stegle2015computational}.
-For example, in Section \ref{gene analysis} we study cell type-specific gene network inference using scRNA-seq data from honey bee brains.
-\blue{Need to do analysis}
+For example, in Section \ref{gene analysis} we study brain region-specific gene networks in mouse using bulk RNA-sequencing data.
 
@@
 In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have lower estimation error than the sample covariance matrix.
-A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix.
-More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}.
-These shrink the eigenvalues of the sample covariance matrix in a data-adaptive fashion.
-Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global means.
+Many of these methods can be interpreted as rotationally invariant estimators, which shrink the sample eigenvalues in a data-adaptive fashion \citep{bun2016rotational, stein1975estimation, stein1986lectures}.
+One example is the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the eigenvalues toward their global mean.
+Some more recently developed nonlinear shrinkage methods have been shown to be asymptotically optimal in this class \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}.
 
-Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance.
-However, most existing nonlinear shrinkage estimators belong to the class of rotation-equivariant estimators \citep{bun2016rotational, stein1975estimation, stein1986lectures}, which modify the sample eigenvalue estimates but still use the sample eigenvectors.
-On the other hand, sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}.
-This suggests that certain classes of non-rotationally invariant estimators may outperform existing unstructured covariance matrix estimators.
+Though these optimal estimators perform extremely well, \citet{bun2016rotational} showed that the eigenvectors of any rotationally invariant estimate must be the same as those of the sample covariance matrix.
+This can be problematic because sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}.
+This suggests that unstructured but non-rotationally invariant covariance matrix estimators may be worth exploring, as these can modify both the sample eigenvectors as well as the eigenvalues.
+The difficulty is to identify a reasonable class of estimators to pursue.
 
@@
 
-Here we propose a new class of unstructured non-rotationally invariant estimators for high-dimensional covariance matrices.
-Our approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}; see Section \ref{sec:compound}.
-This framing motivates us to vectorize the covariance matrix and solve the resulting vector estimation problem using a nonparametric empirical Bayes procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex, efron2019bayes}.
-We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator.
-Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis.
-Our compound decision approach is conceptually similar the empirical Bayes approach taken by \citet{dey2018corshrink} to estimate a high-dimensional correlation matrix.
+In this paper we define and explore one such class of estimators.
+This approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}.
+Section \ref{sec:compound} shows how this framing motivates vectorizing the covariance matrix and solving the resulting vector estimation problem using nonparametric empirical Bayes procedures studied in the compound decision literature \citep{jiang2009general, koenker2014convex, efron2019bayes}.
+The vector estimate can then be reassembled into matrix form and projected into the space of positive-definite matrices to give a final covariance matrix estimator.
+Surprisingly, though this vectorized approach essentially ignores the matrix structure, it can sometimes outperform a number of state-of-the-art proposals in simulations and a real data analysis.
+Our results suggest that vectorized approaches may be preferrable when the sample eigenvectors are poor estimates of the population eigenvectors.
+Our compound decision approach is closely related to the correlation matrix estimator of \citet{dey2018corshrink}.
 We discuss in detail the connections with their work in Section \ref{sec:proposed}.
-\blue{Need to add to numerical results}
 
 The article is organized as follows.
-In Section \eqref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach.
-In Section \eqref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset.
-Finally, Section \eqref{sec:discussion} concludes with a discussion.
-Our procedure is implemented in the R package \verb|cole|, available on GitHub.
-\blue{Is this implemented in the package yet?}
+In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed estimator.
+In Section \ref{numerical results} we illustrate its performance in simulations and a gene coexpression network inference problem.
+Finally, Section \ref{sec:discussion} concludes with a discussion.
+Our procedure is implemented in the R package \verb|cole|, available on GitHub at \url{http://github.com/sdzhao/cole}.
 
-\section{\label{sec:compound}Compound decision problem formulation}
-\subsection{\label{sec:background}Background}
+\section{\label{sec:compound}Covariance matrix estimation as a compound decision problem}
+\subsection{\label{sec:background}Problem formulation}
 
 Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$.
-A common measure of estimation performance is the scaled squared Frobenius risk
+This section argues that covariance matrix estimation can be viewed as a compound decision problem when estimation performance is measured using the scaled squared Frobenius risk
 \begin{equation}
@@
 where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.
-This paper constructs a new non-rotationally invariant $\bs{\delta}$ that in simulations and a data analysis has comparable or lower risk \eqref{frobenius risk} compared to existing methods.
+This compound decision formulation will motivate a new class of unstructured non-rotationally invariant estimators.
 
-Our proposed approach is motivated by the observation that minimizing \eqref{frobenius risk} is a type of compound decision problem, which we briefly review in this section.
-Introduced by \citet{robbins1951asymptotically}, compound decision problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$.
-Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk
+Compound decision problems involve the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$ \citep{robbins1951asymptotically}.
+Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk $ n^{-1} \sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))$, where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$.
+A classical example is the classical Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.
+
+Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can therefore be viewed as a compound decision problem where the goal is to simultaneously estimate the components of the vector
 \begin{equation}
-\label{compound risk}
-R(\bs{\theta},\bs{\delta}) = \frac{1}{n}\sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))
+  \label{eq:sigma}
+  \bs{\sigma} = (\sigma_{11}, \ldots, \sigma_{1p}, \ldots, \sigma_{p1}, \ldots,\sigma_{pp})^\top
 \end{equation}
-where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$.
-A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.
+under average squared error loss.
+One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite.
+Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}.
+As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite.
+This will in fact be true of one of the estimators we propose in Section \ref{sec:proposed}.
+To resolve this issue, we can project any estimate into the space of positive-definite matrices; see Section \ref{posdef}.
 
-A key property of compound decision problems is that while a given $Y_i$ or $\theta_i$ seems like it should offer no help for estimating $\theta_j$ when $j \ne i$, in fact borrowing information across all indices $i = 1, \ldots, n$ to estimate $\theta_j$ is superior to estimating each $\theta_j$ using the corresponding $Y_j$ alone.
-A classical example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$.
+The value of this formulation is that it allows us to apply ideas from the compound decision literature to covariance matrix estimation.
+A key property of compound decision problems is that while information about index $i$ seems like it should offer no help for estimating a parameter with a different index $j$, in fact borrowing information across all indices is superior to considering each index alone.
+An example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$.
 \citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$.
 A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.
+A natural next step is therefore to apply modern ideas for Gaussian sequence estimation to covariance matrix estimation, which we do in Section \ref{sec:method}.
 
-The discussion above shows that covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can be viewed as a compound decision problem.
-Furthermore, one special consequence of choosing \eqref{frobenius risk} as the risk measure is that the matrix estimation problem becomes equivalent to a vector estimation problem, specifically the problem of estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under average squared error loss.
-This interpretation is conceptually very similar to the classical Gaussian sequence problem.
+\subsection{\label{sec:connections}Reinterpretation of existing methods}
 
-We therefore propose to apply modern ideas for Gaussian sequence estimation to covariance matrix estimation.
-One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite.
-Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}.
-As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive semidefinite.
-This will in fact be true of one of the estimators we propose in Section \ref{sec:proposed}.
-To resolve this issue, we project the result estimate into the space of positive-definite matrices; see Section \ref{posdef}.
-
-\subsection{\label{sec:connections}Connections to existing work}
-
-Treating the covariance matrix estimation problem as a vector estimation compound decision problem may seem unintuitive, but many existing matrix estimation methods can also be interpreted as carrying out vector estimation.
-The simplest example is the sample covariance matrix $\bs{S} = \frac{1}{n}\bs{X}^{T}\bs{X}$, which takes this form because the $\bs{X}$ are assumed to have mean zero.
-This can be thought of as estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$ using maximum likelihood.
-Less trivially, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective.
+Treating the covariance matrix estimation problem as a vector estimation compound decision problem may seem unintuitive, but many existing matrix estimation methods can also be reinterpreted as carrying out vector estimation.
+The simplest example is the sample covariance matrix $\bs{S} = n^{-1} \bs{X}^{T}\bs{X}$, which takes this form because the $\bs{X}$ are assumed to have mean zero.
+As another example, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective.
 Their adaptive thresholding method applies a version of the soft thresholding method of \citet{donoho1995adapting}, originally developed to estimate a sparse mean vector in the Gaussian sequence problem, to each entry of the sample covariance matrix.
 
-Interestingly, we can also show that the celebrated linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as a solution to a vector estimation problem where the estimand is the $p^2 \times 1$ vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$.
-Their estimator is defined as
-\begin{equation}
-\label{linear model}
-\hat{\bs{\Sigma}}_{\text{LW}} = \left(1 - \frac{b_n^2}{d_n^2}\right) \bs{S} + \frac{b_n^2}{d_n^2}\hat{\mu} \bs{I}
-\end{equation} 
-for $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, and $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$, where $\Vert \cdot \Vert_F$ denotes the Frobenius norm.
+As a further illustration of the potential of our new formulation, we can show that the linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as the solution to a compound decision problem.
+As described in Section \ref{sec:background}, we first view the problem of minimizing the scaled Frobenius norm \eqref{frobenius risk} as simultaneously estimating the components of the vector $\bs{\sigma}$ \eqref{eq:sigma} under squared error.
+Then a standard approach in the compound decision literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes the compound risk.
+For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule \citep{fourdrinier2018shrinkage, stigler19901988}.
 
-To see this, we first restrict attention to the class of linear decision rules for estimating each component of $(\sigma_{11},\ldots,\sigma_{pp})^\top$, which we define to take the form
+Following this recipe, we also restrict attention to the class of linear decision rules where $\sigma_{jk}$ will be estimated by
 \begin{equation}
@@
 for some parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of $\bs{I}$.
-In other words, the estimate for $\sigma_{jk}$ is linear in $s_{jk}$.
-Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}, and the following proposition shows that it is possible to construct a good estimate of this risk for every fixed value of $\beta_S$ and $\beta_I$.
+Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}.
+Though we cannot directly calculate \eqref{frobenius risk} because it involves the unknown $\sigma_{jk}$, the following proposition shows that we can construct a good estimate of this risk for every fixed value of $\beta_S$ and $\beta_I$.
 
@@
     \label{eq:Rhat}
-    \hat{R}(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ].
+    \hat{R}_L(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ].
   \end{equation}
@@
   \[
-    \lim_{n \rightarrow 0} \{ \bb{E} \hat{R}(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta}) \} = 0
+    \lim_{n \rightarrow 0} \{ \bb{E} \hat{R}_L(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta}) \} = 0
   \]
@@
 
-Proposition \ref{prop:Rhat} shows that $\hat{R}(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}.
-It is then reasonable to estimate the best estimator in this class by minimizing $\hat{R}(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
-Proposition \ref{prop:linear} shows that this estimator is closely related to the \citet{ledoit2004well} estimator \eqref{linear model}.
+Proposition \ref{prop:Rhat} shows that $\hat{R}_L(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}.
+It is then reasonable to estimate the best estimator in this class by minimizing $\hat{R}_L(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$.
+A slight modification of this procedure turns out to recover the \citet{ledoit2004well} estimator
+\begin{equation}
+\label{eq:lw}
+\hat{\bs{\Sigma}}_{\text{LW}} = \left(1 - \frac{b_n^2}{d_n^2}\right) \bs{S} + \frac{b_n^2}{d_n^2}\hat{\mu} \bs{I},
+\end{equation} 
+where $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$, and $\Vert \cdot \Vert_F$ denotes the Frobenius norm.
 
@@
 \label{prop:linear}
-Let $\hat{\beta}_S$ and $\hat{\beta}_I$ denote the minimizers of $\hat{R}(\beta_S, \beta_I)$ and define the linear decision rule
+Let $\hat{\beta}_S$ and $\hat{\beta}_I$ denote the minimizers of $\hat{R}_L(\beta_S, \beta_I)$ and define the linear decision rule
 \begin{equation}
   \label{eq:opt_linear}
-  \hat{\delta}_{jk}(\bs{X}) = \max(\hat{\beta}_S, 0) s_{jk} + \min\{\hat{\beta}_I, \hat{\mu}\} u_{jk}.
+  \hat{\delta}_{jk}(\bs{X}) = \max(\hat{\beta}_S, 0) s_{jk} + \min(\hat{\beta}_I, \hat{\mu}) u_{jk}.
 \end{equation}
-Then $\hat{\delta}_{jk}(\bs{X}) = \hat{\sigma}_{\text{LW}jk}$, where $\hat{\sigma}_{\text{LW}jk}$ is the $jk$th entry of $\bs{\Sigma}_{\text{LW}}$ \eqref{linear model}.
+Then $\hat{\delta}_{jk}(\bs{X}) = \hat{\sigma}_{\text{LW}jk}$, where $\hat{\sigma}_{\text{LW}jk}$ is the $jk$th entry of $\bs{\Sigma}_{\text{LW}}$ \eqref{eq:lw}.
 \end{prop}
 
-Proposition \ref{prop:linear} therefore shows that the \citet{ledoit2004well} covariance matrix estimator can essentially be obtained as the solution to a pure vector estimation problem, without considering the matrix structure of the target vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$.
-There are interesting connections between the estimator in Proposition \ref{prop:linear} and classical estimators of a vector of normal means.
+There are also interesting connections between \eqref{eq:opt_linear} and classical estimators of a vector of normal means.
 First, instead of \eqref{eq:opt_linear}, consider the estimator $\tilde{\delta}_{jk}(\bs{X}) = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, which directly plugs the estimated $\hat{\beta}_S$ and $\hat{\beta}_I$ into \eqref{linear class}.
-It can be shown that  $\tilde{\delta}_{jk}(\bs{X})$ is analogous to the Efron-Morris estimator for the Gaussian sequence problem \citep{efron1973stein}.
-In other words, it can be shown that $\tilde{\delta}(\bs{X})$ is equivalent to estimating the vector $(\sigma_{11}, \ldots, \sigma_{pp})^\top$ by shrinking $(s_{11}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion, stigler19901988}.
+It can be shown that  $\tilde{\delta}_{jk}(\bs{X})$ is equivalent to estimating $\bs{\sigma}$ by shrinking the vector $(s_{11}, \ldots, s_{1p}, \ldots, s_{p1}, \ldots, s_{pp})^\top$ toward the one-dimensional subspace spanned by $(u_{11}, \ldots,  u_{1p}, \ldots, u_{p1}, \ldots, u_{pp})^\top$ \citep{biscarri2019thesis, lindley1962discussion, stigler19901988}.
 A major difference between $\hat{\delta}_{jk}$ \eqref{eq:opt_linear} and $\tilde{\delta}_{jk}$ is that $\hat{\beta}_S$ in the latter is replaced by $\max(\hat{\beta}_S, 0)$ in the former.
-This is equivalent to how the shrinkage factor in the classical James-Stein estimator is replaced by its positive part in the positive-part James-Stein estimator \citep{baranchik1964multiple}.
+This is similar to how replacing the shrinkage factor in the classical James-Stein estimator by its positive part results can improve its accuracy \citep{baranchik1964multiple}.
 
-\section{\label{sec:method}Method}
-\subsection{\label{sec:proposed}Proposed estimator}
+\section{\label{sec:method}Methods}
+\subsection{\label{sec:class}New class of estimators}
 
-Section \ref{sec:compound} argues that treating covariance matrix estimation as a vector estimation problem can be a fruitful strategy, as evidenced by the linear estimator exhibited in Proposition \ref{prop:linear}.
-This suggests that estimators that are nonlinear in $s_{jk}$ may have better performance.
+Proposition \ref{prop:linear} of Section \ref{sec:connections} suggests that applying ideas from the compound decision literature to estimating the vector $\bs{\sigma}$ \eqref{eq:sigma}  can be a fruitful strategy for covariance matrix estimation.
+However, Proposition \ref{prop:linear} considers only a linear decision rule, whereas more recent work in compound decision theory has focused on the much larger class of so-called separable rules \citep{brown2009nonparametric, jiang2009general, zhang2003compound}.
+In the standard compound decision problem of estimating $\bs{\theta} = E \bs{Y}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ for some function $t$ that does not depend on the index $i$ \citep{robbins1951asymptotically}.
+Linear rules such as \eqref{eq:opt_linear} are included in this class, so that the best separable rule should perform at least as well as the best linear rule.
 
-We propose to consider an even larger class of estimators, the class of so-called separable rules.
-In the standard compound decision problem of estimating $\bs{\theta}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ for some function $t$ that does not depend on the index $i$ \citep{robbins1951asymptotically}.
-Here we generalize this idea to the problem of estimating a vectorized matrix.
+Motivated by these ideas, here we propose a new class of covariance matrix estimators by generalizing the notion of separable rules to the problem of estimating a vectorized matrix.
 For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, we define the class of separable rules to be
@@
   \label{separable}
-  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
-   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
+  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
+   \delta_{jj} = t_d(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
 \end{equation}
 where $\bs{X}_{\cdot j} = (X_{1j}, \ldots, X_{nj})^\top$ is the vector of observed values of the $j$th feature.
-In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $\tilde{t}$ observations of the corresponding feature and off-diagonal entries using a function $t$ of observations from the two corresponding features, where $t$ and $\tilde{t}$ do not depend on the indices $j$ and $k$.
+In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $t_d$ and off-diagonal entries using $t_{od}$, where $t_{od}$ and $t_d$ do not depend on the indices $j$ and $k$.
 Furthermore, we enforce rule in $\mathcal{S}$ to give symmetric estimates of the off-diagonal entries.
@@
 
-This class of separable rules is reasonable to consider because it includes several common covariance estimators, including the sample covariance $(s_{11}, \ldots, s_{pp})^\top$, the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}, and the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
+This class of separable rules is reasonable to consider because it includes several common covariance estimators, including the sample covariance, the class of adaptive thresholding estimators for sparse covariance matrices studied by \citet{cai2011adaptive}, and the class of linear estimators \eqref{linear class} used by \citet{ledoit2004well}, which can be expressed as
 \begin{align*}
-t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}) &= \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n, \, 1\leq k<j\leq p,\\
-\tilde{t}(\bs{X}_{\cdot j})&=\beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot j} / n+\beta_I, \, j=1,\ldots,p.
+t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}) &= \beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot k} / n, \, 1\leq k<j\leq p,\\
+t_d(\bs{X}_{\cdot j})&=\beta_S \bs{X}_{\cdot j}^\top \bs{X}_{\cdot j} / n+\beta_I, \, j=1,\ldots,p.
 \end{align*}
@@
 
-We propose to search for the optimal estimator within $\mathcal{S}$.
-The optimal separable estimator $\bs{\delta}^\star$ that minimizes the scaled squared Frobenius risk \eqref{frobenius risk} over all rules in $\mathcal{S}$ will perform at least as well as the three estimators mentioned above, and may perform better.
-Targeting the optimal separable rule is standard in the compound decision literature \citep{zhang2003compound}.
+The class $\mathcal{S}$ constitutes a fundamentally different approach to covariance matrix estimation compared to the existing methods described in Section \ref{introduction}.
+Estimators in $\mathcal{S}$ do not assume that $\bs{\Sigma}$ has any particular structure and also are not necessarily rotationally invariant.
+Intuitively, they should perform better than rotationally invariant estimators when the sample eigenvectors of $\bs{S}$ are very different from the population eigenvectors.
+Results in Section \ref{numerical results} suggest that this is indeed the case.
 
-The optimal $\bs{\delta}^\star$ is an oracle estimator and cannot be calculated in practice, as the true risk is unknown.
-In the classical compound decision framework, empirical Bayes methods are used to estimate the oracle optimal separable rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}.
-We take a similar approach here.
-$f(\cdot \mid \bs{\eta}_{jk})$, the density of $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation.
-When $j\neq k$, $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ is comprised of $n$ independent mean-zero multivariate normals with covariance matrices
-\[
-\bs{C}_{jk}=\begin{bmatrix}
-&\sigma_j^2 & \sigma_j\sigma_k r_{jk} \\
-& \sigma_j\sigma_k r_{jk}  & \sigma_k^2
-\end{bmatrix}.
-\]
+\subsection{\label{sec:proposed}Proposed estimator}
 
-Now consider the Bayesian model for non-diagonal entries
+We propose to search for the optimal estimator $\bs{\delta}^\star$ within $\mathcal{S}$ \eqref{separable}.
+We can first provide a closed-form expression for $\bs{\delta}^\star$.
+Let $f_2(\cdot \mid a, b, \gamma)$ be the density of a bivariate normal centered at zero with standard deviations $a$ and $b$ and correlation $\gamma$, $f_1(\cdot \mid a)$ be the density of a univariate normal with mean zero and standard deviation $a$, and
 \begin{equation}
-  \label{bayesian nondiagonal}
-  (\bs{X},\bs{X}^{'}) \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
-  \quad
-  \bs{\eta} \sim G_{nd}(a,b, \gamma) =  \frac{2}{p(p-1)} \sum_{1\leq k<j\leq p} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
+  \label{eq:G}
+  \begin{aligned}
+    G_{od}(a, b, \gamma)
+    =\,&
+    \frac{1}{p (p - 1)} \sum_{1 \leq j, k \leq p, j \ne k} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma),\\
+    G_d(a)
+    =\,&
+    \int dG_{od}(a,b,\gamma) db d\gamma = \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a),
+  \end{aligned}
 \end{equation}
-and diagonal entries
-\begin{equation}
-  \label{bayesian diagonal}
-  \bs{X} \mid a \sim \widetilde{f}(\cdot \mid a),
-  \quad
-  a \sim G_d(a) =  \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a).
-\end{equation}
-the $G_d$ is the marginal density for the first parameter of $G_{nd}$
-\begin{equation}
-\label{eq:marginal}
-\int dG_{nd}(a,b,\gamma)dbd\gamma = dG_d(a)
-\end{equation}
-By the fundamental theorem of compound decisions \citep{robbins1951asymptotically, jiang2009general}, this is closely related to the vectorized covariance matrix estimation problem under Frobenius risk \eqref{frobenius risk}.
-Then we have the following proposition.
+where $\sigma_j$ and $\sigma_k$ are the true standard deviations of the $j$th and $k$th covariates and $r_{jk} = \sigma_{jk} / (\sigma_j \sigma_k)$ is their true correlation.
 
 \begin{prop}\label{prop:bayes risk}
-For any $\bs{\delta} \in \mathcal{S}$ \eqref{separable}, the Frobenius risk can be written as the convex combination of Bayesian risks with respect to priors $G_{nd}$ and $G_d$.
-\begin{align*}
-  R(\bs{\Sigma}, \bs{\delta})
+  The optimal separable estimator
+  \begin{equation}
+    \label{eq:opt}
+    \bs{\delta}^\star
+    =
+    \argmin_{\bs{\delta} \in \mathcal{S}} R(\bs{\Sigma}, \bs{\delta})
+    =
+    \left\{
+      \begin{array}{ll}
+        t_{od}^\star(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}), & 1 \leq k < j \leq p,\\
+        t_d^\star(\bs{X}_{\cdot j}), & j = 1, \ldots, p,
+      \end{array}
+    \right.
+  \end{equation}
+  obeys
+  \begin{equation}
+    \label{eq:tstar}
+    \begin{aligned}
+      t_{od}^\star(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})
       =\,&
-  \frac{1}{p^2} \{2\sum_{1\leq k<j\leq p}^p
-  \int \{t(\bs{X},\bs{X}^{\prime}) - \sigma_{jk} \}^2 f(t(\bs{X},\bs{X}^{\prime}) \mid \bs{\eta}_{jk}) d\bs{X}d\bs{X}^{\prime}\\
-  +\,&
-  \sum_{j=1}^p
-  \int \{\widetilde{t}(\bs{X}) - \sigma_{jj} \}^2 \widetilde{f}(\bs{X} \mid s_{jj}) d\bs{X}
-  \}\\
+      \frac{\int a b \gamma f_{2n}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k} \mid a, b, \gamma) dG_{od}(a, b, \gamma)}{\int f_{2n}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k} \mid a, b, \gamma) dG_{od}(a, b, \gamma)},\\
+      t_d^\star(\bs{X}_{\cdot j})
       =\,&
-  \frac{p-1}{p}\int \int \{t(\bs{X},\bs{X}^{\prime}) - g(\bs{\eta})\}^2 f((\bs{X},\bs{X}^{\prime})\mid \bs{\eta}) dG_{nd}(\bs{\eta}) d\bs{X}d\bs{X}^{\prime}\\
-  +\,&
-  \frac{1}{p}\int \int \{\widetilde{t}(\bs{X}) - a^2\}^2 \widetilde{f}(\bs{X} \mid a) dG_{d}(a) d\bs{X} \\
-  =\,&
-   \frac{p-1}{p}\bb{E}[\{t(\bs{X},\bs{X}^{\prime}) -  g(\bs{\eta})\}^2]+ \frac{1}{p}\bb{E}[\{\widetilde{t}(\bs{X}) -  a^2\}^2],
-\end{align*}
-where $g(a, b, \gamma) = a b \gamma$ and the final expectation is the Bayes risk of estimating $\sigma_{jk}$. 
+      \frac{\int a^2 f_{1n}(\bs{X}_{\cdot j} \mid a) dG_d(a)}{\int f_{1n}(\bs{X}_{\cdot j} \mid a) dG_d(a)}
+    \end{aligned}
+  \end{equation}
+  where
+  \begin{equation}
+    \label{eq:f}
+    \begin{aligned}
+      f_{2n}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k} \mid a, b, \gamma)
+      =
+      \prod_{i = 1}^n f_2(X_{ij}, X_{ik} \mid a, b, \gamma),
+      \quad
+      f_{1n}(\bs{X}_{\cdot j} \mid a)
+      =
+      \prod_{i = 1}^n f_1(X_{ij} \mid a).
+    \end{aligned}
+  \end{equation}
 \end{prop}
-The decision rule which minimizes the Frobenius risk depends on all the $\sigma_j$ and $r_{jk}$.
-The optimal oracle separable rule $\bs{\delta}^\star$ therefore has $jk$th entry equal to $\delta^\star_{jk}(\bs{X}) = t^\star((\bs{X}_{\cdot j},\bs{X}_{\cdot k}))$ for $j> k$ and $\delta^\star_{jj}(\bs{X}) = \widetilde{t}^\star(\bs{X}_{\cdot j})$, where $t^\star = \mathbb{E}\{g(\bs{\eta}) \mid (\bs{X},\bs{X}^{'})\}$ and $\widetilde{t}^\star = \mathbb{E}\{a^2 \mid \bs{X}\}$ minimizes the Bayes risk.
 
-Based on this result, we propose the following empirical Bayes procedure.
-We first use nonparametric maximum likelihood \citep{kiefer1956consistency} to estimate the priors $G_{nd}$ and $G_d$.
-Under the Bayesian model \eqref{bayesian nondiagonal} and \eqref{bayesian diagonal}, and the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent across $jk$, we estimate $G_{nd}$ using
+Proposition \ref{prop:bayes risk} illustrates the key property of compound decision problems: that borrowing information across indices can help estimate a single parameter.
+In this case, the optimal separable rule for estimating $\sigma_{jk}$ borrows from all of the $\sigma_j$ and $r_{jk}$.
+This rule also has an interesting Bayesian interpretation, as $t^\star_{od}$ and $t^\star_d$ are formally equivalent to posterior expectations when the $(\sigma_j, \sigma_k, r_{jk})$ are independently and identically distributed according to the prior $G_{od}$.
+However, the result in Proposition \ref{prop:bayes risk} holds even when the parameters are not random.
+Proposition \ref{prop:bayes risk} is a direct extension of the fundamental theorem of compound decisions \citep{robbins1951asymptotically, zhang2003compound}.
+
+The optimal $\bs{\delta}^\star$ \eqref{eq:opt} is an oracle estimator and cannot be calculated in practice.
+A common approach in the compound decision literature is to use empirical Bayes ideas to estimate the oracle optimal separable decision rule \citep{robbins1955empirical, zhang2003compound, brown2009nonparametric, jiang2009general, efron2014two, efron2019bayes}.
+We propose to estimate $\bs{\delta}^\star$ using what \citet{efron2014two} calls $g$-modeling, where we proceed as if the $(\sigma_j, \sigma_k, r_{jk})$ were truly random and then estimate the prior $G_{od}$ from the data.
+Specifically, under the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent for different $(j, k)$, we propose to calculate estimate $G_{od}$ and $G_d$ \eqref{eq:G} using nonparametric maximum likelihood \citep{kiefer1956consistency}:
 \begin{equation}
   \label{Gnd hat}
-  \hat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}_{nd}} \{\prod_{1\leq k\leq j<p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\} 
-  \{\prod_{j = 1}^p \int \widetilde{f}(\bs{X} \mid a) dG_d(a)\},
+  \hat{G}_{od} = \argmax_{G_{od} \in \mathcal{G}_{od}}
+  \prod_{1\leq k < j \leq p} \int f_{2n}(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid a, b, \gamma) dG_{od}(a, b, \gamma)
+  \prod_{j = 1}^p \int f_{1n} (\bs{X}_{\cdot j} \mid a) dG_d(a),
 \end{equation}
-where $\mathcal{G}_{nd}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$,   $G_d$ is determined by $G_{nd}$ as indicated in \eqref{eq:marginal}.
-Of course, the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are not independent, so $\hat{G}_{nd}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
-Using $\hat{G}_{nd}$ and $\hat{G}_d$, we estimate the vectorized $\bs{\Sigma}$ using
-$$ \hat{\bs{\delta}}(\bs{X})=
-  (\hat{t}(\bs{X}_{\cdot 2},\bs{X}_{\cdot 1}), \ldots,  \hat{t}(\bs{X}_{\cdot p},\bs{X}_{\cdot (p-1)}),\hat{t}(\bs{X}_{\cdot 1}),\ldots,\hat{t}(\bs{X}_{\cdot p}))$$
-  , where 
+where $\mathcal{G}_{od}$ is the family of all distributions supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$ and $G_d$ is determined by $G_{od}$ as indicated in \eqref{eq:G}.
+Of course, the $(\bs{X}_{\cdot j}, \bs{X}_{\cdot k})$ are not independent across different $(j, k)$ unless the true covariance $\bs{\Sigma}$ is diagonal, so $\hat{G}_{od}$ does not maximize a likelihood but rather a pairwise composite likelihood \citep{varin2011overview}.
+Using $\hat{G}_{od}$ and $\hat{G}_d$, we propose to estimate the vectorized $\bs{\Sigma}$ using
 \begin{equation}
   \label{proposed}
-  \hat{t}(\bs{X}_{\cdot j},\bs{X}_{\cdot k}) = \frac{\int g(\bs{\eta}) f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta})d\hat{G}_{nd}(\bs{\eta})}{\int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) d\hat{G}_{nd}(\bs{\eta})},
-  \quad 
-  \hat{t}(\bs{X}_{\cdot j}) = \frac{\int a^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a)d\hat{G}_d(a)}{\int \widetilde{f}(\bs{X}_{\cdot j} \mid a) d\hat{G}_d(a)}.
+  \hat{\bs{\delta}}(\bs{X})
+  =
+  (
+  \hat{t}_d(\bs{X}_{\cdot 1}), \ldots, \hat{t}_{od}(\bs{X}_{\cdot 1}, \bs{X}_{\cdot p}), \ldots, \hat{t}_{od}(\bs{X}_{\cdot p}, \bs{X}_{\cdot 1}), \ldots, \hat{t}_d(\bs{X}_{\cdot p})
+  )
 \end{equation}
-The $\hat{t}$ estimates the Bayes rule $t^\star$, $\widetilde{t}^\star$ and $\hat{\bs{\delta}}$ estimate the optimal oracle separable rule $\bs{\delta}^\star$.
+where $\hat{t}_{od}$ and $\hat{t}_d$ are obtained by plugging $\hat{G}_{od}$ and $\hat{G}_d$ \eqref{Gnd hat} into the expressions for $t_{od}^\star$ and $t_d^\star$ from \eqref{eq:tstar}.
 
-Our proposed procedure is an example of what \citet{efron2014two} calls $g$-modeling, an approach to empirical Bayes problems that proceeds by modeling the prior.
-A major advantage of nonparametric estimation of the prior is that it allows the data itself to determine how best to shrink the estimator.
-This kind of procedure is also adopted in \citet{dey2018corshrink}.
-In contrast, most existing methods shrink in a pre-determined direction, such as toward a diagonal matrix in the case of \citet{ledoit2004well}.
-Theoretical justification of our proposed $\hat{\bs{\delta}}$ is difficult and is discussed in Section \eqref{discussion}.
-Nevertheless, our numerical results in Section \eqref{numerical results} show that in practice, our $\hat{\bs{\delta}}$ can outperform many existing covariance matrix estimators.
+Our proposed procedure is similar to the method of \citet{dey2018corshrink}, who introduce an empirical Bayes approach to estimating a correlation matrix.
+They first apply Fisher's $Z$-transform to each off-diagonal entry of the sample correlation matrix and model the means of the transformed entries as independent random variables arising independently from some prior distribution.
+They then estimate the prior from the data, calculate the posterior expectations of the means, transform them back into a correlation matrix, and select the closest positive-definite matrix.
+
+The main difference between our approaches is that \citet{dey2018corshrink} do not shrink the variances while we do, as our $\hat{t}_d$ estimates $\sigma_j$ using its supposed posterior expectation assuming a prior distribution of $\hat{G}_d$.
+Furthermore, our Proposition \ref{prop:bayes risk} shows that our shrinkage approach can be interpreted as viewing the triples $(\sigma_j, \sigma_k, r_{jk})$ as if they were drawn from the trivariate prior $G_{od}$, which allows for possible dependencies between $r_{jk}$ and the $\sigma_j$.
+For example, model 2 in our simulations in Section \ref{numerical results} considers a covariance matrix where entries with larger standard deviations also tend to have larger correlations.
+In contrast to the method of \citet{dey2018corshrink}, our approach can learn this structure if it exists and take advantage of it for more accurate estimation.
 
@@
 
-Calculating the estimated prior $\hat{G}_{nd}$ \eqref{Gnd hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}$.
-\citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p^2$ points.
+Calculating the estimated prior $\hat{G}_{od}$ \eqref{Gnd hat} is difficult, as it is an infinite-dimensional optimization problem over the class of all probability distributions $\mathcal{G}_{od}$ supported on $\mathbb{R}_+ \times \mathbb{R}_+ \times [-1, 1]$.
+\citet{lindsay1983geometry} showed that the solution is atomic and is supported on at most $p(p+1)/2$ points.
 The EM algorithm has traditionally been used to estimate the locations of the support points and the masses at those points \citep{laird1978nonparametric}, but this is a difficult nonconvex optimization problem.
@@
 Instead, we maximize the pairwise composite likelihood over a fixed grid of support points, similar to recent $g$-modeling procedures for standard compound decision problems; this restores convexity \citep{jiang2009general, koenker2014convex, feng2018approximate}.
-Specifically, we assume that the prior for the $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$ is supported on $D$ fixed support points $\bs{\xi}_{\tau}$, $\tau=1,\ldots, D$.
-We can then use the EM algorithm to estimate the masses $\hat{\bs{w}}=\{\hat{w}_{1},\ldots, \hat{w}_{D}\}$ at those points via the iteration
-\begin{equation}
+Specifically, we assume that $G_{od}$ is supported on $D$ fixed support points $(a_\tau, b_\tau, \gamma_\tau)$, $\tau=1,\ldots, D$.
+Since $G_{od}$ is symmetric in its first two coordinates, we construct the support points so that $D$ is even and for $\tau > D / 2$, $(a_\tau, b_\tau, \gamma_\tau) = (b_{\tau - D / 2}, a_{\tau - D / 2}, \gamma_{\tau - D / 2})$.
+We also constrain $\hat{G}_{od}$ such that the estimated mass at the point $(a_\tau, b_\tau, \gamma_\tau)$ is equal to the mass at $(b_\tau, a_\tau, \gamma_\tau)$.
+Finally, we use the EM algorithm to estimate the masses $\hat{w}_\tau$.
+Letting $\hat{w}_\tau^{(k)}$ denote the estimate of the $\tau$th mass point as the $k$th iteration, it can be shown that for $\tau = 1, \ldots, D / 2$, the update formula is
+\begin{align*}
   \label{eq:iteration}
-\hat{w}_{\tau}^{(k)} = \frac{1}{p^2}\sum_{j, k = 1}^p \frac{\hat{w}_{\tau}^{(k-1)} f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_{\tau})}{\sum_{l=1}^D\hat{w}_l^{(k-1)} f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)}
-\end{equation}
-over $k$.
-Early stopping of the EM algorithm can be useful \citep{koenker2019comment}, and more sophisticated convex optimization procedures can be used as well \citep{koenker2014convex}.
-Our proposed estimator \eqref{proposed} then becomes
-\[
-\hat{t}(\bs{X}_j,\bs{X}_k) = \frac{ \sum_{\tau=1}^D g(\bs{\xi}_\tau) f(\bs{X}_j,\bs{X}_k \mid \bs{\xi}_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_\tau) \hat{w}_{\tau}},
-\quad
-\hat{t}(\bs{X}_{\cdot j}) = \frac{ \sum_{\tau=1}^D a_{\tau}^2 \widetilde{f}(\bs{X}_{\cdot j} \mid a_\tau) \hat{w}_{\tau}}{ \sum_{\tau=1}^D \widetilde{f}(\bs{X}_{\cdot j}  \mid a_\tau) \hat{w}_{\tau}}.
-\]
+      &
+      \hat{w}_{\tau}^{(k)} = \hat{w}_{\tau+D/2}^{(k)}\\
+  =\,&
+       \frac{2}{p(p+1)} \left[\sum_{j=1}^p \frac{\hat{w}_{\tau}^{(k-1)} \{f_{1n}(\bs{X}_j \mid a_{\tau})+f_{1n}(\bs{X}_j \mid a_{\tau+D/2})\}}{\sum_{l=1}^D \hat{w}_{l}^{(k-1)} f_{1n}(\bs{X}_j \mid a_{l})} \,+\right.\\
+      &
+          \left.
+\sum_{1 \leq k < j \leq p} \frac{\hat{w}_{\tau}^{(k-1)} \{f_{2n}(\bs{X}_j,\bs{X}_k \mid a_\tau, b_\tau, \gamma_\tau)+f_{2n}(\bs{X}_j,\bs{X}_k \mid a_{\tau+D/2}, b_{\tau+D/2}, \gamma_{\tau+D/2})\}}{\sum_{l=1}^{D}\hat{w}_l^{(k-1)} f_{2n}(\bs{X}_j,\bs{X}_k\mid a_l, b_l, \gamma_l)}\right].
+\end{align*}
+In practice, we center each $\bs{X}_{\cdot j}$ and instead of using the density functions $f_{1n}$ and $f_{2n}$ \eqref{eq:f}, we use the densities of their sufficient statistics: the sample variance $s_j^2$, in the case of $f_{1n}$, and the $2 \times 2$ matrices with diagonal entries $s_j^2$ and $s_k^2$ and off-diagonal entries equal to the sample covariances $s_{jk}$, in the case of $f_{2n}$.
 
-Ideally, the grid points should be chosen to densely cover the parameter space.
-However, the fact that $G_{nd}$ is multivariate poses difficulties, as for example using a grid of $d$ points in each dimension requires a total of $D = d^3$ grid points, which requires huge computational cost for even moderate $d$.
-Alternatively, we can use a so-called exemplar algorithm \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions $\hat{\bs{\eta}}_{jk}$ of the $\bs{\eta}_{jk}$.
-This reduces the size of the support set, but even in this case the computation complexity grows like $O(p^2)$.
+The support points need to be carefully chosen.
+Ideally, the points $(a_\tau, b_\tau, \gamma_\tau)$ would densely cover the parameter space.
+However, using a grid of $d$ points in each dimension would require a total of $D = d^3$ points, which would incur a huge computational cost for even moderate $d$.
+Alternatively, we could use the so-called exemplar method \citep{saha2020nonparametric}, which sets the support points to equal the observed sample versions of the $(\sigma_j, \sigma_k, r_{jk})$.
+This would reduce the size of the support set, but the computation complexity would still grow like $O(p^2)$.
 
-Here we propose a clustering-based exemplar algorithm to further improve computational efficiency.
-Let $s_j$, $s_k$, and $\gamma_{jk}$ be the sample variances and correlation between the $j$th and $k$th covariates.
-We first apply $K$-means clustering to identify $K$ clusters among the $p (p-1) / 2$ lower triangular off-diagonal sample points $(s_j,s_k,\gamma_{jk})$ and find their symmetric points by exchanging their first two dimensions.
-We then use the $2K$ cluster centroids as our support points.
-Figure \ref{fig:sim1_frobenius} shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table \eqref{tab:sim1_time} shows that they can be significantly faster.
+Instead, we use a clustering-based exemplar algorithm to further improve computational efficiency.
+Let $s_j$, $s_k$, and $\gamma_{jk}$ be the sample standard deviations and correlation between the $j$th and $k$th covariates.
+We first apply $K$-means clustering to identify $K$ cluster centroids using the $p (p-1) / 2$ lower triangular off-diagonal sample points $(s_j,s_k,\gamma_{jk})$.
+We then swap the first two coordinates of each point to construct another set of $K$ centroids.
+Finally, we use these $2K$ centroids as the support points for $\hat{G}_{od}$.
+We evaluate the impact of changing $K$ in Section \ref{numerical results}.
 
-In our implementation, we use the density function of $\bs{S}_{jk}=\begin{bmatrix}
-s_{jj} & s_{jk} \\
-s_{kj} & s_{kk} 
-\end{bmatrix}$ instead of $(\bs{X}_j,\bs{X}_k)$ since it is the sufficient statistics of $(\bs{X}_j,\bs{X}_k)$ with respect to $\bs{C}_{jk}$. Based on the property of sufficient statistics, 
-\begin{equation}
-\frac{f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_{\tau})}{f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_l)} = \frac{f(\bs{S}_{jk}\mid \bs{\xi}_{\tau})}{f(\bs{S}_{jk}\mid \bs{\xi}_l)}
-\end{equation}
-it is equivalent to use $f(\bs{X}_j,\bs{X}_k\mid \bs{\xi}_{\tau})$ instead to simplify the calculation.
-Although we assume $\bs{X}$ has mean of zero theoretically, in practice, we usually do not know the true means.
-In such case, we can centralize each $\bs{X}_{\cdot j}$ by their sample means.
+Our procedure is implemented in the R package \verb|cole|, available on GitHub.
+In our implementations of the EM algorithm in Section \ref{numerical results} we stopped iterating when the relative difference between two successive log-likelihoods was less than $10^{-4}$, up to a maximum of 200 iterations.
+Instead of EM, we attempted to maximize the composite likelihood \eqref{Gnd hat} using the interior point solver MOSEK, as advocated by \citet{koenker2014convex}, but we ran into computational difficulties because for even moderate sample sizes $n$, the values of the densities $f_{1n}$ and $f_{2n}$ at certain support points could be extremely small.
+These small values caused problems for the R wrapper for MOSEK.
+Our implementation of the EM algorithm was more robust to these small density values.
 
 \subsection{\label{posdef}Positive definiteness correction}
+
 Our proposed estimator \eqref{proposed} is not guaranteed to be positive-definite.
@@
 \citet{huang2017calibration} suggest $c_{\alpha}=10^{-\alpha}\lambda_{\min}^+$, where the parameter $\alpha$ is chosen to minimize $\Vert B - P_{c_{\alpha}}(B) \Vert + \alpha$ over a uniform partition of $\{\alpha_1,\ldots,\alpha_K\}$ of $[0,\alpha_K]$.
-In this paper we chose $K=20$ and $\alpha_K=10$.
+In our implementations we used $K=20$ and $\alpha_K=10$.
 
 \section{\label{numerical results}Numerical Results}
-In this section we refer to our approach using the abbreviation MSG: Matrix Shrinkage via $G$-modeling; we use MSGCor to refer to the version corrected for postive-definiteness.
-In our EM algorithm, the maximum number of iterations \eqref{eq:iteration} is 200 and it early stops when $\frac{L(\bs{w}^{(k+1)})-L(\bs{w}^{(k)})}{L(\bs{w}^{(k)})}\leq 10^{-4}$, where $L(\bs{w}^{(k)}) = \sum_{1\leq k\leq j\leq p}\log f(\bs{X}_{\cdot j},\bs{X}_{\cdot k}|\bs{w}^{(k)})$ is composite log likelihood of the observations.
-\subsection{\label{models}Models}
 
-We considered five models for the population covariance matrix.
-For the first four settings, $\bs{\Sigma} = \text{diag}(\bs{s}) \bs{C} \text{diag}(\bs{s})$, where $\bs{C}$ is correlation matrix and $\bs{s}$ is a vector of standard deviations.
+\subsection{\label{models}Covariance matrix models}
+
+We considered six models for the population covariance matrix to explore the behavior of our proposed method in simulations.
+In each model, we generated certain parameters randomly but then fixed them across all replications of our simulations.
+For readability, below we give each model a short name that we will refer to in the following figures and captions.
+
 \begin{itemize}
-\item Model 1. The standard deviations were independently generated from $\mathcal{U}(1,1.5)$ and the correlation matrix followed Model 2 of \citet{cai2011adaptive}:
+\item Model 1 (sparse): we generated a sparse $\bs{\Sigma}$ where most features were not correlated.
+We let half of the standard deviations equal 1 and the other half equal 1.5.
+We set the correlation matrix following Model 2 of \citet{cai2011adaptive}:
   \[
-  \bs{C}=
   \begin{pmatrix}&\bs{A}_1 & \bs{0} \\ &\bs{0} &\bs{I}_{p/2\times p/2}\end{pmatrix},
   \]
-  where the $jk$th entry of $\bs{A}_1$ is $a_{jk} = \max(1- \vert j - k \vert / 10, 0)$. This setting modeled a sparse covariance matrix.
+  where the $jk$th entry of $\bs{A}_1$ is $a_{jk} = \max(1- \vert j - k \vert / 10, 0)$.
 
-\item Model 2. The first $p / 2$ standard deviations equaled 1, the last $p / 2$ equaled 2, and the correlation matrix was
+\item Model 2 (hypercorrelated): we generated a $\bs{\Sigma}$ that we call ``hypercorrelated'' because its parameters are in a sense dependent on each other, as we let larger standard deviations $\sigma_j$ and $\sigma_k$ correspond to larger $r_{jk}$.
+Specifically, we let the first $p / 2$ standard deviations equal 1, the last $p / 2$ equal 2, and the correlation matrix equal
   \[
-  \bs{C} = \begin{pmatrix} & \bs{C}_{11} & \bs{C}_{12} \\  & \bs{C}_{21} & \bs{C}_{22}\end{pmatrix},
+    \begin{pmatrix} & \bs{C}_{11} & \bs{C}_{12} \\  & \bs{C}_{21} & \bs{C}_{22}\end{pmatrix},
   \]
   where $\bs{C}_{11}$ and $\bs{C}_{22}$ were $p/2 \times p/2$ compound symmetric matrices with correlation parameters 0.8 and 0.2, respectively, and $\bs{C}_{12}$ and $\bs{C}_{21}$ were $p/2 \times p/2$ matrices with entries equal to 0.4.
-This model was designed such that larger $\sigma_j$ and $\sigma_k$ tended to correspond to larger $r_{jk}$.
+As discussed in Section \ref{sec:proposed}, we expect that our proposed approach will be able to take advantage of the dependencies between the $\sigma_j$ and the $r_{jk}$, unlike the empirical Bayes approach of \citet{dey2018corshrink} that does not shrink standard deviations.
 
-\item Model 3.
-The standard deviations were generated independently from $\mathcal{U}(1, 1.5)$ and $\bs{C}$ was a compound symmetric matrix with correlation parameter 0.7.
-This modeled a dense covariance matrix.
+\item Model 3 (dense-0.7): we generated a $\bs{\Sigma}$ where the correlation between all features was 0.7, in contrast to our sparse Model 1.
+Otherwise, we generated the standard deviations as in Model 1.
 
-\item Model 4.
-This setting was the same as Model 3 except with correlation parameter 0.9.
-This high level of dependence tested the robustness of the pairwise composite likelihood estimator \eqref{Gnd hat}.
+\item Model 4 (dense-0.9): this setting was the same as Model 3 except with correlation parameter 0.9.
+This high level of dependence tests the robustness of the pairwise composite likelihood estimator \eqref{Gnd hat}.
 
-\item Model 5.
-With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $p / 2$ equaled 1 and the last $p / 2$ equaled 4.
-This followed simulation settings from \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
+\item Model 5 (orthogonal): instead of generating $\bs{\Sigma}$ by specifying its correlation matrix, we randomly generated a $\bs{U}$ from the Haar measure on the space of all orthonormal matrices and a vector of eigenvalues $\bs{l}$ independently generated from $\mathcal{U}(1, 4)$.
+We then let $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$.
+This simulation setting was used in \citet{lam2016nonparametric} and \citet{ledoit2019quadratic}.
 
-\item Model 6.
-With $\bs{U}$ a randomly generated orthonormal matrix, $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p -3$ entries equaled 1.
+\item Model 6 (spiked): this setting was the same as model 5 except that $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p - 3$ entries equaled 1, so that $\bs{\Sigma}$ was a spiked covariance matrix.
 \end{itemize}
 
-In each scenario, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$.
-Although $\bb{E}\bs{X}=0$ in our setting, we assume it is unknown and use $\bs{S}=\frac{1}{n-1}(\bs{X}-\bs{\bar{X}})^\top(\bs{X}-\bs{\bar{X}})$.
-We generated $200$ replicates and reported median errors under the following Frobenius norms, where $\hat{\bs{\Sigma}}$ is the estimated matrix with entries $\hat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$:
-$$\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_F = \{ \sum_{j,k = 1}^p (\hat{\sigma}_{jk} - \sigma_{jk})^2 \}^{1/2}$$
-which is a version of \eqref{frobenius risk}.
-%\item Spectral: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_2 = \lambda_{\max}(\hat{\bs{\Sigma}} - \bs{\Sigma})$, the largest eigenvalue of $\hat{\bs{\Sigma}} - \bs{\Sigma}$, and
-%
-%\item Matrix $\ell_1$: $\Vert \hat{\bs{\Sigma}} - \bs{\Sigma} \Vert_{L_1} = \max_{k = 1, \ldots, p} \sum_{j = 1}^p \vert \hat{\sigma}_{jk} - \sigma_{jk} \vert$.
+\subsection{\label{compared}Methods compared}
+
+We compared several high-dimensional covariance matrix estimation procedures.
+
+\begin{itemize}
+
+\item Our proposed estimator \eqref{proposed} and its positive-definiteness-corrected version described in Section \ref{posdef}.
+In the following figures and captions we refer to these as MSG (Matrix Shrinkage via $G$-modeling) and MSGCor, respectively.
+
+\item Adap: the adaptive thresholding method of \citep{cai2011adaptive} for sparse covariance matrices, which applies soft thresholding to entries of the sample covariance matrix.
+The threshold method is adaptive to the entry's variance and involves a tuning parameter.
+We fixed the parameter at 2, as recommended.
+
+\item Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \eqref{eq:lw}.
+
+\item QIS: the Quadratic-Inverse Shrinkage estimator of \citet{ledoit2019quadratic}, a recently developed nonlinear shrinkage method.
+QIS performs linear shrinkage on the sample eigenvalues of the covariance matrix in inverse eigenvalue space.
+A bandwidth parameter is required, which we chose following the paper's recommendation.
+
+\item NERCOME: the Nonparametric Eigenvalue-Regularized COvariance Matrix Estimator of \citet{lam2016nonparametric}.
+This nonlinear shrinkage method randomly splits the samples into two groups, one for estimating eigenvectors and the other for estimating eigenvalues.
+Combining the estimates gives a matrix.
+Following \citet{lam2016nonparametric}, we repeated this procedure 50 times and took the final covariance matrix estimator to be the average of the individual matrices.
+
+\item CorShrink: the empirical Bayes correlation matrix estimation method of \citet{dey2018corshrink}.
+To estimate $\bs{\Sigma}$, we used CorShrink to estimate the correlation matrix and then scaled its entries using the sample standard deviations.
+
+\item Sample: the sample covariance matrix.
+\end{itemize}
+
+In addition to the above data-driven estimators, we also implemented the two following oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$.
+
+\begin{itemize}
+\item OracNonlin: the optimal rotation-invariant covariance estimator, derived in \citet{ledoit2019quadratic}, of $\bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$.
+
+\item OracMSG: we implemented our proposed estimator \eqref{proposed} except that the support points are generated by clustering the true parameters $(\sigma_j,\sigma_k,r_{jk})$ instead of their sample versions.
+\end{itemize}
 
 \subsection{\label{optimalK}Clustering-based exemplar algorithm}
-We first studied the behavior of our $K$-means clustering-based exemplar algorithm for different $K$, described in Section \eqref{implementation}.
-For a given $p$, we let $K = rp$ for different ratios ratios $r=2,1,0.5,0.25$.
-We compared these choices for $K$ to the full exemplar method.
-For all these estimators, we show the result after applying positive-definiteness correction.
+
+We first explored the consequences of choosing different numbers of cluster centroids in our clustering-based exemplar algorithm described in Section \ref{implementation}.
+For each model, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$.
+Although $\bb{E}\bs{X}=0$ in our setting, we assume it is unknown and use $\bs{S}=\frac{1}{n-1}(\bs{X}-\bs{\bar{X}})^\top(\bs{X}-\bs{\bar{X}})$.
+We generated $200$ replicates and reported median errors under the loss function
+\begin{equation}
+  \label{eq:sim_loss}
+  \frac{1}{p} \left\{ \sum_{j,k = 1}^p (\hat{\sigma}_{jk} - \sigma_{jk})^2 \right\}^{1/2},
+\end{equation}
+where $\hat{\bs{\Sigma}}$ is the estimated matrix with entries $\hat{\sigma}_{jk}$ and $\bs{\Sigma}$ is the true matrix with entries $\sigma_{jk}$.
+We then fit our proposed method, after positive-definitness correction, for different $K$, where we let $K = rp$ for $r=2,1,0.5,0.25$.
 
@@
 \begin{center}
-\centerline{\includegraphics[width=0.85\textwidth]{img/sim1_frobenius.pdf} }
+\centerline{\includegraphics[width=0.85\textwidth]{img/sim1_frobenius.pdf}}
 \end{center}
-\caption{Average Frobenius norm errors over 200 replications. The Sparse, Block, Dense, Dense2, Orth and Spiked panels correspond to Models 1 through 5, respectively.}
+\caption{Median Frobenius norm errors over 200 replications for our proposed MSGCor.
+Exact numerical results and interquartile ranges are provided in the Supporting Information.
+Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.}
 \label{fig:sim1_frobenius}
@@
 \begin{center}
-\caption{\label{tab:sim1_time} Average running time for different ratios.}
+\caption{\label{tab:sim1_time} Average running time of MSGCor under Model 1 for different $K$.}
 %\setlength{\tabcolsep}{3pt} % Default value: 6pt
-\begin{tabular}{lrrr}
+\begin{tabular}{rrrr}
 \Hline
@@
 \hline
-Exemplar method   & 0.1357          & 6.0338        & 91.8942       \\
+Full example method   & 0.1357          & 6.0338        & 91.8942       \\
 $K=2p$	      & 0.0539 	     & 1.1883	   & 10.7687         \\
@@
 
-Figure \ref{fig:sim1_frobenius} presents the Frobenius norm error estimates from Model 1 to Model 6.
-Table \eqref{tab:sim1_time} shows the running time only for Model 1, because the running time does not vary much across different models.
-The results show that different $K$ exhibit similar performance and are comparable to the full exemplar method.
+Figure \ref{fig:sim1_frobenius} shows that across all models, different $K$ give similar estimation accuracies compared to using all sample points in the exemplar algorithm.
+Table \ref{tab:sim1_time} shows that they can be significantly faster; we only provide results from Model 1 because the times did not vary much across different models.
 Letting $K = p$ seemed to provide a good balance between accuracy and speed, so we implement our proposed method with $K = p$ in the rest of this paper.
 
-\subsection{\label{compared}Methods compared}
+\subsection{\label{sec:accuracies}Estimation accuracies}
 
-In this subsection, we compared MSG and MSGCor to several existing high-dimensional covariance matrix estimation methods:
-
-\begin{itemize}
-\item Sample: the sample covariance matrix.
-
-\item Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \eqref{linear model}.
-
-\item QIS: the Quadratic-Inverse Shrinkage estimator of \citet{ledoit2019quadratic}, a recently developed nonlinear shrinkage method.
-QIS performs linear shrinkage on the sample eigenvalues of the covariance matrix in inverse eigenvalue space.
-A bandwidth parameter is required, which we choose following the paper's recommendation.
-
-\item NERCOME: the Nonparametric Eigenvalue-Regularized COvariance Matrix Estimator of \citet{lam2016nonparametric}.
-This nonlinear shrinkage method randomly splits the samples into two groups, one for estimating eigenvectors and the other for estimating eigenvalues.
-Combining the estimates gives a matrix.
-Following the article, we repeated this procedure 50 times and took the final covariance matrix estimator to be the average of the individual matrices.
-
-\item Adap: the adaptive thresholding method of \citep{cai2011adaptive} for sparse covariance matrices, which applies soft thresholding to entries of the sample covariance matrix.
-The threshold method is adaptive to the entry's variance and involves a tuning parameter.
-We fixed the parameter at 2, as recommended.
-
-\item CorShrink: Empirical Bayes shrinkage estimation of correlation matrix \citep{dey2018corshrink}.
-\end{itemize}
-
-In addition to the above estimators, we also implemented the two following oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$.
-\begin{itemize}
-\item OracNonlin: the optimal rotation-invariant covariance estimator, defined in \citet{ledoit2019quadratic}, with $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$.
-The sample covariance, the linear shrinkage estimator of \citet{ledoit2004well}, and the nonlinear shrinkage estimators QIS and NERCOME are all rotation-invariant.
-
-\item OracMSG: It equals our proposed estimator \eqref{proposed} except sample grid points are cluster centers of true parameters $(\sigma_j,\sigma_k,r_{jk})$.
-\end{itemize}
+We next compared the estimation accuracies of the methods in Section \ref{compared} under loss \eqref{eq:sim_loss}, following the same simulation scheme as in Section \ref{optimalK}.
+The results are visualized in Figure \ref{fig:sim2_frobenius}.
+Our proposed methods could outperform existing methods, especially in Models 1 through 4.
+This pattern also held when we ran simulations with $p = 1000$ in the Supporting Information.
+The corrected version consistently outperformed the uncorrected version.
+Our methods could outperform the CorShrink method of \citet{dey2018corshrink} in many cases, likely because it was able to accurately shrink the standard deviations.
 
@@
 \end{center}
-\caption{Average Frobenius norm errors over 200 replications.
-The Sparse, Block, Dense, Dense2, and Orth panels correspond to Models 1 through 5, respectively.
-In MSG, $K$-means clustering is applied with $K=p$}
+\caption{Median Frobenius norm errors over 200 replications.
+Exact numerical results and interquartile ranges are provided in the Supporting Information.
+MSG and MSGCor were implemented with $K = p$.
+Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.}
 \label{fig:sim2_frobenius}
@@
 
-Figure \ref{fig:sim2_frobenius} presents the Frobenius loss for different estimators.
-Our MSG methods had the lowest or near-lowest errors across all settings except for Model 4 which has high correlations 0.9.
-This is not surprising because our method assumed independence of $\bs{A}_{jk}$.
-In some cases, for example in Models 1 and 2, the improvement was substantial.
-Model 2 was especially interesting because the standard deviation and correlations were related.
-Our proposed empirical Bayes estimator was able to capture this dependence in its estimate of the prior $G_{nd}$ \eqref{bayesian nondiagonal} and leverage it to provide much more accurate estimates.
-The nonlinear shrinkage estimators very slightly outperformed MSG in Model 5.
-In every setting, correcting MSG for positive-definiteness never increased the risk and decreased the risk in some cases.
-We also did experiments for Spectral norm and Matrix $\ell_1$.
-The results for this two norms are very similar to Frobenius norm.
-One exception is Adap has the lowest error in terms of Matrix $\ell_1$ norm in Model 1 because of its sparsity.
-Though our estimator was motivated in terms of the Frobenius norm error, it performed extremely well in terms of the other two norms as well.
+MSG and MSGCor were especially good in Model 2, where the standard deviation and correlations were related, because it was able to capture this dependence in its estimate of the prior $G_{od}$ \eqref{eq:G} and leverage it to provide much more accurate estimates.
+It was substantially better than CorShrink in this setting, further illustrating the advantage of shrinking the standard deviations.
+In the Supporting Information we also compare MSG and MSGCor to CorShrink for estimating correlation instead of covariance matrices.
+We find that in that case, CorShrink performs slightly better than MSG and MSGCor except in Model 2, likely because the additional flexibility that our method trades off lower bias for higher variance.
 
-Finally, the simulations show that the class of separable estimators \eqref{separable} proposed in this paper is fundamentally different from the class of rotation-invariant estimators, as the oracle optimal estimators in these two classes behave very differently.
-For example, the oracle separable estimator had vanishing risk in Model 2, while the oracle rotiation-invariant estimator does not.
-Separable estimators seemed better for Models 1 and 2 while rotation-invariant estimators were superior in Models 3 and 4.
-They seem comparable in Model 5.
+However, in Models 5 and 6 MSG and MSGCor were worse than the best of the other methods, though they were still comparable.
+In the Supporting Information we found this pattern was exacerbated when $p = 1000$.
+This observation is consistent with the behavior of the oracle estimator OracMSG, which was dramatically better than the oracle rotationally-invariant estimator OracNonlin in Models 1 through 4 but was comparable to it in Models 5 and 6.
+As discussed in Section \ref{sec:class}, this suggests that the population eigenvectors may be closer to the sample eigenvectors in Models 5 and 6.
+To explore this possibility, we recorded the Frobenius norms of the differences between the matrices of sample and population eigenvectors in each of our simulation settings.
+The results in Figure \ref{fig:eigenvectors} seem to support our hypothesis that our class of separable estimators is best-suited for estimating covariance matrices whose sample and population eigenvalues differ greatly.
 
-\subsection{Data analysis}
+\begin{figure}
+\begin{center}
+\centerline{  \includegraphics[width=0.95\textwidth]{img/eigenvectors.png}}
+\end{center}
+\caption{Estimation errors of sample eigenvectors over 200 replications. Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.}
+\label{fig:eigenvectors}
+\end{figure}
+
+\section{Data analysis}
 \label{gene analysis}
+
 Covariance matrix estimation is often used to reconstruct gene networks \citep{markowetz2007inferring}.
-We applied our MSG and the other covariance matrix estimators described in Section \eqref{compared} to gene network estimation using data from a small round blue-cell tumor microarray experiment \citep{khan2001classification}, which was also studied by \citet{cai2011adaptive}.
-\citet{osareh2009classification} report the expression of 2308 genes from 63 samples from four groups: 12 neuroblastoma, 20 rhabdomyosarcoma, 8 Burkitt lymphoma, and 23 Ewing's sarcoma patients.
-In MSG, the clustering parameter $K$ is still set as $p=200$.
-We followed the same data preprocessing as \citet{cai2011adaptive} and sorted the genes in decreasing order according to their $F$-statistic
-\begin{equation}
-F = \frac{1}{k-1}\sum_{m=1}^kn_m(\overline{x}_m - \overline{x})^2 /   \frac{1}{n-k}\sum_{m=1}^k (n_m-1)\hat{\sigma}_m^2 
-\end{equation}
-where $k = 4$ is the number of patient categories, $n_m$, $\overline{x}_m$, and $\hat{\sigma}_k$ represent the sample size, sample mean, and sample variance of the gene's expression in the $m$th category, respectively, and $\overline{x}$ is the global mean.
-We proceeded with the top 40 genes and bottom 160 genes.
+We used our proposed methods and the other covariance matrix estimators described in Section \ref{compared} for gene network estimation in a study of the transcriptomic response to social aggression in various brain regions in mouse \citep{saul2017transcriptional}.
+As part of the control condition, mice were exposed to a paper cup and bulk RNA-sequencing data was collected from the amygdala, frontal cortex, and hypothalamus at 30, 60, and 90 minutes after exposure.
+Data were collected from five mice at each time point, but two of the 30-minute hypothalamus samples were sequenced using a different library preparation than the other samples and so were left out of this analysis.
+The data are available from the NCBI Gene Expression Omnibus (GEO; \url{http://www.ncbi.nlm.nih.gov/geo/}) under accession number GSE80345.
 
-We applied various methods to estimate the covariance matrix of these 200 genes.
-To measure the accuracy of the estimators, we split the 63 samples into two subsets $\bs{X}_1$ and $\bs{X}_2$, ensuring that each subset consisted of the same number of subjects from each of the four disease groups.
-After centering the variables to have zero mean, we used $\bs{X}_1$ to calculate covariance matrix estimates and compared these to the sample covariance matrix $\bs{S}_2$ of $\bs{X}_2$, which served as a proxy for the unknown true covariance matrix.
-We measured the errors using the Frobenius, spectral, and matrix $\ell_1$ norms.
-We repeated this process 200 times.
+Our goal was to estimate brain region-specific gene networks in these mice, assuming that the networks were the same across time points.
+Understanding how these networks differ between brain regions may provide insight into the different functions of the different regions.
+We built coexpression networks of the top 200 genes that were most differentially expressed between the regions.
+To identify these genes, we first considered only genes that were expressed in each of our samples and had at least one count per million mapped reads in at least 3 samples.
+We then performed a Kruskal-Wallis test and retained the 200 genes with the lowest p-values.
 
-Table \eqref{tab:tab1} reports the average errors across the replications.
-Our MSG methods had the lowest average error.
-The positive-definiteness correction slightly reduced the risk as well.
-The linear shrinkage estimate was almost as accurate, but the other methods were much less accurate.
-These results suggest that our estimator can perform well in realistic settings, where the mean-zero multivariate normal distributional assumption on the data may not be met.
+We then converted the gene expression values to log-counts per million mapped reads and applied the methods from Section \ref{compared} to estimate the covariance matrix of these 200 genes in each brain region.
+When implementing our $K$-means clustering-based exemplar algorithm, we let $K = 400$.
+We also only implemented MSGCor, since it was the best performer in our simulations.
 
@@
 \begin{center}
-\caption{\label{tab:tab1} Average gene expression covariance matrix estimation errors. Bold entries highlight the smallest errors in each column.}
+\caption{\label{tab:tab1}Median gene expression covariance matrix estimation errors (25\% and 75\% quantiles in parentheses). Bold text highlights the smallest median errors in each column.}
 %\setlength{\tabcolsep}{3pt} % Default value: 6pt
-\begin{tabular}{lrrr}
+\begin{tabular}{rll}
   \Hline
-            & Frobenius & Spectral & Matrix $\ell_1$ \\
+  Brain region & Amygdala & Frontal cortex \\
   \hline
-MSG   & 24.13 (2.54)         & 13.14(2.96)         & 46.53(8.49)         \\
-MSGCor & \textbf{24.04(2.56)} & \textbf{13.12(2.94)}& \textbf{46.00(8.67)}         \\
-Adap        & 28.93(2.03)          & 17.54(3.85)         & 49.66(8.47)         \\
-Linear      & 24.49(2.53)         & 13.65(3.14)         & 49.46(9.04)         \\
-QIS         & 28.76(2.79)          & 16.09(3.35)         & 58.73(9.74)         \\
-NERCOME     & 24.52(2.52)          & 13.19(3.07)         & 48.27(9.67)         \\
-Sample      & 28.43(2.54)          & 16.21(3.15)         & 56.08(9.89)         \\
+  MSGCor & \textbf{2.26 (1.98, 2.57)} & \textbf{2.3 (2.15, 2.55)} \\
+  Adap & 2.6 (2.18, 2.97) & 2.44 (2.18, 2.71) \\
+  Linear & 2.3 (2.05, 2.55) & 2.35 (2.19, 2.53)\\
+  QIS & 2.61 (2.33, 2.86) & 2.79 (2.64, 2.93)\\
+  NERCOME & 2.37 (2.14, 2.62) & \textbf{2.3 (2.14, 2.53)}\\
+  CorShrink & \textbf{2.26 (2, 2.56)} & 2.36 (2.2, 2.56)\\
+  Sample & 2.61 (2.33, 2.85) & 2.79 (2.64, 2.95)\\
   \hline
-
 \end{tabular}
@@
 
-In addition to comparing the numerical accuracies, we also investigated whether our estimator gave qualitatively different gene networks compared to the other approaches.
-First, Figure \ref{network} illustrates the covariance matrices in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects.
-To avoid completely connected graphs, we sparsified the matrix estimates by thresholding the smaller entries of each matrix to zero.
-Since the adaptive thresholding method of \citet{cai2011adaptive} naturally produced a sparse estimated matrix, we thresholded the other matrix estimates to match the sparsity level of the \citet{cai2011adaptive} estimate.
+To measure the accuracy of the estimators, we randomly split the samples into a training and testing set.
+We estimated the covariance matrix on the training set and measured accuracy using the Frobenius norm of the difference between the estimated matrix and the sample covariance matrix calculated from the test set.
+We set the training sample size equal to 10, which left five samples in the test sets for the amygdala and frontal cortex and only three samples in the test set for hypothalamus.
+We therefore dropped hypothalamus from the analysis because we felt that three samples was unsuitable for calculating a surrogate of the true covariance matrix.
+We repeated this process 100 times.
+Table \eqref{tab:tab1} reports the median errors and interquartile ranges across the replications.
+Many of the different methods had comparable performance, and the different brain regions favored different methods.
+Nevertheless, our proposed MSGCor had among the best performances in both regions.
 
@@
   \begin{center}
-\centerline{ \includegraphics[width=0.85\textwidth]{img/network.pdf}}
+    \includegraphics[width=0.9\textwidth]{img/networks_A.png}
   \end{center}
-\caption{Gene networks recovered by the different covariance matrix estimation methods.}
-\label{network}
+\caption{Amygdala gene networks recovered by the different covariance matrix estimation methods.}
+\label{network_A}
 \end{figure}
-
-The results show several interesting features.
-First, there appear to be two major clusters, which are disconnected in every estimated network except for the one produced by the adaptive thresholding approach.
-Second, the larger cluster appears to contain two sub-clusters, and this finer structure was only recovered by MSG and QIS, and to a lesser extent the linear shrinkage estimator and NERCOME.
-Finally, the nodes in the networks estimated by QIS and NERCOME appear to be clustered more tightly together compared to in the other networks.
-These observations suggest that MSG produces qualitatively different networks, in addition to lower estimation errors.
-
-Finally, we also compared the estimated degrees of the genes in the different networks.
-For each estimated network, we ordered the 200 genes by degree and then selected the top 20\%, denoting this set $J_k$ for the $k$th network.
-For each pair of networks $k$ and $k'$, we calculated the similarity between their most connected genes using Jaccard index $\vert J_k \cap J_{k'} \vert / \vert J_k \cup J_{k'} \vert$.
-Figure \ref{top20} visualizes these similarities.
-Interestingly, however, among all estimators, they were also the most similar to the unbiased sample covariance matrix.
-Together with the above results, this indicates that MSG may simultaneously give the lowest error and, at least in terms of degree estimation, the most unbiased results.
 
@@
   \begin{center}
-\centerline{\includegraphics[width=0.85\textwidth]{img/top20.pdf}}
+    \includegraphics[width=0.9\textwidth]{img/networks_FC.png}
   \end{center}
-\caption{Similarities of gene degrees between the estimated networks. Each number reports the Jaccard index between the top 20\% most connected genes of each pair of networks.}
-\label{top20}
+\caption{Frontal cortex gene networks recovered by the different covariance matrix estimation methods.}
+\label{network_FC}
 \end{figure}
+
+In addition to comparing the numerical accuracies, we also investigated whether our estimator gave qualitatively different gene networks compared to the other approaches.
+Figures \ref{network_A} and \ref{network_FC} illustrate the estimated amygdala and frontal cortex covariance matrices in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects.
+To avoid completely connected graphs, we sparsified the matrix estimates by thresholding the smaller entries of each matrix to zero.
+Since the adaptive thresholding method of \citet{cai2011adaptive} naturally produced a sparse estimated matrix, we thresholded the other matrix estimates to match the sparsity level of the \citet{cai2011adaptive} estimate.
+The network representations suggest that the resulting networks look fairly similar except for the ones produced by linear shrinkage and NERCOME.
+That our procedure produces networks that look similar to the unbiased sample covariance matrix networks is comforting.
 
@@
 
-The class of separable covariance matrix estimators \eqref{separable} that we proposed in this paper appears to be very promising.
-Many existing procedures already explicitly or implicitly target this class, and our proposed estimate \eqref{proposed} of the optimal separable estimator outperforms a number of existing covariance matrix estimators.
-This is surprising because our approach vectorizes the matrix and therefore cannot take matrix structure, such as positive-definiteness, into account.
+The class of separable covariance matrix estimators \eqref{separable} that we explored in this paper appears to be promising.
+This is somewhat surprising because our approach vectorizes the matrix and therefore cannot take matrix structure, such as positive-definiteness, into account.
 This suggests that a vectorized approach combined with a positive-definiteness constraint may have improved performance.
 The resulting estimator would necessarily not be separable, because the estimate of the $jk$th entry would depend on more than just the $j$th and $k$th observed features, so the $g$-modeling estimation strategy is insufficient.
-More work is needed.
 
-Though our estimator performs well in simulations and in real data, providing theoretical guarantees is difficult.
+Providing theoretical guarantees for our estimator is difficult.
 In the standard mean vector estimation problem with $Y_i \sim N(\theta_i, 1)$, \citet{jiang2009general} showed that an empirical Bayes estimator based on a nonparametric maximum likelihood estimate of the prior on the $\theta_i$ can indeed asymptotically achieve the same risk as the oracle optimal separable estimator.
@@
 
-Finally, we have so far assumed that our data are multivariate normal.
+Finally, we have so far assumed that our data are multivariate normal, though simulations in the Supporting Information suggest that our procedures can still provide relatively accurate estimates when the normality assumption is violated.
 To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gnd hat} and in our proposed estimator \eqref{proposed}.
@@
 \section*{Acknowledgments}
-We thank Dr. Roger Koenker for his valuable comments.\vspace*{-8pt}
+We thank Dr. Roger Koenker and three anonymous referees for their very valuable comments.\vspace*{-8pt}
 
@@
 \section*{Supporting Information}
-Web Appendices, Tables, and Figures referenced in Section 3 are available with this paper at the Biometrics website on Wiley Online Library.
+Web Appendices, Tables, and Figures are available with this paper at the Biometrics website on Wiley Online Library.
 
@@
 
-\subsection{Proof of Proposition \eqref{prop:Rhat}}
-\begin{proof}
+\subsection{Proof of Proposition \ref{prop:Rhat}}
 Using the fact that when $\bb{E}\bs{X}=\bs{0}$, $\bb{E}s_{jk} = \sigma_{jk}$ for all $j,k=1,\ldots,p$.
@@
 \begin{align*}
-  \bb{E} \hat{R}(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta})
+  \bb{E} \hat{R}_L(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta})
   =\,&
@@
 where the last result follows because by assumption, $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$.
-\end{proof}
 
-\subsection{Proof of Proposition \eqref{prop:linear}}
+\subsection{Proof of Proposition \ref{prop:linear}}
 
-\begin{proof}
-The estimator in \citet{ledoit2004well} is
-\begin{equation} 
-\label{eq:lw}
-\bs{\Sigma}_{LW} = (\frac{d_n^2-b_n^2}{d_n^2})_+\bs{S}+ \min\{1,\frac{b_n^2}{d_n^2}\}\hat{\mu}\bs{I}
-\end{equation}
-where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{pn^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.
+% The estimator in \citet{ledoit2004well} is
+% \begin{equation} 
+% \label{eq:lw}
+% \bs{\Sigma}_{LW} = (\frac{d_n^2-b_n^2}{d_n^2})_+\bs{S}+ \min\{1,\frac{b_n^2}{d_n^2}\}\hat{\mu}\bs{I}
+% \end{equation}
+% where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{pn^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.
 
-  We first rewrite the risk estimate $\hat{R}(\beta_S, \beta_I)$.
+  We first rewrite the risk estimate $\hat{R}_L(\beta_S, \beta_I)$.
 Define $\bs{M}=(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$.
-Then the unbiased risk estimator can be re-written as
+Then the risk estimate can be re-written as
   \[
- p^2 \hat{R}(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
+ p^2 \hat{R}_L(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
   \]
@@
   \begin{align*}
-    \hat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\hat{R}(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
+    \hat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\hat{R}_L(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
     &
@@
 
-  We will need to show $\hat{\mu} b_n^2 / d_n^2 \to \hat{\beta}_I$ and $\hat{\beta}_I/\hat{\mu} + \hat{\beta}_S = 1$. Since
+  Since
 $$\bs{Z}^\top \bs{Z} = \begin{pmatrix}
@@
 \end{align*}
-The second component of $\hat{\bs{\beta}}$ equals $\hat{\beta}_I$, so
+
+The second component of $\hat{\bs{\beta}}$ equals $\hat{\beta}_I$, and
 \begin{align*}
@@
 &=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)/p \} / d_n^2
-=\hat{\mu}\frac{b_n^2}{d_n^2}.
+=\hat{\mu}\frac{b_n^2}{d_n^2},
 \end{align*}
-which is very closed to the coefficient of $\bs{I}$ in \eqref{eq:lw}
-Furthermore,
+so $\min(\hat{\mu}, \hat{\beta}_I) = \hat{\mu} b_n^2 / d_n^2$, which is the coefficient of the second term in the Ledoit and Wolf estimator \eqref{eq:lw}. Furthermore,
 \begin{align*}
@@
 =
-1.
+1,
 \end{align*}
-\end{proof}
+so $\min(\hat{\beta}_S, 0) = 1 - b_n^2 / d_n^2$, which is the coefficient of the first term in \eqref{eq:lw}.
 
-\subsection{Proof of Proposition \eqref{prop:bayes risk}}
+\subsection{Proof of Proposition \ref{prop:bayes risk}}
+
+For any $\bs{\delta} \in \mathcal{S}$ \eqref{separable}, the Frobenius risk \eqref{frobenius risk} can be written as
+\begin{align*}
+  R(\bs{\Sigma}, \bs{\delta})
+  =\,&
+       \frac{2}{p^2} \sum_{1\leq k<j\leq p}^p
+       \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - \sigma_{jk} \}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid \sigma_j, \sigma_k, r_{jk}) d\bs{X}d\bs{X}^{\prime} \,+\\
+  &
+       \frac{1}{p^2}
+       \sum_{j=1}^p
+       \int \{t_d(\bs{X}) - \sigma_{jj} \}^2 f_{1n}(\bs{X} \mid s_{jj}) d\bs{X}\\
+  =\,&
+       \frac{p-1}{p}\int \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - a b \gamma \}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) d\bs{X}d\bs{X}^{\prime} \, +\\
+     &
+       \frac{1}{p}\int \int \{t_d(\bs{X}) - a^2\}^2 f_{1n}(\bs{X} \mid a) dG_{d}(a) d\bs{X}.
+\end{align*}
+Therefore for any $t_{od}$,
+\begin{align*}
+  &
+    \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - a b \gamma \}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \\
+  =\,&
+       \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - t_{od}^\star(\bs{X}, \bs{X}^\prime)\}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \,+\\
+  &
+    2 \{t_{od}(\bs{X},\bs{X}^{\prime}) - t_{od}^\star(\bs{X}, \bs{X}^\prime)\} \int 
+  \{t_{od}^\star(\bs{X}, \bs{X}^\prime) - a b \gamma \}
+  f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \,+\\
+  &
+    \int \{t_{od}^\star(\bs{X}, \bs{X}^\prime) - a b \gamma\}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \\
+  \geq\,&
+          \int \{t_{od}^\star(\bs{X}, \bs{X}^\prime) - a b \gamma\}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma),
+\end{align*}
+because the middle cross term is equal to zero by the definition of $t^\star_{od}$. Similarly, it can be shown that for any $t_d$,
+\begin{align*}
+  \int \{t_d(\bs{X}) - a^2\}^2 f_{1n}(\bs{X} \mid a) dG_{d}(a)
+  \geq
+  \int \{t_d^\star(\bs{X}) - a^2\}^2 f_{1n}(\bs{X} \mid a) dG_{d}(a).
+\end{align*}
+This implies that $R(\bs{\Sigma}, \bs{\delta}) \geq R(\bs{\Sigma}, \bs{\delta}^\star)$ for any $\bs{\delta} \in \mathcal{S}$.
 

### 2021-12-07 msg.tex
small changes for manuscript and response of reviews

@@
   \]
-  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}.
+  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}. $\Vert \cdot \Vert_F$ denotes the Frobenius norm $\Vert \bs{A} \Vert_F^2=\sum_{j,k=1}^p A_{jk}^2$.
 \end{prop}
@@
 \end{equation} 
-where $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$, and $\Vert \cdot \Vert_F$ denotes the Frobenius norm.
+where $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$.
 
@@
 =
-\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert
+\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert_F
 =
@@
 \]
-where $\Vert \cdot \Vert$ denotes the Frobenius norm, $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues.
+where $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues.
 
@@
   \label{near posdef}
-  P_0(\bs{B})
+  P_c(\bs{B})
   =
@@
 Finally, we have so far assumed that our data are multivariate normal, though simulations in the Supporting Information suggest that our procedures can still provide relatively accurate estimates when the normality assumption is violated.
-To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gnd hat} and in our proposed estimator \eqref{proposed}.
-If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.
+To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f_1$ and $f_2$ in the nonparametric maximum compositive likelihood problem \eqref{Gnd hat} and in our proposed estimator \eqref{proposed}.
+If $f_1$ and $f_2$ are unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.
 

### 2022-03-07 msg.tex
change of to for
biometrics second revision, updated to satisfy biometrics editor comments only

@@
 For example, in Section \ref{gene analysis} we study brain region-specific gene networks in mouse using bulk RNA-sequencing data.
-
 A common approach to estimating gene expression networks is to use the standard sample covariance or corrlation matrices \citep{langfelder2008wgcna, zhang2005general}.
-However, these estimators behave poorly in the high-dimensional regime.
-To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrices.
-These can roughly be divided into two groups, according to whether they impose assumptions of the structure of population covariance matrix.
+However, these estimators behave poorly in the high-dimensional regime, and there are roughly two types of approaches to address this issue.
 
@@
 This approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}.
-Section \ref{sec:compound} shows how this framing motivates vectorizing the covariance matrix and solving the resulting vector estimation problem using nonparametric empirical Bayes procedures studied in the compound decision literature \citep{jiang2009general, koenker2014convex, efron2019bayes}.
-The vector estimate can then be reassembled into matrix form and projected into the space of positive-definite matrices to give a final covariance matrix estimator.
+Section \ref{sec:compound} shows how this framing motivates vectorizing the covariance matrix and solving the resulting vector estimation problem using nonparametric empirical Bayes procedures studied in the compound decision literature.
+The vector estimate can then be reassembled into matrix form and projected into the space of positive-definite matrices to give a final estimate.
 Surprisingly, though this vectorized approach essentially ignores the matrix structure, it can sometimes outperform a number of state-of-the-art proposals in simulations and a real data analysis.
-Our results suggest that vectorized approaches may be preferrable when the sample eigenvectors are poor estimates of the population eigenvectors.
 Our compound decision approach is closely related to the correlation matrix estimator of \citet{dey2018corshrink}.
@@
 In Section \ref{numerical results} we illustrate its performance in simulations and a gene coexpression network inference problem.
-Finally, Section \ref{sec:discussion} concludes with a discussion.
-Our procedure is implemented in the R package \verb|cole|, available on GitHub at \url{http://github.com/sdzhao/cole}.
+Section \ref{sec:discussion} concludes with a discussion, and proofs can be found in the Supporting Information.
 
@@
 Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$.
-This section argues that covariance matrix estimation can be viewed as a compound decision problem when estimation performance is measured using the scaled squared Frobenius risk
+A common measure of estimation performance is the scaled squared Frobenius risk
 \begin{equation}
@@
 where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.
-This compound decision formulation will motivate a new class of unstructured non-rotationally invariant estimators.
 
-Compound decision problems involve the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$ \citep{robbins1951asymptotically}.
+Minimizing \eqref{frobenius risk} can be viewed as a compound decision problem.
+These involve the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$ \citep{robbins1951asymptotically}.
 Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk $ n^{-1} \sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))$, where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$.
@@
 
-Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can therefore be viewed as a compound decision problem where the goal is to simultaneously estimate the components of the vector
+Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} is thus a compound decision problem where the goal is to simultaneously estimate the components of
 \begin{equation}
@@
 As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite.
-This will in fact be true of one of the estimators we propose in Section \ref{sec:proposed}.
 To resolve this issue, we can project any estimate into the space of positive-definite matrices; see Section \ref{posdef}.
@@
 A key property of compound decision problems is that while information about index $i$ seems like it should offer no help for estimating a parameter with a different index $j$, in fact borrowing information across all indices is superior to considering each index alone.
-An example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$.
-\citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$.
-A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.
-A natural next step is therefore to apply modern ideas for Gaussian sequence estimation to covariance matrix estimation, which we do in Section \ref{sec:method}.
+An example of this phenomenon is the James-Stein shrinkage estimator \citep{james1961estimation}% , which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$.
+\citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$
+, and a long line of subsequent work has led to much more sophisticated and accurate procedures \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.
+In Section \ref{sec:method} we apply these ideas to covariance matrix estimation,
 
@@
 
-Treating the covariance matrix estimation problem as a vector estimation compound decision problem may seem unintuitive, but many existing matrix estimation methods can also be reinterpreted as carrying out vector estimation.
+Treating covairance matrix estimation as a vector estimation problem may seem unintuitive, but many existing matrix estimation methods can also be reinterpreted as carrying out vector estimation.
 The simplest example is the sample covariance matrix $\bs{S} = n^{-1} \bs{X}^{T}\bs{X}$, which takes this form because the $\bs{X}$ are assumed to have mean zero.
 As another example, \citet{cai2011adaptive} studied sparse high-dimensional covariance matrices and explicitly appealed to the vector perspective.
-Their adaptive thresholding method applies a version of the soft thresholding method of \citet{donoho1995adapting}, originally developed to estimate a sparse mean vector in the Gaussian sequence problem, to each entry of the sample covariance matrix.
+Their adaptive thresholding method applies a version of the soft thresholding method of \citet{donoho1995adapting}, originally developed to estimate the mean of a Gaussian sequence, to each entry of the sample covariance matrix.
 
 As a further illustration of the potential of our new formulation, we can show that the linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as the solution to a compound decision problem.
-As described in Section \ref{sec:background}, we first view the problem of minimizing the scaled Frobenius norm \eqref{frobenius risk} as simultaneously estimating the components of the vector $\bs{\sigma}$ \eqref{eq:sigma} under squared error.
-Then a standard approach in the compound decision literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes the compound risk.
-For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule \citep{fourdrinier2018shrinkage, stigler19901988}.
-
-Following this recipe, we also restrict attention to the class of linear decision rules where $\sigma_{jk}$ will be estimated by
+A standard approach in this literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes the compound risk % .
+For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule
+\citep{fourdrinier2018shrinkage, stigler19901988}. Following this recipe, we consider the class of linear decision rules where $\sigma_{jk}$ will be estimated by
 \begin{equation}
@@
   \]
-  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}. $\Vert \cdot \Vert_F$ denotes the Frobenius norm $\Vert \bs{A} \Vert_F^2=\sum_{j,k=1}^p A_{jk}^2$.
+  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}, where $\Vert \bs{A} \Vert_F$ denotes the Frobenius norm $\bs{A}$.
 \end{prop}
+
 The condition on $\bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ requires that the variances of the entries of $\bs{S}$ do not grow too quickly as $n$ grows.
 It is also required by \citet{ledoit2004well} and is implied by their Lemma 3.1.
-
 Proposition \ref{prop:Rhat} shows that $\hat{R}_L(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}.
@@
 
-There are also interesting connections between \eqref{eq:opt_linear} and classical estimators of a vector of normal means.
+% There are also interesting connections between \eqref{eq:opt_linear} and classical estimators of a vector of normal means.
 First, instead of \eqref{eq:opt_linear}, consider the estimator $\tilde{\delta}_{jk}(\bs{X}) = \hat{\beta}_S s_{jk} + \hat{\beta}_I u_{jk}$, which directly plugs the estimated $\hat{\beta}_S$ and $\hat{\beta}_I$ into \eqref{linear class}.
@@
 However, Proposition \ref{prop:linear} considers only a linear decision rule, whereas more recent work in compound decision theory has focused on the much larger class of so-called separable rules \citep{brown2009nonparametric, jiang2009general, zhang2003compound}.
-In the standard compound decision problem of estimating $\bs{\theta} = E \bs{Y}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ for some function $t$ that does not depend on the index $i$ \citep{robbins1951asymptotically}.
+% In the standard compound decision problem of estimating $\bs{\theta} = E \bs{Y}$ using $\bs{Y}$, a separable decision rule $\bs{\delta}(\bs{Y})$ is one where $\delta_i(\bs{Y}) = t(Y_i)$ for some function $t$ that does not depend on the index $i$ \citep{robbins1951asymptotically}.
 Linear rules such as \eqref{eq:opt_linear} are included in this class, so that the best separable rule should perform at least as well as the best linear rule.
 
-Motivated by these ideas, here we propose a new class of covariance matrix estimators by generalizing the notion of separable rules to the problem of estimating a vectorized matrix.
-For decision rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, we define the class of separable rules to be
+We generalize these ideas to propose a new class of covariance matrix estimators.
+For rules $\bs{\delta}(\bs{X}) = (\delta_{11}(\bs{X}),\ldots,\delta_{pp}(\bs{X}))$ that estimate $(\sigma_{11}, \ldots, \sigma_{pp})^\top$, we define the class of separable estimators to be
 \begin{equation}
   \label{separable}
-  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
+  \mathcal{S} = \{\bs{\delta} :
+  \quad
+  \delta_{kj} = \delta_{jk} = t_{od}(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p;
+  \quad 
    \delta_{jj} = t_d(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
@@
 In other words, rules in $\mathcal{S}$ estimate diagonal entries of $\bs{\Sigma}$ using a function $t_d$ and off-diagonal entries using $t_{od}$, where $t_{od}$ and $t_d$ do not depend on the indices $j$ and $k$.
-Furthermore, we enforce rule in $\mathcal{S}$ to give symmetric estimates of the off-diagonal entries.
+Furthermore, we enforce rules in $\mathcal{S}$ to give symmetric estimates of the off-diagonal entries.
 
@@
 \end{align*}
-Furthermore, whereas these three existing estimators are ultimately all functions of the $s_{jk}$, the class $\mathcal{S}$ allows for much more general rules that can be any function of the observations $\bs{X}_{\cdot j}$ and $\bs{X}_{\cdot k}$.
+Furthermore, whereas these three existing estimators are ultimately all functions of the $s_{jk}$, the class $\mathcal{S}$ allows for much more general rules that can be any function of the $\bs{X}_{\cdot j}$ and $\bs{X}_{\cdot k}$.
+Finally, the class $\mathcal{S}$ constitutes a fundamentally different approach to covariance matrix estimation compared to the existing methods described in Section \ref{introduction}.
+Estimators in $\mathcal{S}$ do not assume that $\bs{\Sigma}$ has any particular structure and also are not necessarily rotationally invariant.
 
-The class $\mathcal{S}$ constitutes a fundamentally different approach to covariance matrix estimation compared to the existing methods described in Section \ref{introduction}.
-Estimators in $\mathcal{S}$ do not assume that $\bs{\Sigma}$ has any particular structure and also are not necessarily rotationally invariant.
-Intuitively, they should perform better than rotationally invariant estimators when the sample eigenvectors of $\bs{S}$ are very different from the population eigenvectors.
+ % Intuitively, they should perform better than rotationally invariant estimators when the sample eigenvectors of $\bs{S}$ are very different from the population eigenvectors.
 Results in Section \ref{numerical results} suggest that this is indeed the case.
@@
 We first apply $K$-means clustering to identify $K$ cluster centroids using the $p (p-1) / 2$ lower triangular off-diagonal sample points $(s_j,s_k,\gamma_{jk})$.
-We then swap the first two coordinates of each point to construct another set of $K$ centroids.
-Finally, we use these $2K$ centroids as the support points for $\hat{G}_{od}$.
-We evaluate the impact of changing $K$ in Section \ref{numerical results}.
+We then swap the first two coordinates of each point to construct another set of $K$ centroids and use these $2K$ centroids as the support points for $\hat{G}_{od}$.
+Section \ref{numerical results} evaluates the impact of changing $K$.
 
-Our procedure is implemented in the R package \verb|cole|, available on GitHub.
 In our implementations of the EM algorithm in Section \ref{numerical results} we stopped iterating when the relative difference between two successive log-likelihoods was less than $10^{-4}$, up to a maximum of 200 iterations.
@@
 
-\begin{itemize}
-\item Model 1 (sparse): we generated a sparse $\bs{\Sigma}$ where most features were not correlated.
+Model 1 (sparse): we generated a sparse $\bs{\Sigma}$ where most features were not correlated.
 We let half of the standard deviations equal 1 and the other half equal 1.5.
-We set the correlation matrix following Model 2 of \citet{cai2011adaptive}:
-  \[
-  \begin{pmatrix}&\bs{A}_1 & \bs{0} \\ &\bs{0} &\bs{I}_{p/2\times p/2}\end{pmatrix},
-  \]
-  where the $jk$th entry of $\bs{A}_1$ is $a_{jk} = \max(1- \vert j - k \vert / 10, 0)$.
+We set the correlation matrix following Model 2 of \citet{cai2011adaptive}, a block-diagonal matrix where the $jk$th entry of the first $p/2 \times p/2$ block was $\max(1- \vert j - k \vert / 10, 0)$ and the second $p/2 \times p/2$ block was an identity matrix.
 
-\item Model 2 (hypercorrelated): we generated a $\bs{\Sigma}$ that we call ``hypercorrelated'' because its parameters are in a sense dependent on each other, as we let larger standard deviations $\sigma_j$ and $\sigma_k$ correspond to larger $r_{jk}$.
-Specifically, we let the first $p / 2$ standard deviations equal 1, the last $p / 2$ equal 2, and the correlation matrix equal
-  \[
-    \begin{pmatrix} & \bs{C}_{11} & \bs{C}_{12} \\  & \bs{C}_{21} & \bs{C}_{22}\end{pmatrix},
-  \]
-  where $\bs{C}_{11}$ and $\bs{C}_{22}$ were $p/2 \times p/2$ compound symmetric matrices with correlation parameters 0.8 and 0.2, respectively, and $\bs{C}_{12}$ and $\bs{C}_{21}$ were $p/2 \times p/2$ matrices with entries equal to 0.4.
+Model 2 (hypercorrelated): we generated a $\bs{\Sigma}$ that we call ``hypercorrelated'' because its parameters are in a sense dependent on each other, as we let larger standard deviations $\sigma_j$ and $\sigma_k$ correspond to larger $r_{jk}$.
+Specifically, we let the first $p / 2$ standard deviations equal 1, the last $p / 2$ equal 2, and the correlation matrix have four $p/2 \times p/2$ blocks.
+The diagonal blocks were compound symmetric matrices with correlation parameters 0.8 and 0.2 and the off-diagonal blocks had entries equal to 0.4.
 As discussed in Section \ref{sec:proposed}, we expect that our proposed approach will be able to take advantage of the dependencies between the $\sigma_j$ and the $r_{jk}$, unlike the empirical Bayes approach of \citet{dey2018corshrink} that does not shrink standard deviations.
 
-\item Model 3 (dense-0.7): we generated a $\bs{\Sigma}$ where the correlation between all features was 0.7, in contrast to our sparse Model 1.
+Models 3 and 4 (dense-0.7 and dense-0.9): we generated a $\bs{\Sigma}$ where the correlation between all features were 0.7 and 0.9, in contrast to our sparse Model 1.
 Otherwise, we generated the standard deviations as in Model 1.
 
-\item Model 4 (dense-0.9): this setting was the same as Model 3 except with correlation parameter 0.9.
-This high level of dependence tests the robustness of the pairwise composite likelihood estimator \eqref{Gnd hat}.
-
-\item Model 5 (orthogonal): instead of generating $\bs{\Sigma}$ by specifying its correlation matrix, we randomly generated a $\bs{U}$ from the Haar measure on the space of all orthonormal matrices and a vector of eigenvalues $\bs{l}$ independently generated from $\mathcal{U}(1, 4)$.
+Model 5 (orthogonal): instead of generating $\bs{\Sigma}$ by specifying its correlation matrix, we randomly generated a $\bs{U}$ from the Haar measure on the space of all orthonormal matrices and a vector of eigenvalues $\bs{l}$ independently generated from $\mathcal{U}(1, 4)$.
 We then let $\bs{\Sigma} = \bs{U}^T\text{diag}(\bs{l})\bs{U}$.
@@
 
-\item Model 6 (spiked): this setting was the same as model 5 except that $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p - 3$ entries equaled 1, so that $\bs{\Sigma}$ was a spiked covariance matrix.
-\end{itemize}
+Model 6 (spiked): this setting was the same as model 5 except that $\bs{l}$ was a vector of eigenvalues where the first $3$ entries were 4,3,2 and the remaining $p - 3$ entries equaled 1, so that $\bs{\Sigma}$ was a spiked covariance matrix.
 
@@
 We compared several high-dimensional covariance matrix estimation procedures.
+Several require tuning parameters, which we fixed based on the original papers' recommendations: (a) Our proposed estimator \eqref{proposed} and its positive-definiteness-corrected version described in Section \ref{posdef}.
+In the following figures and captions we refer to these as MSG (Matrix Shrinkage via $G$-modeling) and MSGCor, respectively; (b) Adap: the adaptive thresholding method of \citet{cai2011adaptive} for sparse covariance matrices; (c) Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \eqref{eq:lw}; (d) QIS: the Quadratic-Inverse Shrinkage estimator of \citet{ledoit2019quadratic}, which performs linear shrinkage on the sample eigenvalues in inverse eigenvalue space; (e) NERCOME: the Nonparametric Eigenvalue-Regularized COvariance Matrix Estimator of \citet{lam2016nonparametric}, which randomly splits the samples into two groups to estimate eigenvectors and eigenvalues separately; (f) CorShrink: the empirical Bayes correlation matrix estimation method of \citet{dey2018corshrink}, where we estimated $\bs{\Sigma}$ by using CorShrink to estimate the correlation matrix and then scaling its entries using the sample standard deviations; and (g) Sample: the sample covariance matrix.
 
-\begin{itemize}
-
-\item Our proposed estimator \eqref{proposed} and its positive-definiteness-corrected version described in Section \ref{posdef}.
-In the following figures and captions we refer to these as MSG (Matrix Shrinkage via $G$-modeling) and MSGCor, respectively.
-
-\item Adap: the adaptive thresholding method of \citep{cai2011adaptive} for sparse covariance matrices, which applies soft thresholding to entries of the sample covariance matrix.
-The threshold method is adaptive to the entry's variance and involves a tuning parameter.
-We fixed the parameter at 2, as recommended.
-
-\item Linear: the linear shrinkage estimator of \citet{ledoit2004well} given in \eqref{eq:lw}.
-
-\item QIS: the Quadratic-Inverse Shrinkage estimator of \citet{ledoit2019quadratic}, a recently developed nonlinear shrinkage method.
-QIS performs linear shrinkage on the sample eigenvalues of the covariance matrix in inverse eigenvalue space.
-A bandwidth parameter is required, which we chose following the paper's recommendation.
-
-\item NERCOME: the Nonparametric Eigenvalue-Regularized COvariance Matrix Estimator of \citet{lam2016nonparametric}.
-This nonlinear shrinkage method randomly splits the samples into two groups, one for estimating eigenvectors and the other for estimating eigenvalues.
-Combining the estimates gives a matrix.
-Following \citet{lam2016nonparametric}, we repeated this procedure 50 times and took the final covariance matrix estimator to be the average of the individual matrices.
-
-\item CorShrink: the empirical Bayes correlation matrix estimation method of \citet{dey2018corshrink}.
-To estimate $\bs{\Sigma}$, we used CorShrink to estimate the correlation matrix and then scaled its entries using the sample standard deviations.
-
-\item Sample: the sample covariance matrix.
-\end{itemize}
-
-In addition to the above data-driven estimators, we also implemented the two following oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$.
-
-\begin{itemize}
-\item OracNonlin: the optimal rotation-invariant covariance estimator, derived in \citet{ledoit2019quadratic}, of $\bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$.
-
-\item OracMSG: we implemented our proposed estimator \eqref{proposed} except that the support points are generated by clustering the true parameters $(\sigma_j,\sigma_k,r_{jk})$ instead of their sample versions.
-\end{itemize}
+In addition to the above data-driven estimators, we also implemented two oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$: (a) OracNonlin: the optimal rotation-invariant covariance estimator, derived in \citet{ledoit2019quadratic}, of $\bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$; and (b) OracMSG: we implemented our proposed estimator \eqref{proposed} except that the support points are generated by clustering the true parameters $(\sigma_j,\sigma_k,r_{jk})$ instead of their sample versions.
 
@@
 Exact numerical results and interquartile ranges are provided in the Supporting Information.
-Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.}
+Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.
+This figure appears in color in the electronic version of this article, and any mention of color refers to that version.}
 \label{fig:sim1_frobenius}
@@
 Table \ref{tab:sim1_time} shows that they can be significantly faster; we only provide results from Model 1 because the times did not vary much across different models.
-Letting $K = p$ seemed to provide a good balance between accuracy and speed, so we implement our proposed method with $K = p$ in the rest of this paper.
+Letting $K = p$ seemed to provide a good balance between accuracy and speed, so we implement our proposed method with $K = p$ in this paper.
 
@@
 MSG and MSGCor were implemented with $K = p$.
-Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.}
+Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.
+This figure appears in color in the electronic version of this article, and any mention of color refers to that version.}
 \label{fig:sim2_frobenius}
@@
 This observation is consistent with the behavior of the oracle estimator OracMSG, which was dramatically better than the oracle rotationally-invariant estimator OracNonlin in Models 1 through 4 but was comparable to it in Models 5 and 6.
-As discussed in Section \ref{sec:class}, this suggests that the population eigenvectors may be closer to the sample eigenvectors in Models 5 and 6.
-To explore this possibility, we recorded the Frobenius norms of the differences between the matrices of sample and population eigenvectors in each of our simulation settings.
-The results in Figure \ref{fig:eigenvectors} seem to support our hypothesis that our class of separable estimators is best-suited for estimating covariance matrices whose sample and population eigenvalues differ greatly.
+We conjecture that this may be because the population eigenvectors may be closer to the sample eigenvectors in Models 5 and 6, while our class of separable estimators is best-suited for estimating covariance matrices whose sample and population eigenvalues differ greatly.
+We leave this for future work.
 
-\begin{figure}
-\begin{center}
-\centerline{  \includegraphics[width=0.95\textwidth]{img/eigenvectors.png}}
-\end{center}
-\caption{Estimation errors of sample eigenvectors over 200 replications. Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.}
-\label{fig:eigenvectors}
-\end{figure}
+% \begin{figure}
+% \begin{center}
+% \centerline{  \includegraphics[width=0.95\textwidth]{img/eigenvectors.png}}
+% \end{center}
+% \caption{Estimation errors of sample eigenvectors over 200 replications. Sparse: Model 1; Hypercorrelated: Model 2; Dense-0.7: Model 3; Dense-0.9: Model 4; Orthogonal: Model 5; Spiked: Model 6.}
+% \label{fig:eigenvectors}
+% \end{figure}
 
@@
 Data were collected from five mice at each time point, but two of the 30-minute hypothalamus samples were sequenced using a different library preparation than the other samples and so were left out of this analysis.
-The data are available from the NCBI Gene Expression Omnibus (GEO; \url{http://www.ncbi.nlm.nih.gov/geo/}) under accession number GSE80345.
+The data are available from the NCBI Gene Expression Omnibus under accession number GSE80345.
 
@@
 
+\section*{Supporting Information}
+Web Appendices, Tables, and Figures referenced in Sections \ref{introduction} and \ref{numerical results} are available with this paper at the Biometrics website on Wiley Online Library.
+The procedures are implemented in the R package \verb|cole|, available on GitHub at \url{http://github.com/sdzhao/cole}.
+
+\section*{Data Availability Statement}
+The data are available from the NCBI Gene Expression Omnibus at \url{https://www.ncbi.nlm.nih.gov/geo/} under accession number GSE80345.
+
 \bibliographystyle{biom} 
 \bibliography{biblio}
-
-\section*{Supporting Information}
-Web Appendices, Tables, and Figures are available with this paper at the Biometrics website on Wiley Online Library.
-
-\appendix
-\section{}
-
-\subsection{Proof of Proposition \ref{prop:Rhat}}
-Using the fact that when $\bb{E}\bs{X}=\bs{0}$, $\bb{E}s_{jk} = \sigma_{jk}$ for all $j,k=1,\ldots,p$.
-Therefore for the class of linear decision rules \eqref{linear class}, the scaled Frobenius risk \eqref{frobenius risk} equals
-\begin{align*}
-R(\bs{\Sigma}, \bs{\delta}) &=\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} \{ (\beta_{S}s_{jk}+\beta_Iu_{jk}-\sigma_{jk})^2 \} \\
-& =\frac{1}{p^2}\sum_{j,k=1}^p \bb{E} [s_{jk}-\sigma_{jk} - \{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}]^2\\
-& =\frac{1}{p^2}\sum_{j,k=1}^p [\bb{E}\{(s_{jk}-\sigma_{jk})^2\}+ \bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2-2(1-\beta_{S})\bb{E}\{s_{jk}(s_{jk}-\sigma_{jk})\}]\\
-& =\frac{1}{p^2}\sum_{j,k=1}^p [(2\beta_{S}-1)\text{Var}(s_{jk})+\bb{E}\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2] \\
-& =\frac{1}{p^2}\sum_{j,k=1}^p\bb{E}[(2\beta_S-1)\frac{n}{n-1}\hat{\Delta}_{jk}^2+\{(1-\beta_{S})s_{jk}-\beta_Iu_{jk}\}^2],
-\end{align*}
-with $\hat{\Delta}_{jk}$ defined in Proposition \ref{prop:linear}. Therefore
-\begin{align*}
-  \bb{E} \hat{R}_L(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta})
-  =\,&
-       \frac{-1}{n - 1} (2\beta_S-1) \frac{1}{p^2} \sum_{j,k=1}^{p}\bb{E}\hat{\Delta}_{jk}^2
-       =
-       \frac{-1}{n} (2\beta_S-1) \frac{1}{p^2} \sum_{j,k=1}^{p} \text{Var}(s_{jk}) \\
-  =\,&
-       \frac{-1}{n} (2\beta_S-1) \frac{1}{p^2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2
-  \rightarrow 0,
-\end{align*}
-where the last result follows because by assumption, $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$.
-
-\subsection{Proof of Proposition \ref{prop:linear}}
-
-% The estimator in \citet{ledoit2004well} is
-% \begin{equation} 
-% \label{eq:lw}
-% \bs{\Sigma}_{LW} = (\frac{d_n^2-b_n^2}{d_n^2})_+\bs{S}+ \min\{1,\frac{b_n^2}{d_n^2}\}\hat{\mu}\bs{I}
-% \end{equation}
-% where $d_n^2=\frac{1}{p}\sum_{j,k=1}^p s_{jk}^2-\frac{1}{p^2}(\sum_{j=1}^p s_{jj})^2$, $b_n^2=\frac{1}{pn^2}\sum_{j,k=1}^p\sum_{i=1}^n (X_{ij}X_{ik}-s_{jk})^2$, $\hat{\mu}=\frac{1}{p}\sum_{j=1}^p s_{jj}$.
-
-  We first rewrite the risk estimate $\hat{R}_L(\beta_S, \beta_I)$.
-Define $\bs{M}=(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2, 0)^\top$, $\bs{\beta} = (\beta_S, \beta_I)^\top$, and the vectorized covariance matrices $\bs{v_S}=(s_{11},\ldots, s_{pp})^\top $, $\bs{v_I}=(u_{11},\ldots, u_{pp})^\top $, and $\bs{v_{\Sigma}} = (\sigma_{11},\ldots,\sigma_{pp})^\top$.
-Then the risk estimate can be re-written as
-  \[
- p^2 \hat{R}_L(\beta_S,\beta_I) = \bs{\beta}^\top (\bs{Z}^\top \bs{Z})\bs{\beta} - 2(\bs{Z}^\top\bs{v_S}-\bs{M})^\top \bs{\beta} - \bs{1}^\top \bs{M}+\bs{v}_S^\top\bs{v}_S,
-  \]
-  where $\bs{Z} =(\bs{v_S}, \bs{v_I})$. Therefore
-  \begin{align*}
-    \hat{\bs{\beta}} &= \arg\min_{ \bs{\beta} }\hat{R}_L(\beta_S,\beta_I) = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z}^\top\bs{v_S}-\bs{M}),\\
-    &
-    \hat{\bs{v}}_{\bs{\Sigma}} = \bs{Z\hat{\beta}} = \bs{v_S} - \bs{Z}(\bs{Z}^\top \bs{Z})^{-1}\bs{M}.
-  \end{align*}
-
-  Since
-$$\bs{Z}^\top \bs{Z} = \begin{pmatrix}
-s_{11} &\ldots &s_{pp} \\
-u_{11} &\ldots &u_{pp} 
-\end{pmatrix}\begin{pmatrix}
-s_{11}  &u_{11} \\
-\ldots&\ldots\\
-s_{pp} &u_{pp} 
-\end{pmatrix}=\begin{pmatrix}
-\sum_{j,k=1}^ps_{jk}^2 & \sum_{j=1}^ps_{jj}\\
-\sum_{j=1}^ps_{jj} & p
-\end{pmatrix}$$
-and
-$\text{det}( \bs{Z}^\top \bs{Z} ) = p\sum_{j,k=1}^ps_{jk}^2 - (\sum_{j=1}^p s_{jj})^2 = p^2d_n^2$, it follows that
-$$(\bs{Z}^\top \bs{Z})^{-1} = \frac{1}{p^2d_n^2}\begin{pmatrix}
-p & -\sum_{j=1}^p s_{jj}\\
--\sum_{j=1}^p s_{jj} & \sum_{j,k=1}^ps_{jk}^2
-\end{pmatrix},$$
-and in addition
-$$\bs{Z}^\top \bs{v_S}= \begin{pmatrix}
-\sum_{j,k=1}^ps_{jk}^2\\
-\sum_{j=1}^p s_{jj}
-\end{pmatrix},
-\quad 
-\bs{Z^\top \bs{v_S}}-\bs{M} =  \begin{pmatrix}
-&\sum_{j,k=1}^ps_{jk}^2 - \hat{\Delta}_{jk}^2\\
-&\sum_{j=1}^p s_{jj}
-\end{pmatrix}.$$
-
-Therefore
-\begin{align*}
-\hat{\bs{\beta}}& = (\bs{Z}^\top \bs{Z})^{-1}(\bs{Z^\top \bs{v_S}}-\bs{M} ) \\
-&= \frac{1}{p^2d_n^2}\begin{pmatrix}
-p & -\sum_{j=1}^p s_{jj}\\
--\sum_{j=1}^p s_{jj} & \sum_{j,k=1}^ps_{jk}^2
-\end{pmatrix}\begin{pmatrix}
-\sum_{j,k=1}^p(s_{jk}^2 - \hat{\Delta}_{jk}^2)\\
-\sum_{j=1}^p s_{jj}
-\end{pmatrix}\\
-&= \frac{1}{p^2d_n^2}\begin{pmatrix}
- p\sum_{j,k=1}^p s_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2\\
- (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)
-\end{pmatrix}.
-\end{align*}
-
-The second component of $\hat{\bs{\beta}}$ equals $\hat{\beta}_I$, and
-\begin{align*}
-\hat{\beta}_I &= \frac{1}{p^2d_n^2} (\sum_{j=1}^ps_{jj}) (\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)\\
-&=\{(\sum_{j=1}^ps_{jj})/p\}   \{(\sum_{j,k=1}^p\hat{\Delta}_{jk}^2)/p \} / d_n^2
-=\hat{\mu}\frac{b_n^2}{d_n^2},
-\end{align*}
-so $\min(\hat{\mu}, \hat{\beta}_I) = \hat{\mu} b_n^2 / d_n^2$, which is the coefficient of the second term in the Ledoit and Wolf estimator \eqref{eq:lw}. Furthermore,
-\begin{align*}
-\hat{\beta}_I /\hat{\mu} + \hat{\beta}_S &= \frac{1}{p^2d_n^2}\sum_{j,k=1}^p \{ps_{jk}^2 -p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2-(\sum_{j=1}^ps_{jj})^2+p\sum_{j,k=1}^p\hat{\Delta}_{jk}^2\}\\
-&=\frac{1}{p^2d_n^2}\{p\sum_{j,k=1}^p s_{jk}^2 -(\sum_{j=1}^ps_{jj})^2\}
-=
-1,
-\end{align*}
-so $\min(\hat{\beta}_S, 0) = 1 - b_n^2 / d_n^2$, which is the coefficient of the first term in \eqref{eq:lw}.
-
-\subsection{Proof of Proposition \ref{prop:bayes risk}}
-
-For any $\bs{\delta} \in \mathcal{S}$ \eqref{separable}, the Frobenius risk \eqref{frobenius risk} can be written as
-\begin{align*}
-  R(\bs{\Sigma}, \bs{\delta})
-  =\,&
-       \frac{2}{p^2} \sum_{1\leq k<j\leq p}^p
-       \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - \sigma_{jk} \}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid \sigma_j, \sigma_k, r_{jk}) d\bs{X}d\bs{X}^{\prime} \,+\\
-  &
-       \frac{1}{p^2}
-       \sum_{j=1}^p
-       \int \{t_d(\bs{X}) - \sigma_{jj} \}^2 f_{1n}(\bs{X} \mid s_{jj}) d\bs{X}\\
-  =\,&
-       \frac{p-1}{p}\int \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - a b \gamma \}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) d\bs{X}d\bs{X}^{\prime} \, +\\
-     &
-       \frac{1}{p}\int \int \{t_d(\bs{X}) - a^2\}^2 f_{1n}(\bs{X} \mid a) dG_{d}(a) d\bs{X}.
-\end{align*}
-Therefore for any $t_{od}$,
-\begin{align*}
-  &
-    \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - a b \gamma \}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \\
-  =\,&
-       \int \{t_{od}(\bs{X},\bs{X}^{\prime}) - t_{od}^\star(\bs{X}, \bs{X}^\prime)\}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \,+\\
-  &
-    2 \{t_{od}(\bs{X},\bs{X}^{\prime}) - t_{od}^\star(\bs{X}, \bs{X}^\prime)\} \int 
-  \{t_{od}^\star(\bs{X}, \bs{X}^\prime) - a b \gamma \}
-  f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \,+\\
-  &
-    \int \{t_{od}^\star(\bs{X}, \bs{X}^\prime) - a b \gamma\}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma) \\
-  \geq\,&
-          \int \{t_{od}^\star(\bs{X}, \bs{X}^\prime) - a b \gamma\}^2 f_{2n}(\bs{X}, \bs{X}^{\prime} \mid a, b, \gamma) dG_{od}(a, b, \gamma),
-\end{align*}
-because the middle cross term is equal to zero by the definition of $t^\star_{od}$. Similarly, it can be shown that for any $t_d$,
-\begin{align*}
-  \int \{t_d(\bs{X}) - a^2\}^2 f_{1n}(\bs{X} \mid a) dG_{d}(a)
-  \geq
-  \int \{t_d^\star(\bs{X}) - a^2\}^2 f_{1n}(\bs{X} \mid a) dG_{d}(a).
-\end{align*}
-This implies that $R(\bs{\Sigma}, \bs{\delta}) \geq R(\bs{\Sigma}, \bs{\delta}^\star)$ for any $\bs{\delta} \in \mathcal{S}$.
 

### 2022-03-08 response_reviews.tex
small changes for manuscript and response of reviews
edited to satisfy reviewer 2's comments

--- 2021-11-02 response_reviews.tex
@@
 \usepackage[margin=1in]{geometry}
+\usepackage[round]{natbib}
 \usepackage{amsmath}
@@
 
-\section{Response to the Associate Editor}
+\section{Response to Reviewer 2}
+
+Thank you for your very careful second reading and the excellent comments. We respond to each point below.
 
 \begin{enumerate}
-\item \emph{As a journal, Biometrics presents papers that focus on the development of new methods and results of use in biosciences.
-As of now, the application seems more an afterthought, the use of a convenient (easily available) dataset to fit the scope of the journal.
-I believe that the authors should motivate their approach from the gene expression data or other bio-related applications starting from the Introduction.}
-\item \emph{It is not really intuitively clear why the vectorization works, despite the approach not taking into consideration the structure of the covariance matrix.
-Also, the manuscript proposes a series of ad-hoc expedients to implement the method.
-See the clustering-based exemplar algorithm and replacing non-positive eigenvalues with a chosen positive value $c$ in the positive-definite correction.
-These expedients do not seem to be completely justified by the proposed separable compound-decision theory framework, and their effects on the estimation are unclear.
-A referee also commented on the impossibility to follow closely the implementation of the EM algorithm.
-In short, while I appreciate that the theoretical justification for the asymptotic efficiency of the proposed vectorization is not clear, the current implementation reads more like a clever algorithm than a well-founded statistical method.
-The authors should justify all the steps of the algorithm and provide an assessment of their contribution to its overall performance.}
+\item \emph{Population and sample eigenvectors: A key thesis of the revised manuscript that is emphasized multiple times (even in the abstract) is that the new method outperforms optimal rotationally invariant methods in situations where sample eigenvectors are far from population eigenvectors.
+I think this claim is not suffciiently supported \ldots I think either more compelling evidence should be provided, or the claim can be downemphasized and perhaps mentioned once in the manuscript as a possible conjecture.}
 
-\item \emph{I agree with one of the referees that the title should be re-thought, since compound decision theory or compound decisions appear quite relevant to the approach, more than the empirical Bayes aspect of the paper.}\\
-We agree that compound decision theory should be more emphasized than empirical Bayes in our paper.
-Following your suggestions, we have changed the title from "A Empirical Bayes Approach to Covariance Matrix Estimation" to "A Compound Decision Approach to Covariance Matrix Estimation".
+  This makes sense. We have downplayed this point and merely mention it as a conjecture; \textbf{see the last paragraph of Section 4}.
 
-\item \emph{on page 4, it appears to me that the choice of the risk function $\hat{R}$ does not correspond to the Frobenius risk.
-The rationale for choosing this risk function should be provided.
-It somehow appears the choice is dictated by Ledoit et al (2004) and Proposition 1, and not by independent first principles.
-If so, it should be explicitly stated before.}\\
-We squared and scaled the Frobenius norm $\|A\|_F = \sqrt{\sum_{j,k=1}^pa_{jk}^2}$ to $\frac{1}{p^2}\sum_{j,k=1}^p a_{jk}^2$ in order to apply fundamental theorem of compound decisions, for which the risk function should be scaled by the number of parameters because the empirical distribution has mass probability of $\frac{1}{d}$ where $d$ is the number of parameters.\\
-The transformation of Frobenius norm will not change the value of $(\widehat{\beta}_S,\widehat{\beta}_I)$ since it is determined by minimizing $RE(\beta_S,\beta_I)$.
-\item \emph{Figures 1 and 2 do not allow to appreciate uncertainty. Perhaps a table would be better.}\\
-We add the tables of the medians and uncertainty of Frobenius errors in supporting information.
-The uncertainty of Frobenius error is measured by the difference of 75\% quantile and 25\% quantile in 200 replicates.
-We use this range to measure the uncertainty because the errors of most methods are not following normal distribution.
-\item \emph{The manuscript requires careful proofreading and improvements. in the structure of the presentation and the visualization of the results. Non-exhaustive examples are reported below:}
+\item \emph{Gene networks in Figures 4 and 5: I found these networks diffciult to interpret.
+It would greatly help if the presentation of these figures could be improved.
+Perhaps one could present less methods, but superimpose resulting edges onto a joint set of nodes?}
 
-\emph{- p.1 “To overcome these issues, various methods have been developed to estimate high-dimensional covariance matrix” $\to$ matrices}
+  Thanks for the great suggestion, we have adopted your approach; \textbf{see Figures 3 and 4 and a discussion of the figures at the end of Section 5}.
 
-It has been fixed.
+\item \emph{counts per million mapped reads: My understanding is that this method of normalizing counts is not recommended for RNA-Seq experiments.
+For example, DESeq2 uses another method of normalization.
+Why was this method chosen?
+Do the results change with other normalization methods?
+}
 
-\emph{- p. 1 “about the structure of population covariance matrix” $\to$ of the population covariance matrix}
+  There are indeed a number of normalization methods, but most of the work in this area has been for identifying differentially expressed genes.
+It is not always clear what normalization to use for more complex analyses, such as the coexpression network study in this paper.
+In particular, the most recent version of the edgeR user's guide states in Section 2.16 that
 
-It has been fixed.
+  ``Inputing RNA-seq counts to clustering or heatmap routines designed for microarray data is not straight-forward, and the best way to do this is still a matter of research.
+To draw a heatmap of individual RNA-seq samples, we suggest using moderated log-counts-per-million.''
 
-\emph{- p. 2 “Other common structured methods assume that the covariance matrix is banding” $\to$ banded?}
+  We have followed this CPM approach in our analyses. CPM normalization for coexpression has also been shown to have good performance \citep{johnson2022robust}.
 
-It has been edited to 'banded'.
+\item \emph{Could the authors provide a bit more detail on splitting into training and testing set?
+Is time ignored when splitting the samples?
+Also is the sample covariance (that first centers observations; as in the simulation) used for both test and training samples?}
 
-\emph{- p. 2 “yet can still outperform the sample covariance matrix” What does this mean?}
+  We do ignore time, for the sake of sample size.
+This was also done in the original Saul et al.
+(2017) paper.
+We also state that we use the sample covariance matrix, which first centers the observations; \textbf{see the second-to-last paragraph of Section 5}.
 
-It means that these unstructured model have been shown to have lower estimation risk comparing to sample covariance matrix.
-For example, in \cite{ledoit2004}, the numerical result shows that their linear shrinkage method has much lower risk than sample covariance matrix.
-To make it clear, we have changed to "yet still have less estimation error comparing to the sample covariance matrix".
+\item \emph{OracMSG: (method defined on Page 16): Why do support points need to be generated here?
+Since all the ($\sigma_j$, $\sigma_k$, $r_{jk}$) are known for the oracle method, I would think that the oracle rule can be computed directly without any need for optimization.}
 
-\emph{- p. 3 “is squared error loss”$\to$ is the squared error loss}
+  We performed clustering even though the true parameters are all known in order to reduce the number of support points to ease the computational burden.
+When $p$ is large the number of true support points could be very large, which dramatically increased the computation time.
+We have now mentioned this; \textbf{see Section 4.2}.
 
-It has been fixed.
+\item \emph{Some notation on indexing the population/sample covariance matrix should be made more explicit in the manuscript.
+In particular, it appears that throughout the text there is the convention on single/double-indexing that $\sigma^2_j = \sigma_{jj}$ and $s^2_j = s_{jj}$ that confused me in light of e.g.
+(2).
+(Similarly for $r_{jk}$.)}
 
-\emph{- p.
-4 “Now consider the problem of estimating the vectorized $\boldsymbol{\Sigma}$ under risk 1” $\to$ risk (1) Equations are typically referenced between brackets.
-The same issue appears in many other points where equations are referenced throughout the manuscript.}
+  We have now clarified that $\sigma_{jj} = \sigma_j^2$; \textbf{see the first paragraph of Section 2.1}.
 
-We used to use \textbackslash ref to reference equation, we have changed it to \textbackslash eqref so that they are referenced between brackets.
+\item \emph{Page 17, ``The corrected version consistently outperformed the uncorrected version.'': I think this should hold provably (not only just empirically) by a convex analysis projection argument.}
 
-\emph{- p. 8 “A major advantage of nonparametric estiation of the prior”$\to$ estimation}
+  Thanks for this insight!
+However, we do not exactly project our matrix onto the space of positive semi-definite matrices because we want positive-definiteness.
+Our minimum eigenvalue is $c_\alpha$, as we describe in (14), and we haven't yet thought through how this would affect the argument.
 
-It has been fixed.
+\item \emph{Page 3, ``The article is organized as follows.'', something is off with the numbering of Sections in this paragraph. }
 
-\emph{- page 8, “Figure 1 shows that different $K$ have similar estimation accuracy compared to the exemplar algorithm, while Table 1 shows that they can be significantly faster” However, at this point Figure 1 refers to models that have not yet been presented.
-Also, the reference to Table 1 anticipates point made in Section 3.2}
+  Thank you for pointing this out, we have fixed the numbering.
 
-We have moved the setting of the experiment to the start of Section 3 Numerical Results.
+\item \emph{Page 4, ``incentived'' typo }
 
-\emph{- p. 16 “Finally, we have so far assumed that our data multivariate normal” $\to$ are normal}
+  Fixed, thanks.
 
-It has been fixed.
+\item \emph{Page 4, ``about index i ... different index j'': Maybe phrase this in terms of different parameters rather than indices?}
 
-\emph{- In Equations (1) and (2) the arguments of the risk functions are $\theta$ and $\delta$. However, on page 5 the arguments of $\hat{R}$ are $\beta_{S}$, and $\beta_{I}$.}
+  We have followed your suggestion; \textbf{see the last paragraph of Section 2.1}.
 
-We have changed the notation $\hat{R}$ to $\text{RE}$ which means risk estimator and this differentiates it from $R$. 
+\item \emph{Page 5, ``that minimizes the compound risk.” $\rightarrow$ “that minimizes an estimate of the compound risk''?}
+
+  Fixed, thanks.
+
+\item \emph{Page 5, after equation 3: Maybe explicitly write that $\beta_S$, $\beta_I$ are scalars (i.e., $\beta_S, \beta_I \in \mathbb{R}$)?
+Has $\bs{I}$ been explicitly defined as the identity matrix?
+}
+
+  We have now written that $\beta_S$ and $\beta_I$ are scalars are have identified $\bs{I}$ as the $p \times p$ identity matrix; \textbf{see after equation (3)}.
+
+\item \emph{Page 9: ``In this case, the optimal separable rule for estimating $\sigma_{jk}$ borrows from all of the $\sigma_j$ and $r_{jk}$'' $\rightarrow$ Perhaps use different indices when referring to the first $\sigma_{jk}$ and when referring to all parameters?
+}
+
+  We have now reworded this as ``In this case, the optimal separable rule for estimating the $jk$th entry of $\bs{\Sigma}$ borrows from all of the $\sigma_j$ and $r_{jk}$''; \textbf{see after (11)}.
+
+\item \emph{Page 10: ``We propose to calculate estimate'' $\rightarrow$ ``we propose to estimate''?}
+
+  Fixed, thanks.
+
+\item \emph{Page 18: ``The data are from the NCBI'' $\rightarrow$ ``The data are available from the NCBI''?}
+
+  Fixed, thanks.
+
+\item \emph{Page 21: ``These imposes significant'' $\rightarrow$ ``impose'' }
+
+  Fixed, thanks.
+
 \end{enumerate}
 
-\section{Comments to Referee1}
-Thank you so much for the detailed and inspiring suggestions on our paper.
-We learnt a lot through them.
-We agree with your comments and address them in the updated draft version.
-Please find the following responses for more details.
+\bibliographystyle{abbrvnat}
+\bibliography{biblio}
 
-\subsection{Comments to section "CorShrink"}
-After carefully reviewed \cite{dey2018}, we think it is very related to our paper and it is valuable to make the comparison.
-
-As mentioned in Section 1, our method is quite similar to \cite{dey2018}.
-Both of us used empirical Bayes estimator and nonparametric estimation of prior distribution.
-The difference is that our method aim to estimate covariance matrix which requires to also shrink standard deviations.
-
-We also compared our two methods by numerical experiments which are shown in supporting information.
-We designed two simulations to compare the performance on correlation matrix estimation and covariance matrix estimation.
-In the first simulation, we compare covariance matrix estimation.
-For corShrink, $\widehat{\bs{\Sigma}} = \text{diag}(\widehat{SD}) \widehat{\bs{R}} \text{diag}(\widehat{SD})$, where $\widehat{SD}$ are sample standard deviations.
-In the second simulation, $\bs{\widehat{R}} = \text{diag}(1/\widehat{SD})\widehat{\bs{\Sigma}}\text{diag}(1/\widehat{SD})$, where $\widehat{SD}$ are standard deviations derived from $\widehat{\bs{\Sigma}}$.
-For each setting, we take $n$ as 20 and 100.
-
-\subsection{Comments to section "Separable rules, symmetry and the diagonal"}
-\begin{equation}\label{R1}
-\widehat{\Sigma} = (\widehat{\sigma}_{jk})_{1\leq j,k\leq p},\quad \widehat{\sigma}_{jk} = t(X_{\cdot,j}, X_{\cdot,k}), \quad t:\bb{R}^n\times \bb{R}^n\to \bb{R}
-\end{equation}
-\begin{enumerate}
-\item \emph{While reading the paper I was a bit confused by the fact that (R1) treats diagonal entries and off-diagonal in the same way.
-It would feel more natural to me to define a separable estimator as in (R1) for $j\neq k$, and as follows on the diagonal:}
-$$\hat{\sigma}_{jj} = \tilde{t}(X_{\cdot j}),\quad \tilde{t}:\bb{R}^n \to \bb{R}$$
-We have changed the rule to 
-\begin{equation}
-  \label{separable}
-  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
-   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
-\end{equation}
-\emph{In the Multivariate Gaussian case this is not an issue, since $X_{\cdot j}\neq X_{\cdot k}$ for $j\neq k$ with probability 1.
-But the proposed approach would also apply with minor modifications to e.g., discrete measurements (with known likelihood) in which case $X_{\cdot j}\neq X_{\cdot k}$ could happen with non-zero probability.} \\
-We agree that it is not very precise to use $I(\bs{X}_{\cdot j}=\bs{X}_{\cdot k})$ to differentiate on-diagonal entries and off-diagonal entries.
-So we changed separable rule to the above class of rules.
-
-\emph{Furthermore, while this issue is ignored for most of the paper, diagonal and off-diagonal entries are treated in a different way at the end of Section 2.3 (in the context of the clustering based algorithm).
-I believe it would be valuable to make the above distinction more explicit!}
-\item \emph{A related question I had concerns the symmetry of the estimator. Does the estimator satisfy the symmetry property $\hat{\sigma}_{jk} =\hat{\sigma}_{kj}$? }
-
-Yes, we only calculate estimates of lower triangular matrix of $\Sigma$, i.e., $\hat{\sigma}_{jk}$ where $j\geq k$.
-Then the upper triangular matrix estimation is calculated by the symmetry of covariance matrix $\sigma_{kj} = \sigma_{jk}$.
-Correspondingly, we should fix the Bayesian model in Section2.2 from
-\begin{equation}
-  \label{bayesian}
-  \bs{A} \mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
-  \quad
-  \bs{\eta} \sim G_p(a,b, \gamma) =  \frac{1}{p^2} \sum_{j,k=1}^p I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
-\end{equation}
-to 
-\begin{equation}
-  \label{bayesian nondigaonal}
- (\bs{X}_{\cdot j},\bs{X}_{\cdot k})\mid \bs{\eta} \sim f(\cdot \mid \bs{\eta}),
-  \quad
-  \bs{\eta} \sim G_{nd}(a,b, \gamma) =  \frac{2}{p(p-1)} \sum_{1\leq k< j \leq p} I(\sigma_j \leq a, \sigma_k \leq b, r_{jk} \leq \gamma).
-\end{equation}
-and
-\begin{equation}
-  \label{bayesian nondigaonal}
-  \bs{X}_{\cdot j} \mid a \sim \widetilde{f}(\cdot \mid a),
-  \quad
-  a \sim G_{d}(a) =  \frac{1}{p} \sum_{j=1}^p I(\sigma_j \leq a).
-\end{equation}
-, as well as the proposed estimator from
-\begin{equation}
-  \label{Gp hat}
-  \widehat{G}_p = \argmax_{G \in \mathcal{G}} \prod_{j,k = 1}^p \int f(\bs{A}_{jk} \mid \bs{\eta}) dG(\bs{\eta}),
-\end{equation}
-to
-\begin{equation}
-  \label{Gp hat revision}
-  \widehat{G}_{nd} = \argmax_{G_{nd} \in \mathcal{G}} \prod_{1\leq k< j \leq p} \int f(\bs{X}_{\cdot j},\bs{X}_{\cdot k} \mid \bs{\eta}) dG_{nd}(\bs{\eta})\prod_{j=1}^p \int \widetilde{f}(\bs{X}_{\cdot j} \mid a) dG_{d}(a),
-\end{equation}
-\emph{Similarly, the reader may wonder whether the estimation procedure enforces symmetry in the prior $G$ in (6) with respect to $\sigma_j,\sigma_k$.}
-
-In correspondence, we modified our sample exemplar algorithm, we first clusters sample points $(s_j,s_k, r_{jk})$ of lower triangular part of sample covariance matrix and get the $K$ centroids $(a,b,\gamma)$, then find their symmetric points $(b,a,\gamma)$.
-We also updated EM algorithm to guarantee each $(a,b,\gamma)$ and $(b,a,\gamma)$ has the same weight.
-\item \emph{The paper emphasizes the Multivariate Gaussian problem.
-In that case, the sample covariance is sufficient, and so, it would seem more natural to me to define a separable rule, as one that satisfies,
-$$\hat{\sigma}_{jk} = \bar{t}(X_{\cdot j}^TX_{\cdot k}),\quad \bar{t}: \bb{R}\to \bb{R}_{+}$$
-instead of (R1). I am wondering why the authors chose to use (R1) instead.}
-
-$X_{\cdot j}^TX_{\cdot k}$ is sufficient statistics of the single entry $\sigma_{jk}$.
-In fact, for $(\bs{X}_{\cdot,j}, \bs{X}_{\cdot,k})$, the sufficient statistics for $\eta$ is $\bs{S}_{jk} = \begin{pmatrix} s_{jj}&s_{jk} \\  s_{kj}&s_{kk} \end{pmatrix}$.
-We also tried density based on single entries $s_{jk}$ before but the performance is not as good as using $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$.
-\item \emph{If I understand correctly, one of the benefits of estimator (3) is that it works without specific assumptions on the noise distribution.
-Instead the approach here appears to more heavily rely on a known likelihood (and specifically a multivariate Gaussian likelihood).
-It would be great if some simulations could be conducted under misspecification, wherein the data is not multivariate Gaussian, but the approach is applied assuming multivariate Gaussian data.}
-
-This is a good point.
-To measure the performance of our method on nonGaussian distribution, we did some simulations on negative binomial distribution.
-We first generate independent distributed $Y$ and then let $X = L Y$, where L is the Cholesky decomposition of $\Sigma$.
-The simulation results are presented in supporting information.
-\end{enumerate}
-
-\subsection{Comments to section "Compound decision theory or empirical Bayes?"}
-\begin{enumerate}
-\item \emph{The result following equation (6) could be presented separately as a Theorem (or Proposition).
-It should be highlighted that part of the statement is that the estimator in (R1) that optimizes (R2) depends on all the $\sigma_{j}$ and $\rho_{jk}$.}
-
-We rewrote the equation after (6) as a separate proposition to show the equivalence between the Frobenius risk and Bayes risk.
-\item \emph{The term “compound decision theory” or “compound decisions” etc. deserves to be in the title!}
-
-Yes, we agree that "compound decision" is more deserved to be highlighted in the paper so we replaced "nonparametric empirical Bayes" with "compound decision" in the title.
-\item \emph{On the other hand, I would downemphasize the empirical Bayes aspect of the paper (and my preference would be to even remove it from the title).
-To me, an empirical Bayes analysis (especially since here the goal is to perform well, in a frequentist sense, in terms of (R2)) has the following flavor: “Let us mimic what a Bayesian would do, when we do not know the true prior, and in a way that we can get frequentist guarantees as well”.
-However, I do not think a Bayesian would start approaching this problem by first positing (6) and then using (8)2.
-Instead a more natural Bayesian approach would be to set a prior for the full covariance matrix $\Sigma$, e.g., as in Berger et al.
-[2020] and references therein.
-I would restrict the discussion of empirical Bayes to just saying that model (6) appears formally when optimizing over separable rules (R1).
-To approximate the unknown optimal rule, one follows an empirical Bayesian “Generalized Maximum Likelihood” approach with respect to the pairwise composite likelihood.3}
-\end{enumerate}
-
-\subsection{comments to section "other remarks"}
-\begin{enumerate}
-\item \emph{It would be of great service to the community if the authors could provide code implementing the method and reproducing the results!}
-
-We put our code of methods and simulation in supporting information.
-\item \emph{Related to the above point, based on the description in the manuscript it would not be possible for me to fully reproduce the results.
-For example, does the actual implementation use the EM algorithm or an interior point convex solver?
-If the EM algorithm is used, how many steps are taken or what is the termination criterion?
-And why is the EM algorithm used (since for these empirical Bayes problem it typically converges extremely slowly compared to interior point solvers, as carefully demonstrated by Koenker and Mizera [2014])?
-On the other hand, if an interior point solver is used, which one?}
-
-In our implementation, we used EM algorithm to estimate the prior distribution.
-We added the related details to Section 3 Numerical Results.
-The maximum number of iterations is set as 200.
-Iteration early stops when the criterion $\frac{L(\bs{w}^{(k+1)}) - L(\bs{w}^{(k)})}{L(\bs{w}^{(k)})}\leq 10^{-4}$ is met.
-We also tried interior point, but it always failed to find the optimal solution.
-The possible reason is the density $f(\cdot|a,b,\gamma)$ has very different scales, from $10^{-22}$ to $2$ and this makes interior point method difficult to solve.
-\item \emph{First paragraph of page 3, “they modify only the sample eigenvalues and not the sample eigenvectors”: The situation here is a bit more nuanced, since e.g., sometimes using the sample eigenvectors is provably the optimal thing to do (e.g., if one constrains themselves to rotationally invariant estimators as in Bun et al.
-[2016] and references therein)}
-
-The sentence "However, they modify only the sample eigenvalues and not the sample eigenvectors." is not very appropriate so we changed it to "However, as rotation-equivariant estimators (Stein, 1975, 1986), shrinkage estimators modify only the sample eigenvalues and keep the sample eigenvectors.".
-\item \emph{Page 5, “it is straightforward to show that”: The “straightforward” could be replaced by a short proof in the appendix.}
-
-We added the proof in the appendix.
-
-\emph{Also could it be that the risk estimate is not exactly unbiased (because of dividing by $n$ and not $n-1$ in the definition of $\hat{\Delta}_{jk}^2$)?}
-You're right.
-We realized that the previous defined $\hat{\Delta}_{jk}^2$ is not unbiased estimator of Frobenius risk of $\bs{\delta}$ (1).
-So we changed the text to describe it as an "approximately unbiased risk estimator" since the factor $\frac{n-1}{n}$ is closed to 1.
-It is also shown in appendix B.
-
-\emph{Where is the sample covariance $\bs{S}$ defined?}
-
-In the updated manuscript, $\bs{S}$ is defined as $\bs{S} = \frac{1}{n}\bs{X}^\top \bs{X}$ as it is first mentioned in page 4, section 2.1.
-
-\item \emph{Section 2.5, “Projections in terms of other matrix norms are also possible”. Is this actually being done in the implementation?}
-
-No. We did not implement it in terms of other norms.
-
-\emph{Else perhaps remove this sentence.}
-
-It has been removed.
-\end{enumerate}
-\section{Comments to Referee2}
-\begin{enumerate}
-\item \emph{In the numerical experiments and the real data analysis, what was the size of the grid that was used? I may have missed it but I did not see it mentioned in the caption for figure 2.}\\
-In section 3.3, methods compared and section 3.4, data analysis, we used clustering-based exemplar algorithm, where we find $K$ clusters of all off-diagonal sample points.
-$K$ is set to be $p$ to balance accuracy and speed according to result of section 3.2.
-\item \emph{In the numerical experiments, the dimension p goes as far as 200. I believe it will be useful to consider a setting where $n = 100$ and $p = 1000$.} \\
-We did the simulation for $n=100$ and $p=1000$ using the different covariance matrix models in Section 3.1. Please see the supporting information for the simulation results.  
-\item \emph{Continuing on the numerical experiments, models 1 to 5 impose various structures on $\bs{\Sigma}$.
-It will be interesting to see the performance of the proposed method if a spiked covariance structure is imposed on $\bs{\Sigma}$.} \\
-We added the spiked model as the sixth matrix models in Section 3.1.
-Spiked covariance matrix is generated by $\Sigma = \bs{U}\text{diag}(\bs{l})\bs{U}^\top$, where $\bs{U}$ is a randomly generated orthogonal unit matrix, $\bs{l} = \{4,3,2,1,\ldots,1\}$.
-\item \emph{Typo on page 6: "The density of $f(\cdot \mid \bs{\eta}_{jk})$ of ...".}\\
-Thanks for poiting this out. We've fixed this sentence as "$f(\cdot \mid \bs{\eta}_{jk})$, the density of $\bs{A}_{jk}$, depends on $\bs{\eta}_{jk} = (\sigma_j, \sigma_k, r_{jk})^\top$, ...".
-\end{enumerate}
-\begin{thebibliography}{9}
-\bibitem{dey2018}
-Dey, K. K., \& Stephens, M. (2018). CorShrink: Empirical Bayes shrinkage estimation of correlations, with applications. bioRxiv, 368316.
-\bibitem{ledoit2004}
-Ledoit, O., \& Wolf, M. (2004). A well-conditioned estimator for large-dimensional covariance matrices. Journal of multivariate analysis, 88(2), 365-411.
-
-\end{thebibliography}
 \end{document}

### 2022-03-08 msg.tex
edited to satisfy reviewer 2's comments

@@
 The article is organized as follows.
-In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed estimator.
-In Section \ref{numerical results} we illustrate its performance in simulations and a gene coexpression network inference problem.
+In Section \ref{sec:compound} we briefly review compound decision theory, and in Section \ref{sec:method} we introduce our proposed estimator.
+In Section \ref{numerical results} we illustrate its performance in simulations and Section \ref{gene analysis} contains an analysis of the gene coexpression network inference problem.
 Section \ref{sec:discussion} concludes with a discussion, and proofs can be found in the Supporting Information.
@@
 \end{equation}
-where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate.
+where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate and $\sigma_{jj} = \sigma^2_j$ is the variance of the $j$th feature.
 
@@
 One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite.
-Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}.
+Notably, however, this structure is not incentivized by the Frobenius risk \eqref{frobenius risk}.
 As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite.
@@
 The value of this formulation is that it allows us to apply ideas from the compound decision literature to covariance matrix estimation.
-A key property of compound decision problems is that while information about index $i$ seems like it should offer no help for estimating a parameter with a different index $j$, in fact borrowing information across all indices is superior to considering each index alone.
+A key property of compound decision problems is that while information about a parameter $\theta_i$ seems like it should offer no help for estimating any other $\theta_j$, in fact borrowing information across all $\theta_i$ is superior to considering each parameter separately.
 An example of this phenomenon is the James-Stein shrinkage estimator \citep{james1961estimation}% , which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$.
@@
 , and a long line of subsequent work has led to much more sophisticated and accurate procedures \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.
-In Section \ref{sec:method} we apply these ideas to covariance matrix estimation,
+In Section \ref{sec:method} we apply these ideas to covariance matrix estimation.
 
@@
 As a further illustration of the potential of our new formulation, we can show that the linear shrinkage covariance matrix estimator of \citet{ledoit2004well} can be essentially recovered as the solution to a compound decision problem.
-A standard approach in this literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes the compound risk % .
+A standard approach in this literature is to first define a class of decision rules for estimating each component and then identify the member of that class that minimizes an estimate of the compound risk % .
 For example, in the Gaussian sequence problem the Efron-Morris estimator \citep{efron1973stein} can be viewed as the optimal linear decision rule
@@
 \end{equation}
-for some parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of $\bs{I}$.
+for some scalar parameters $\beta_S$ and $\beta_I$, where $s_{jk}$ is the $jk$th entry of $\bs{S}$ and $u_{jk}$ is the $jk$th entry of the $p \times p$ identity matrix $\bs{I}$.
 Ideally we would like to estimate the optimal $\beta_S$ and $\beta_I$ by choosing the values that minimize the Frobenius risk \eqref{frobenius risk}.
@@
 Proposition \ref{prop:bayes risk} illustrates the key property of compound decision problems: that borrowing information across indices can help estimate a single parameter.
-In this case, the optimal separable rule for estimating $\sigma_{jk}$ borrows from all of the $\sigma_j$ and $r_{jk}$.
+In this case, the optimal separable rule for estimating the $jk$th entry of $\bs{\Sigma}$ borrows from all of the $\sigma_j$ and $r_{jk}$.
 This rule also has an interesting Bayesian interpretation, as $t^\star_{od}$ and $t^\star_d$ are formally equivalent to posterior expectations when the $(\sigma_j, \sigma_k, r_{jk})$ are independently and identically distributed according to the prior $G_{od}$.
@@
 We propose to estimate $\bs{\delta}^\star$ using what \citet{efron2014two} calls $g$-modeling, where we proceed as if the $(\sigma_j, \sigma_k, r_{jk})$ were truly random and then estimate the prior $G_{od}$ from the data.
-Specifically, under the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent for different $(j, k)$, we propose to calculate estimate $G_{od}$ and $G_d$ \eqref{eq:G} using nonparametric maximum likelihood \citep{kiefer1956consistency}:
+Specifically, under the working assumption that the $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$ are independent for different $(j, k)$, we propose to estimate $G_{od}$ and $G_d$ \eqref{eq:G} using nonparametric maximum likelihood \citep{kiefer1956consistency}:
 \begin{equation}
@@
 In addition to the above data-driven estimators, we also implemented two oracle estimators, which cannot be implemented in practice as they require the unknown $\bs{\Sigma}$: (a) OracNonlin: the optimal rotation-invariant covariance estimator, derived in \citet{ledoit2019quadratic}, of $\bs{U}^T\text{diag}(\bs{l})\bs{U}$, where $\bs{U}=(\bs{u}_1\ldots\bs{u}_p)$ is the sample eigenvector matrix and $\bs{l} = (d_1, \ldots, d_p)$ is composed of oracle eigenvalues $d_i = \bs{u}_i^T\bs{\Sigma} \bs{u}_i$; and (b) OracMSG: we implemented our proposed estimator \eqref{proposed} except that the support points are generated by clustering the true parameters $(\sigma_j,\sigma_k,r_{jk})$ instead of their sample versions.
+We performed clustering even though the true parameters are all known in order to reduce the number of support points to ease the computational burden.
 
@@
 To measure the accuracy of the estimators, we randomly split the samples into a training and testing set.
+For the sake of sample size, we have ignored the timepoints at which the gene expressions were measured, following \citet{saul2017transcriptional}.
 We estimated the covariance matrix on the training set and measured accuracy using the Frobenius norm of the difference between the estimated matrix and the sample covariance matrix calculated from the test set.
@@
   \begin{center}
-    \includegraphics[width=0.9\textwidth]{img/networks_A.png}
+    \includegraphics[width=0.45\textheight]{img/networks_A_2.png}
   \end{center}
-\caption{Amygdala gene networks recovered by the different covariance matrix estimation methods.}
+\caption{Amygdala gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
 \label{network_A}
@@
   \begin{center}
-    \includegraphics[width=0.9\textwidth]{img/networks_FC.png}
+    \includegraphics[width=0.45\textheight]{img/networks_FC_2.png}
   \end{center}
-\caption{Frontal cortex gene networks recovered by the different covariance matrix estimation methods.}
+\caption{Frontal cortex gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
 \label{network_FC}
@@
 
-In addition to comparing the numerical accuracies, we also investigated whether our estimator gave qualitatively different gene networks compared to the other approaches.
-Figures \ref{network_A} and \ref{network_FC} illustrate the estimated amygdala and frontal cortex covariance matrices in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects.
+In addition to comparing the numerical accuracies, we also compared the gene networks produced by applying the various approaches to all samples.
 To avoid completely connected graphs, we sparsified the matrix estimates by thresholding the smaller entries of each matrix to zero.
 Since the adaptive thresholding method of \citet{cai2011adaptive} naturally produced a sparse estimated matrix, we thresholded the other matrix estimates to match the sparsity level of the \citet{cai2011adaptive} estimate.
-The network representations suggest that the resulting networks look fairly similar except for the ones produced by linear shrinkage and NERCOME.
-That our procedure produces networks that look similar to the unbiased sample covariance matrix networks is comforting.
+Figures \ref{network_A} and \ref{network_FC} illustrate the estimated amygdala and frontal cortex covariance matrix estimates from the best performers from Table \ref{tab:tab1} in network form, where each node represents a gene and each edge represents a non-zero covariance between the genes it connects.
+The figures show that our MSGCor approach gave qualitatively different gene networks.
+This suggests a useful strategy for estimating gene networks in practice: high-confidence edges can be identified as those edges recovered by multiple covariance matrix estimators.
 
@@
 In contrast, our covariance matrix estimator is built from arbitrarily dependent $(\bs{X}_{\cdot j},\bs{X}_{\cdot k})$.
-These imposes significant theoretical difficulties that will require substantial work to address; we leave this for future research.
+These impose significant theoretical difficulties that will require substantial work to address; we leave this for future research.
 

### 2022-03-15 response_reviews.tex
changed response to reviewer title
update QIS, supp

@@
 
-\title{Response to reviews of BIOM2021236M}
+\title{Response to reviews of BIOM2021236M.R1}
 \author{}
@@
 
-  We have now written that $\beta_S$ and $\beta_I$ are scalars are have identified $\bs{I}$ as the $p \times p$ identity matrix; \textbf{see after equation (3)}.
+  We have now written that $\beta_S$ and $\beta_I$ are scalars and have identified $\bs{I}$ as the $p \times p$ identity matrix; \textbf{see after equation (3)}.
 

### 2022-03-15 msg.tex
update QIS, supp

@@
 For each model, we generated $n=100$ samples from a $p$-variate $\mathcal{N}(\bs{0}, \bs{\Sigma})$, where $p = 30, 100,$ or $200$.
-Although $\bb{E}\bs{X}=0$ in our setting, we assume it is unknown and use $\bs{S}=\frac{1}{n-1}(\bs{X}-\bs{\bar{X}})^\top(\bs{X}-\bs{\bar{X}})$.
+Although $\bb{E}\bs{X}=0$ in our setting, we assume it is unknown and use $\bs{S}=\frac{1}{n-1}(\bs{X}-\bs{\overline{X}})^\top(\bs{X}-\bs{\overline{X}})$.
 We generated $200$ replicates and reported median errors under the loss function
@@
   Linear & 2.3 (2.05, 2.55) & 2.35 (2.19, 2.53)\\
-  QIS & 2.61 (2.33, 2.86) & 2.79 (2.64, 2.93)\\
+  QIS & 2.48 (2.07, 2.85) & 2.42 (2.18, 2.70)\\
   NERCOME & 2.37 (2.14, 2.62) & \textbf{2.3 (2.14, 2.53)}\\
@@
 
-\section*{Supporting Information}
-Web Appendices, Tables, and Figures referenced in Sections \ref{introduction} and \ref{numerical results} are available with this paper at the Biometrics website on Wiley Online Library.
-The procedures are implemented in the R package \verb|cole|, available on GitHub at \url{http://github.com/sdzhao/cole}.
-
 \section*{Data Availability Statement}
@@
 \bibliography{biblio}
+
+\section*{Supporting Information}
+Web Appendices, Tables, and Figures referenced in Sections \ref{introduction} and \ref{numerical results} are available with this paper at the Biometrics website on Wiley Online Library.
+The procedures are implemented in the R package \verb|cole|, available on GitHub at \url{http://github.com/sdzhao/cole}.
 

### 2022-03-16 msg.tex
update reference & abstrct

@@
 This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class.
-Simulation results and gene network inference in an RNA-seq experiment in mouse show that our approach is comparable to or can outperform a number of state-of-the-art proposals, particularly when the sample eigenvectors are poor estimates of the population eigenvectors.
+Simulation results and gene network inference in an RNA-seq experiment in mouse show that our approach is comparable to or can outperform a number of state-of-the-art proposals.
 \end{abstract}
@@
   \end{center}
-\caption{Amygdala gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
+\caption{Amygdala gene networks recovered by the top performers in Table \ref{tab:tab1}.
+Line types indicate the method(s) that recovered each each edge.
+This figure appears in color in the electronic version of this article, and any mention of color refers to that version.}
 \label{network_A}
@@
   \end{center}
-\caption{Frontal cortex gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
+\caption{Frontal cortex gene networks recovered by the top performers in Table \ref{tab:tab1}.
+Line types indicate the method(s) that recovered each each edge.
+This figure appears in color in the electronic version of this article, and any mention of color refers to that version.}
 \label{network_FC}
